---
title: "Synoptic_CB_Nutrients_2023_AnalysisTemplate"
author: "Stephanie J. Wilson"
date: "2025-05-05"
output: html_document
---

---
title: "COMPASS Synoptic Discrete Data Workflow: Example Worflow"
author: "Stephanie J. Wilson & Stephanie Pennington"
date: '2024-10-08'
output: html_document
---

##Setup

```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#a link to the Gitbook or whatever protocol you are using for this analysis 


#Maybe using pacman instead?
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  ggplot2,
  ggpubr,
  stringr,
  readr,
  broom,
  data.table)

#Packages that are required 
  library(dplyr)
  library(ggplot2)
  library(ggpubr)
  library(stringr)
  library(readr)
  library(stringr)

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
P_mw <- 30.973762  # molecular weight of P 
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 


#expected ranges for sample concentrations used for flags 
  # min_conc = int
  # max_conc = int
  # cv_flag = 10% 
  # rep_flag = 10%

#check standard concentrations 
  # chk_std = 10
  

#any reference to other code 



```

## Import Data

```{r Import Data, echo=FALSE}
cat("Import Data")

#setwd("/Raw Data")

#set working directory
  #could be from google drive or local drive etc. 
path <- ("file path")

#---------------------

#---------------------NOX---------------------
#Read in Raw Data 
dat_NOx <- readr::read_csv("Raw Data/SEAL_COMPASS_Synoptic_NOx_April2023_1.csv") 
dat_NH3_PO4 <- readr::read_csv("Raw Data/SEAL_COMPASS_Synoptic_NH3_PO4_April2023_1.csv")
# Combining Files
df_combo <- bind_rows(dat_NOx, dat_NH3_PO4)

# Rename columns
df_combo_rename <- df_combo %>%
  rename('Conc' = ...6,
         'Absorbance' = ...7,
         'Dilution' = ...9,
         'Sample_Name' = ...4,
         'Test' = ...13,
         'Run_Time' = ...15,
         'Unit' = ...12)

#making list to remove columns
cols_to_remove <- c("RUNSTARTED", "...5", "...8", "...10", "...11", "...14")

#Removing columns
df_all <- df_combo_rename %>%
  select(-all_of(cols_to_remove)) %>%
  select(!c(1,2))%>%
  select(!c(8,9)) 

#Checking column headers
head(df_all)


```

## Assessing standard Curves

```{r Assess Standard Curves, echo=FALSE}
cat("Assess Standard Curves")


#Pull out standards 
stds <- df_all %>%
  filter(str_detect(Sample_Name, "Standard"))

#Making standards dataframe based on the test
stds_NOx <- stds[stds$Test == "Vanadium NOx", ]
stds_NH3 <- stds[stds$Test == "Ammonia 2", ]
stds_PO4 <- stds[stds$Test == "o-PHOS 0.3", ]

#Inputting Standard Values Based on Protocol (LINK PROTOCOL)
stds_NOx <- stds_NOx %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard_1" ~ 1.0,
    Sample_Name == "Standard_90" ~ 0.0222,
    Sample_Name == "Standard_91" ~ 0.05,
    Sample_Name == "Standard_92" ~ 0.1,
    Sample_Name == "Standard_93" ~ 0.25,
    Sample_Name == "Standard_94" ~ 0.5,
    Sample_Name == "Standard_95" ~ 0.75,
    Sample_Name == "Standard_96" ~ 1.0,
    TRUE ~ `Conc`  # leave unchanged if no match
  ))

stds_NH3 <- stds_NH3 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard_1" ~ 2.0,
    Sample_Name == "Standard_90" ~ 0.0389,
    Sample_Name == "Standard_91" ~ 0.1,
    Sample_Name == "Standard_92" ~ 0.2,
    Sample_Name == "Standard_93" ~ 0.5,
    Sample_Name == "Standard_94" ~ 1.0,
    Sample_Name == "Standard_95" ~ 1.5,
    Sample_Name == "Standard_96" ~ 2,
    TRUE ~ `Conc`  # leave unchanged if no match
  ))

stds_PO4 <- stds_PO4 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard_1" ~ 0.3,
    Sample_Name == "Standard_90" ~ 0.0060,
    Sample_Name == "Standard_91" ~ 0.0150,
    Sample_Name == "Standard_92" ~ 0.0300,
    Sample_Name == "Standard_93" ~ 0.0750,
    Sample_Name == "Standard_94" ~ 0.1500,
    Sample_Name == "Standard_95" ~ 0.2250,
    Sample_Name == "Standard_96" ~ 0.3000,
    TRUE ~ `Conc`  # leave unchanged if no match
  ))


#generating line of best fit aka standard curves
#NOx
lin_reg_NOx <- lm(stds_NOx$Conc ~stds_NOx$Absorbance, stds_NOx)
NOx_coefs <- coef(lin_reg_NOx)                        # intercept and slope
NOx_r2 <- summary(lin_reg_NOx)$r.squared              # R squared
# Store in a clean dataframe
std_curve_NOx_1 <- data.frame(
  Test = "NOx",
  Intercept = NOx_coefs["(Intercept)"],
  Slope = NOx_coefs[2],
  R_squared = NOx_r2
)

#NH3
lin_reg_NH3 <- lm(stds_NH3$Conc ~stds_NH3$Absorbance, stds_NH3)
NH3_coefs <- coef(lin_reg_NH3)                        # intercept and slope
NH3_r2 <- summary(lin_reg_NH3)$r.squared              # R squared
# Store in a clean dataframe
std_curve_NH3_1 <- data.frame(
  Test = "NH3",
  Intercept = NH3_coefs["(Intercept)"],
  Slope = NH3_coefs[2],
  R_squared = NH3_r2
)

#PO4
lin_reg_PO4 <- lm(stds_PO4$Conc ~stds_PO4$Absorbance, stds_PO4)
PO4_coefs <- coef(lin_reg_PO4)                        # intercept and slope
PO4_r2 <- summary(lin_reg_PO4)$r.squared              # R squared
# Store in a clean dataframe
std_curve_PO4_1 <- data.frame(
  Test = "PO4",
  Intercept = PO4_coefs["(Intercept)"],
  Slope = PO4_coefs[2],
  R_squared = PO4_r2
) # WANT TO FIND A WAY TO GET THE DATE IN THERE


#Combining standards & curve data frames
stds_combo <- bind_rows(stds_NOx, stds_NH3, stds_PO4)
std_curve_combo <- bind_rows(std_curve_NOx_1, std_curve_NH3_1, std_curve_PO4_1)


#Plot standard Curve or Curves 
#NOx INCORRECT 
std_curve_NOx_test <- ggplot(stds_NOx, aes(x = Absorbance, y = Conc)) +
  geom_point(size = 3) +  # scatter plot of standards
  geom_abline(
    intercept = std_curve_NOx_1$Intercept,
    slope = std_curve_NOx_1$Slope,
    color = "orange2",
    linetype = "solid"
  ) +
  theme_minimal() +
  labs(title = "NOx Standard Curve") 
print(std_curve_NOx_test)

#NH3
std_curve_NH3 <- ggplot(stds_NH3, aes(x = Absorbance, y = Conc)) +
  geom_point(size = 3) +  # scatter plot of standards
  geom_abline(
    intercept = std_curve_NH3_1$Intercept,
    slope = std_curve_NH3_1$Slope,
    color = "green4",
    linetype = "solid"
  ) +
  theme_minimal()+
  labs(title = "NH3 Standard Curve")
print(std_curve_NH3)

std_curve_PO4 <- ggplot(stds_PO4, aes(x = Absorbance, y = Conc)) +
  geom_point(size = 3) +  # scatter plot of standards
  geom_abline(
    intercept = std_curve_PO4_1$Intercept,
    slope = std_curve_PO4_1$Slope,
    color = "blue3",
    linetype = "solid"
  ) +
  theme_minimal()+
  labs(title = "PO4 Standard Curve")
print(std_curve_PO4)


#Report out a flag if the run has an R2 lower than appropriate 
  #what is going to be our cut off



#compare slopes to previous runs (from log) in order to assess drift 
  #want to get a way to have the run date in the stds_combo file so we can continuously export it into a sheet


#write out slopes from current run to log 

```

## Sample Processing - Calculating Concentrations

```{r Sample Processing, echo=FALSE}

cat("Sample Processing")

#Remove standards or other checks not needed here if necessary 

#Pull out samples and adding metadata
df_all_clean_1 <- df_all[df_all$Sample_Name %like% "MSM_", ]
df_all_clean_2 <- rbind(df_all_clean_1, (df_all[df_all$Sample_Name %like% "GWI_", ]))
df_all_clean_3 <- rbind(df_all_clean_2, (df_all[df_all$Sample_Name %like% "GCW_", ])) 

df_all_clean <- df_all_clean_3 %>%
  mutate(separate(
    df_all_clean_3,
    col = Sample_Name,
    sep = "_",
    into = c("Site", "Samp_Date", "Position", "Replicate", "Depth"))) %>%
   mutate(Samp_Date = lubridate::ym(Samp_Date)) %>%
   mutate(Run_Time = lubridate::mdy_hm(Run_Time))

#Calculate Concentrations from Standard Curve & Unit Conversion
df_all_calc <- df_all_clean %>% 
mutate(Conc_Calc = case_when(
      Test == "Ammonia 2" ~ lin_reg_NH3$coefficients["(Intercept)"] + 
              lin_reg_NH3$coefficients["stds_NH3$Absorbance"] * Absorbance,
      Test == "o-PHOS 0.3" ~ lin_reg_PO4$coefficients["(Intercept)"] + 
              lin_reg_PO4$coefficients["stds_PO4$Absorbance"] * Absorbance,
     # Test == "Vanadium NOx "~ lin_reg_NOx$coefficients["(Intercept)"] + 
     #         lin_reg_NOx$coefficients["stds_NOx$Absorbance"] * Absorbance,
      TRUE ~ NA_real_)) %>%
  mutate(`Conc_Calc_uM` = case_when(
    Test == "Ammonia 2" ~ ((Conc_Calc/Con1)/N_mw)*2,
    Test == "o-PHOS 0.3" ~ ((Conc_Calc/Con1)/P_mw)*2,
    #Test == "Vanadium NOx" ~ ((Conc_Calc/Con1)/N_mw)*2
  ))


#add step to make negative values equal to bd (below detection) and replace with zeros 
df_all_calc$Conc_Calc_uM <- ifelse(df_all_calc$Conc_Calc_uM<0, 0, (df_all_calc$Conc_Calc_uM) )
head(df_all_calc)


```

## Sample Flagging - Within standards range?

```{r Sample Processing, echo=FALSE}

cat("Sample Flagging")

#Flagging data if the concentration is outside the standards range 

df_all_calc_chk <- df_all_calc %>% 
mutate(Range = case_when(
      Test == "Ammonia 2" ~ ifelse(Conc < 0.02,"bdl", ifelse(Conc > 2, "adl", "Within_Range")),
      Test == "o-PHOS 0.3" ~ ifelse(Conc < 0.003,"bdl", ifelse(Conc > 3, "adl", "Within_Range")),
     # Test == "Vanadium NOx "~ ifelse(Conc < 0.025,"bdl", ifelse(Conc > 1, "adl", "Within_Range")),
      TRUE ~ NA_character_)) %>%
  mutate(`Conc_Calc_uM` = case_when(
    Test == "Ammonia 2" ~ ((Conc_Calc/Con1)/N_mw)*2,
    Test == "o-PHOS 0.3" ~ ((Conc_Calc/Con1)/P_mw)*2,
    #Test == "Vanadium NOx" ~ ((Conc_Calc/Con1)/N_mw)*2
  ))


#Flagging for CV of analyticalreplicates 





```

## Analyze the Check Standards

```{r Check Standards, echo=FALSE}

cat("Analysis Check Standards")

#this pechk was made on 10/30/24. 
#check PECHK_Concentrations file in the Seal teams to make sure the correct concentrations are below
pechk_pred_NH3 = 0.97
pechk_pred_PO4 = 0.74
pechk_pred_NOx = 1.40


#Pull out check standards from raw file 


#Calculate their concentrations if not done already 


#change this if check standard concentrations are different 


#calculate percent difference between check standards & expected concentration 


#flag if the percent difference is over X% (defined in setup)


#Visualize check std concentrations vs. the expected concentrations


#report out if flags indicate need for rerun


```

## Dilution Corrections

```{r Dilution Corrections}

cat("Dilution Corrections")

#Calculate the concentration when accounting for the dilution factor if applicable



```

## Visualize Data

```{r Visualize Data}

cat("Visualize Data")

#Plot samples to get a first look at concentrations (sanity check)


```

## Export Processed Data

```{r, Export Processed Data}

cat("Export Processed Data")

#set working directory 
finalpath <- "file path"

#Prepare data to be exported 

#Add any necessary identifiers to the samples  ### VERY IMPORTANT AND STANDARD FOR PROJECT ####
  #example read in sample IDs list and merge 
  #create required ID columns in R, etc. 

#Write out data frame 
write.csv(alldat, "file_name")

```

#end
