---
title: "2022_Lysimeter_vs_Sipper"
author: "Zoe Read"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(lubridate)
```

##Pull in 2022 Data
```{r}

##Lysimeter Data
lys <- read.csv("../COMPASS_SynopticCB_AllData_AllYears.csv")
lys_2022 <- lys %>% 
  filter(str_detect(Sample_ID, "2022")) %>% 
  mutate(Replicate = paste0("LYS", Lysimeter)) %>% 
  select(-Lysimeter) %>% 
  filter(Zone != "SW")
colnames(lys_2022)

##Sipper Data
sip_2022 <- read.csv("COMPASS_SynopticCB_Sipper_Data.csv")
colnames(sip_2022)

all_2022 <- bind_rows(lys_2022, sip_2022)

##Add labels for sampler types
all_data_long <- all_2022 %>%  
  mutate(Sampler = substring(Replicate, 1, 3),
         Replicate = substring(Replicate, 4, nchar(Replicate))) %>%  
  mutate(Depth_cm = as.character(Depth_cm), 
         Month = as.character(Month))

```



##Boxplots SO4/Cl
```{r}

boxplot_SO4_all <- ggplot(all_data_long, aes(x = Depth_cm, y = SO4_ppm, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_grid(Zone ~ Site) +
  labs(y = "SO4 (ppm)", fill = "Sampler", title = "Sulfate from all data") +
  theme_bw()

boxplot_SO4_all

##Subset for only site and zone that overlap
df_sub <- subset(all_data_long, Site == "GCW" & Zone == "WC")

boxplot_SO4_GCW_WC <- ggplot(df_sub, aes(x = Depth_cm, y = SO4_ppm, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_grid(Zone ~ Site) +
  labs(y = "SO4 (ppm)", fill = "Sampler", title = "Sulfate from GCW WC") +
  theme_bw()

boxplot_SO4_GCW_WC

##Subset for only dates that overlap
dates <- c("202206","202207", "202208") 
df_sub <- all_data_long %>% 
  filter(Site == "GCW" & Zone == "WC" & Month == "6" | Site == "GCW" & Zone == "WC" & Month == "7"| Site == "GCW" & Zone == "WC" & Month == "8")

boxplot_Cl <- ggplot(df_sub, aes(x = Depth_cm, y = Cl_ppm, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_wrap(~ Month) +
  labs(y = "Cl (ppm)", fill = "Sampler", title = "Chloride from GCW WC 2022 by Month") +
  theme_bw()

boxplot_Cl

boxplot_SO4 <- ggplot(df_sub, aes(x = Depth_cm, y = SO4_ppm, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_wrap(~ Month) +
  labs(y = "SO4 (ppm)", fill = "Sampler", title = "Sulfate from GCW WC 2022 by Month") +
  theme_bw()

boxplot_SO4

##Average by sampler type at each depth
df_avg <- df_sub %>% 
  group_by(Sampler, Month, Depth_cm) %>% 
  summarize(SO4_ppm_avg = mean(SO4_ppm, na.rm = TRUE), 
            Cl_ppm_avg = mean(Cl_ppm, na.rm = TRUE))

df_avg_lys <- df_avg %>% 
  filter(Sampler == "LYS") %>%
  rename_with(
    .cols = c(4:(ncol(.))),
    .fn = ~paste0("LYS_", .))  %>%
  ungroup(Sampler)%>%
  select(-Sampler)

df_avg_sip <- df_avg %>% 
  filter(Sampler == "SIP") %>%
  rename_with(
    .cols = c(4:(ncol(.))),
    .fn = ~paste0("SIP_", .)) %>%
  ungroup(Sampler) %>% 
  select(-Sampler)

df_avg1 <- df_avg_lys %>% 
  left_join(df_avg_sip)

##1:1 Plot
so4_scatter_plot <- ggplot(df_avg1, aes(x = LYS_SO4_ppm_avg, y = SIP_SO4_ppm_avg, color = as.character(Month))) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # coord_cartesian(xlim = combined_limits, ylim = combined_limits) +
  coord_fixed(ratio = 1) + 
  labs(
    x = "Lysimeter SO4 (ppm)",
    y = "Sipper SO4 (ppm)",
    title = ""
  ) + 
  theme_classic()

so4_scatter_plot

cl_scatter_plot <- ggplot(df_avg1, aes(x = LYS_Cl_ppm_avg, y = SIP_Cl_ppm_avg, color = as.character(Month))) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # coord_cartesian(xlim = combined_limits, ylim = combined_limits) +
  coord_fixed(ratio = 1) + 
  labs(
    x = "Lysimeter Cl (ppm)",
    y = "Sipper Cl (ppm)",
    title = ""
  ) + 
  theme_classic()

cl_scatter_plot
```
#Read in and collate all the files in the Processed Data folder 
```{r}
files <- list.files(path = "../Nutrients/Synoptic-CB/2022/Processed Data", pattern = "\\.csv$", full.names = TRUE)

# 3. Read and combine all CSVs
all_data <- files %>%
  lapply(read.csv, colClasses = "character") %>%    # or read.csv if you prefer base R
  bind_rows()             # combine into one data frame

all_data <- all_data %>%
  select("Date" ,"Site", "Zone" , "Replicate", "Depth", "Sample_Name",  "NH3_uM", "PO4_uM", "NOx_uM")

colnames(all_data)

all_data$NH3_uM <- as.numeric(all_data$NH3_uM)
all_data$PO4_uM <- as.numeric(all_data$PO4_uM)
all_data$NOx_uM <- as.numeric(all_data$NOx_uM)

##Fix sample IDs
all_data$Site <- gsub("GCerw", "GCW", all_data$Site)
all_data$Site <- gsub("Gcrew", "GCW", all_data$Site)
all_data$Site <- gsub("GCrew", "GCW", all_data$Site)

##Add labels for sampler types
all_data_long <- all_data %>%  
  mutate(Sampler = substring(Replicate, 1, 3),
         Replicate = substring(Replicate, 4, nchar(Replicate)))%>%  
  filter(Sampler %in% c("Lys", "Sip"))

```

##Boxplots Nutrients
```{r}

boxplot_NH3_all <- ggplot(all_data_long, aes(x = Depth, y = NH3_uM, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_grid(Zone ~ Site) +
  labs(y = "NH3_uM (uM)", fill = "Sampler", title = "NH3_uM from all data") +
  theme_bw()

boxplot_NH3_all

##Subset for only site and zone that overlap
df_sub <- subset(all_data_long, Site == "GCW" & Zone == "WC")

boxplot_NH3_GCW_WC <- ggplot(df_sub, aes(x = Depth, y = NH3_uM, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_grid(Zone ~ Site + Date) +
  labs(y = "NH3 (uM)", fill = "Sampler", title = "NH3_uM from GCW WC") +
  theme_bw()

boxplot_NH3_GCW_WC

##Subset for only dates that overlap
dates <- c("202206","202207", "202208") 
df_sub <- subset(all_data_long, Site == "GCW" & Zone == "WC" & Date %in% dates)
# str(df_sub)

boxplot_NH3 <- ggplot(df_sub, aes(x = Depth, y = NH3_uM, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_wrap(~ Date) +
  labs(y = "NH3 (uM)", fill = "Sampler", title = "NH3 from GCW WC 2022") +
  theme_bw()

boxplot_NH3

boxplot_SO4 <- ggplot(df_sub, aes(x = Depth, y = SO4_ppm, fill = Sampler)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_wrap(~ Date) +
  labs(y = "SO4 (ppm)", fill = "Sampler", title = "Sulfate from GCW WC 2022") +
  theme_bw()

boxplot_SO4

# Save as a PNG
png("Lysimeter_vs_Sipper_Cl.png", width = 2000, height = 1200, res = 200) 
print(boxplot_Cl)  # For ggplot objects, use print()
dev.off() 

# Save as a PNG
png("Lysimeter_vs_Sipper_SO4.png", width = 2000, height = 1200, res = 200) 
print(boxplot_SO4)  # For ggplot objects, use print()
dev.off() 

```



## Plotting values against each other
```{r Plotting values against each other, echo=FALSE, warning=FALSE}

data_abbrev <- data_plotting %>% 
  dplyr::select(Acid_state_1, FE_2_mean, FE_3_mean, FE_tot_mean, new_ID) %>% 
  filter(Acid_state_1 != "No Acid")

# head(data_abbrev)


##Reshape Data
df_long <- data_abbrev %>%
  dplyr::select(new_ID, Acid_state_1, FE_2_mean, FE_3_mean, FE_tot_mean) %>%
  pivot_longer(
    cols = c(FE_2_mean, FE_3_mean, FE_tot_mean),
    names_to = "Fe_type",
    values_to = "value"
  )

df_wide <- df_long %>%
  pivot_wider(
    names_from = Acid_state_1,
    values_from = value
  )

##Select only Total FE Data
df_Fe_tot <- df_wide %>% 
    filter(Fe_type == "FE_tot_mean")

##Plot Total FE Data
scatter_Fetot_plot <- ggplot(df_Fe_tot, aes(x = `Acidified in Field`, y = `Acidified Later in Lab`)) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # coord_cartesian(xlim = combined_limits, ylim = combined_limits) +
  coord_fixed(ratio = 1) + 
  labs(
    x = "Acidified in Field",
    y = "Acidified Later in Lab",
    title = "Total Fe Measurements"
  ) + 
  theme_classic()

scatter_Fetot_plot


##Select only TEMPEST Data
df_TMP <- df_Fe_tot %>% 
    filter(str_detect(new_ID, "TMP"))

#Plot TEMPEST Fe Data: 
scatter_Fe_plot_TMP <- ggplot(df_TMP, aes(x = `Acidified in Field`, y = `Acidified Later in Lab`)) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # coord_cartesian(xlim = combined_limits, ylim = combined_limits) +
  coord_fixed(ratio = 1) + 
  labs(
    x = "Acidified in Field",
    y = "Acidified Later in Lab",
    title = "TMP Total Fe Measurements"
  ) + 
  theme_classic()

scatter_Fe_plot_TMP


```


####Still figuring these out####
```{r}
##Only Sip Samples
all_data_sip <- all_data %>%  
  filter(str_detect(Replicate, "Sip"))%>%  
  rename(
    Replicate_sip = Replicate, 
    Sample_ID_sip = Sample_ID, 
    SO4_ppm_sip = SO4_ppm, 
    Cl_ppm_sip = Cl_ppm, 
    SO4_mM_sip = SO4_mM, 
    Cl_mM_sip = Cl_mM, 
    salinity_sip = salinity
  )%>%  
  mutate(Replicate_sip = gsub("Sip", "", Replicate_sip))

##Only Lys Samples
all_data_lys <- all_data %>%  
  filter(str_detect(Replicate, "Lys"))%>%  
  rename(
    Replicate_lys = Replicate, 
    Sample_ID_lys = Sample_ID, 
    SO4_ppm_lys = SO4_ppm, 
    Cl_ppm_lys = Cl_ppm, 
    SO4_mM_lys = SO4_mM, 
    Cl_mM_lys = Cl_mM, 
    salinity_lys = salinity
  )%>%  
  mutate(Replicate_lys = gsub("Lys", "", Replicate_lys))

##Merge Datasets by Date, Site, Depth, and Zone
all_data_combined <- merge(all_data_lys, all_data_sip, by = c("Date", "Site", "Depth", "Zone"))
```


```{r}
##Metrics for each site, zone, depth for each year
summary_by_year_sip <- all_data_sip %>%
  group_by(Date, Site, Zone, Depth) %>%
   reframe(
     min_Cl_mM  = min(Cl_mM_sip , na.rm=TRUE),
     max_Cl_mM  = max(Cl_mM_sip , na.rm=TRUE), 
     mean_Cl_mM  = mean(Cl_mM_sip , na.rm=TRUE), 
     sd_Cl_mM  = sd(Cl_mM_sip , na.rm=TRUE), 

     min_SO4_mM  = min(SO4_mM_sip , na.rm=TRUE),
     max_SO4_mM  = max(SO4_mM_sip , na.rm=TRUE), 
     mean_SO4_mM  = mean(SO4_mM_sip , na.rm=TRUE), 
     sd_SO4_mM  = sd(SO4_mM_sip , na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 4))  %>%
  mutate(Type = "Sip") 

summary_by_year_lys <- all_data_lys %>%
  group_by(Date, Site, Zone, Depth) %>%
   reframe(
     min_Cl_mM  = min(Cl_mM_lys , na.rm=TRUE),
     max_Cl_mM  = max(Cl_mM_lys , na.rm=TRUE), 
     mean_Cl_mM  = mean(Cl_mM_lys , na.rm=TRUE), 
     sd_Cl_mM  = sd(Cl_mM_lys , na.rm=TRUE), 

     min_SO4_mM  = min(SO4_mM_lys , na.rm=TRUE),
     max_SO4_mM  = max(SO4_mM_lys , na.rm=TRUE), 
     mean_SO4_mM  = mean(SO4_mM_lys , na.rm=TRUE), 
     sd_SO4_mM  = sd(SO4_mM_lys , na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 4)) %>%
  mutate(Type = "Lys") 

##Merge Datasets by Date, Site, Depth, and Zone
summary_by_year_combined <- bind_rows(summary_by_year_sip, summary_by_year_lys)

##Convert Collection Date into Date format
summary_by_year_combined$Date <- as.character(summary_by_year_combined$Date)
# Append '01' for the first day of the month and convert
summary_by_year_combined$Date_formatted <- as.Date(paste0(summary_by_year_combined$Date, "01"), format = "%Y%m%d")

# ##Extract Year and month from Collection Date
# raw_dat$Year <- format(raw_dat$Collection_Date_formatted, format = "%Y")
# raw_dat$Month <- format(raw_dat$Collection_Date_formatted, format = "%m")
# raw_dat$Day <- format(raw_dat$Collection_Date_formatted, format = "%d")


```

## Summarized data for Site and Zone 
```{r Summarize and Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Summarize SO4 and Cl by Date and Zone
# Dionex_avg <- data_plotting %>%
#   group_by(Collection_Date_formatted, Zone_1) %>%
#    dplyr::summarize(mean_SO4 = mean(SO4_ppm, na.rm=TRUE),
#             sd_SO4 = sd(SO4_ppm, na.rm=TRUE),
#             mean_Cl = mean(Cl_ppm, na.rm=TRUE),
#             sd_Cl = sd(Cl_ppm, na.rm=TRUE))

# Define the specific date breaks you want to display
specific_breaks <- as.Date(c("2024-01-01", "2025-01-01"))

viz_cl_plot_avgs <- ggplot(summary_by_year_combined, aes(x = Date_formatted, y = mean_Cl_mM, color = Type)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_Cl_mM - sd_Cl_mM, ymax = mean_Cl_mM + sd_Cl_mM), width = 0.2) + 
  geom_line() + 
  facet_wrap( Site ~ Zone, scales="free_y") +
  # scale_color_manual(values = c(
  #   "Control" = "springgreen2",
  #   "Freshwater" = "cyan2",
  #   "Saltwater" = "violetred2"
  # )) +
  theme_classic() +
  labs(x = "Date", y = "Chloride (ppm)", title = "Samples: Chloride") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") + # Format x-axis labels
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

viz_cl_plot_avgs


viz_so4_plot_avgs <- ggplot(Dionex_avg, aes(x = Collection_Date_formatted, y = mean_SO4, color = Zone_1)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_SO4 - sd_SO4, ymax = mean_SO4 + sd_SO4), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ Zone_1, scales="free_y", ncol = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = "Date", y = "Sulfate (ppm)", title = "Samples: Sulfate") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") + # Format x-axis labels
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

viz_so4_plot_avgs

# ggarrange(viz_cl_plot_avgs, viz_so4_plot_avgs, nrow=1, ncol=2)

```

## Boxplot summary
```{r Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}
##Want boxplot of Sip vs Lys data for same Date, Site, Depth, and Zone
##Looks like only comparisons possible for GCW WC for different dates and depths
##Need to make date actual date value and maybe facet by depth? 

summary_by_year_combined_GCW <- subset(summary_by_year_combined, Site == "GCW")
summary_by_year_combined_GCW_WC <- subset(summary_by_year_combined_GCW, Zone == "WC")


##Summarize SO4 and Cl by Date, Site, and Zone
viz_cl_boxplot <- ggplot(summary_by_year_combined_GCW_WC, aes(x = Depth, y = mean_Cl_mM, fill = Type)) +
  geom_boxplot() + 
  facet_grid( Site ~ Date ) +
  scale_x_discrete(drop = TRUE) +
  theme_classic() +
  labs(x = " ", y = "Chloride (mM)", title = "Synoptic CB Porewater: Chloride") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))

viz_cl_boxplot


viz_so4_boxplot <- ggplot(all_dat_dates, aes(x = Zone_1, y = SO4_mM, fill = Zone_1, group = Zone_1)) +
  geom_boxplot() +
  facet_grid( ~ Site, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "Sulfate (mM)", title = "Samples: Sulfate") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 
  theme(legend.position = "none")

viz_so4_boxplot

# ggarrange(viz_cl_boxplot, viz_so4_boxplot, nrow=2, ncol=1)

# Save as a PNG
png("COMPASS_SynopticCB_Cl_alldata.png", width = 2000, height = 1200, res = 100) 
print(viz_cl_boxplot)  # For ggplot objects, use print()
dev.off() 

# Save as a PNG
png("COMPASS_SynopticCB_SO4_alldata.png", width = 2000, height = 1200, res = 100) 
print(viz_so4_boxplot)  # For ggplot objects, use print()
dev.off() 

```

```{r}
# Summary stats for Sip only
sip_sum <- df_sub %>%
  filter(Sampler == "Sip") %>%
  group_by(Depth, Date) %>%
  summarise(
    mean = mean(SO4_ppm, na.rm = TRUE),
    sd = sd(SO4_ppm, na.rm = TRUE),
    .groups = "drop"
  )


##Plot: raw data + Sip mean ± SD + Lys points
ggplot() +
  geom_jitter(
    data = subset(df_sub, Sampler == "Sip"),
    aes(x = Depth, y = SO4_ppm, color = Sampler),
    width = 0.15,
    size = 3
  ) +

  #Sip mean ± SD
  geom_errorbar(
    data = sip_sum,
    aes(x = Depth, ymin = mean - sd, ymax = mean + sd),
    width = 0.2
  ) + 

  geom_point(
    data = sip_sum,
    aes(x = Depth, y = mean),
    size = 4,
    shape = 21,
    fill = "steelblue"
  ) +

  # Lys single points
  geom_point(
    data = subset(df_sub, Sampler == "Lys"),
    aes(x = Depth, y = SO4_ppm, shape = Sampler),
    shape = 24,
    size = 3,
    fill = "white",
    color = "black"
  ) +

  labs(
    x = "Depth",
    y = "Sulfate (ppm)",
    title = "Sulfate concentration by sampler and depth",
    subtitle = "Sip shown as replicates and mean ± SD; Lys shown as single samples"
  ) +

  theme_bw() + 
  
  facet_wrap(~Date) + 
  
  labs(
  color = "Sampler type",
  shape = "Sampler type")


```


```

## Write the 2022 file out to the folder 
```{r, echo=FALSE}

head(merged_data)

final_data <- merged_data %>%
  rename(
    Time = Time..24hr.,
    Time_Zone = Time.Zone_EDT.EST,
    Field_notes = Field.Notes, 
    SO4_mM = SO4_Conc_mM, 
    Cl_mM = Cl_Conc_mM
    # add more rename pairs as needed
  ) %>%
  select(
    Project, Region, Site, Zone, Replicate, Depth_cm, 
         Sample_ID, Year, Month, Day, Time, Time_Zone,
         SO4_ppm, SO4_mM, SO4_Conc_flag, SO4_QAQC_flag, Cl_ppm, Cl_mM, Cl_Conc_flag, Cl_QAQC_flag, salinity,
         Analysis_rundate,  Run_notes, Field_notes
        # list columns in the order you want them
  )

head(final_data)

write.csv(final_data, "COMPASS_SynopticCB_SO4_Cl_2022.csv")

```

##END