---
title: "Synoptic CB: Soil Fluxes"
author: "2025 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

output_dir: "To Be Reviewed/PDF"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(openxlsx)
library(arrow)
library(purrr)
library(lubridate)
```

#Read in and collate all the files in the Processed Data folder 
```{r, include = FALSE}

ch4 <- read.csv("Processed Data/COMPASS_Synoptic_CB_SmartChamber_Fluxes_Final_2025_CH4.csv")
ch4 <- ch4 %>%
  rename_with(
    .cols = 8:(ncol(.) - 6),
    .fn = ~paste0("ch4_", .)
  ) %>%
  rename(ch4_qaqc_flag = qaqc_flag) %>% 
  mutate(
    TIMESTAMP <- ymd_hms(TIMESTAMP)
  )

##Need to figure out why some timestamps, probably the ones I edited, are output differently to the processed Excel file and then read in as long decimal places. 

co2 <- read.csv("Processed Data/COMPASS_Synoptic_CB_SmartChamber_Fluxes_Final_2025_CO2.csv")
co2 <- co2 %>%
  rename_with(
    .cols = 8:(ncol(.) - 6),
    .fn = ~paste0("co2_", .)
  ) %>%
  rename(co2_qaqc_flag = qaqc_flag)

# Combine all files
all_dat <- merge(ch4, co2, by = c("RepNumber", "Month", "Site", "Zone", "Replicate", "Plot", "TIMESTAMP", "TIMESTAMP_min", "TIMESTAMP_max", "soilp_c", "soilp_m", "soilp_t"), all = TRUE)

##Rename Zone
all_dat$Zone <- gsub("W", "WC", all_dat$Zone)
all_dat$Plot <- gsub("_W_", "_WC_", all_dat$Plot)

```


##Pull in soil probe data
```{r, echo = FALSE}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
zip_contents <- unzip("Raw Data.zip", list = TRUE)

# Keep only JSON files
json_files <- zip_contents$Name[grepl("json$", zip_contents$Name, ignore.case = TRUE)]

# Extract only the JSON files
unzip(
  "Raw Data.zip",
  files = json_files,
  exdir = tmp_dir
)

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "json$",
  full.names = TRUE,
  recursive = TRUE
)

head(file.list)

# Read in the data
alldat1 <- lapply(file.list, ffi_read_LIsmartchamber)

# Combine into one data.table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)


##Create a dataframe with averaged soil probe data per Plot and add to flux data
soildat <- alldat %>% 
    group_by(Plot) %>%
  dplyr::summarize(
     soilp_c = mean(soilp_c, na.rm=TRUE),
     soilp_m = mean(soilp_m, na.rm=TRUE),
     soilp_t = mean(soilp_t, na.rm=TRUE))

# allflux <- left_join(fluxes, soildat, by = "Plot")

```


##pull TEROS data for the zone during the time of the fluxes and put it into another column so that we can use that too.
```{r, include = FALSE}
# Soil sensors

##Read in L1 Sensor Data from S Drive and only save necessary data
# Path to zip file
L2 <- "S:/Biogeochemistry/COMPASS/L2 Data/v2-1"

# List CSV files inside the zip
files <- list.files(path = L2, recursive = TRUE, full.names = TRUE)

files <- files[
  grepl("soil", files, ignore.case = TRUE) &
  grepl("10cm", files, ignore.case = TRUE) &
  grepl("SWH|GCW|MSM|GWI", files, ignore.case = TRUE) &
  grepl("2025", files, ignore.case = TRUE) &
  grepl("\\.parquet$", files, ignore.case = TRUE)
]

teros <- files %>%
  map_dfr(read_parquet)

#Change W Plot to WC
teros <- teros %>%
  mutate(Plot = ifelse(Plot == "W", "WC", Plot))

##Want to go from long to wide format
teros_wide <- pivot_wider(teros, id_cols = c("TIMESTAMP", "Plot", "Site", "Instrument", "Instrument_ID", "Sensor_ID"), names_from = research_name, values_from = Value )

##Average for each TIMESTAMP by Site, Plot, and Timestamp
teros_wide_avg <- teros_wide %>%
  group_by(Site, Plot, TIMESTAMP) %>%
   dplyr::summarize(
     soil.EC.10cm.TEROS = mean(`soil-EC-10cm`, na.rm=TRUE),
     soil.salinity.10cm.TEROS = mean(`soil-salinity-10cm`, na.rm=TRUE),
     soil.temp.10cm.TEROS = mean(`soil-temp-10cm`, na.rm=TRUE),
     soil.vwc.10cm.TEROS = mean(`soil-vwc-10cm`, na.rm=TRUE))

##Write out this dataframe
write.csv(teros_wide_avg, file="Raw Data/TEROS_data.csv", row.names = FALSE)

teros_wide_avg <- read.csv("Raw Data/TEROS_data.csv")
colnames(teros_wide_avg)
```

## Match up the TEROS soil data and add it to the flux dataframe
```{r, include = FALSE}
#Going to need to use site and timestamp in the metadata in order to find a date 

#first we need to make a timestamp column in metadat so that we can 
final_dat1$TIMESTAMP <- ymd_hms(final_dat1$TIMESTAMP)

#make sure time stamps are in the correct format 
final_dat1$TIMESTAMP <- as.POSIXct(final_dat1$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")
teros_wide_avg$TIMESTAMP <- as.POSIXct(teros_wide_avg$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")

#time rounding step to closest 15 minutes
final_dat1$DateTime_round <- round_date(final_dat1$TIMESTAMP, "15 mins")
teros_wide_avg$TIMESTAMP_round <- round_date(teros_wide_avg$TIMESTAMP, "15 mins")
  
#ensure there is only one data observation per hour 
teros_wide_avg_clean <- teros_wide_avg %>% 
    group_by(TIMESTAMP_round, Site, Plot) %>%
    summarize(
     soil.EC.10cm.TEROS = mean(soil.EC.10cm.TEROS, na.rm=TRUE),
     soil.salinity.10cm.TEROS = mean(soil.salinity.10cm.TEROS, na.rm=TRUE),
     soil.temp.10cm.TEROS = mean(soil.temp.10cm.TEROS, na.rm=TRUE),
     soil.vwc.10cm.TEROS = mean(soil.vwc.10cm.TEROS, na.rm=TRUE))

#join them 
final_dat2 <- final_dat1 %>% 
  left_join(teros_wide_avg_clean, by=c("Site"="Site", "Zone"="Plot", "DateTime_round"="TIMESTAMP_round")) %>% 
  select(-DateTime_round)

```

##Remove bad fluxes and calculate how many there were
```{r, echo=FALSE}
##Remove any samples with "remove" in the qaqc_flag columns
final_dat <- final_dat2 %>% 
  filter(!grepl("remove", ch4_qaqc_flag, ignore.case = TRUE)) %>% 
  filter(!grepl("remove", co2_qaqc_flag, ignore.case = TRUE)) 

##Calculate how many bad fluxes in each category
##Percent fluxes removed: 
#Total
remove_dat <- all_dat %>% 
  filter(grepl("remove", ch4_qaqc_flag, ignore.case = TRUE) | grepl("remove", co2_qaqc_flag, ignore.case = TRUE))
perc_removed = nrow(remove_dat)/nrow(all_dat)*100

#Negative CO2
neg_co2 <- all_dat %>% 
  filter(grepl("negative CO2", co2_qaqc_flag, ignore.case = TRUE))
perc_neg_co2 = nrow(neg_co2)/nrow(all_dat)*100

#Unusually high CH4/CO2 values
unusual <- all_dat %>% 
  filter(grepl("Unusual", ch4_qaqc_flag, ignore.case = TRUE))
perc_unusual = nrow(unusual)/nrow(all_dat)*100

#Poor/messy flux
messy <- all_dat %>% 
  filter(grepl("messy", ch4_qaqc_flag, ignore.case = TRUE) | grepl("poor", co2_qaqc_flag, ignore.case = TRUE))
perc_messy = nrow(messy)/nrow(all_dat)*100

##Table with  percentage and type of issue

cat(paste0("Number flux measurements removed: ", nrow(remove_dat)))
cat(paste0("Number negative CO2 fluxes removed: ", nrow(neg_co2)))
cat(paste0("Number unusually high starting value fluxes removed: ", nrow(unusual)))
cat(paste0("Number messy fluxes removed: ", nrow(messy)))

cat(paste0("Percentage flux measurements removed: ", round(perc_removed, digits = 2), "%"))
cat(paste0("Percentage negative CO2 fluxes removed: ", round(perc_neg_co2, digits = 2), "%"))
cat(paste0("Percentage unusually high starting value fluxes removed: ", round(perc_unusual, digits = 2), "%"))
cat(paste0("Percentage messy fluxes removed: ", round(perc_messy, digits = 2), "%"))


```

## Write out files
```{r write out combined data and metadata, include=FALSE}

#write out a csv of all the data to the main folder: 
write.csv(final_dat, "COMPASS_SynopticCB_MC_Fluxes_2025.csv")

```


## All Data by Year, Site and Zone
```{r Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting <- final_dat

#Order by Zone from Upland to Surface Water
data_plotting$Zone = factor(data_plotting$Zone, levels = c("UP", "SWAMP", "TR", "WC", "SW"))
data_plotting <- data_plotting[order(data_plotting$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting$Site = factor(data_plotting$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting <- data_plotting[order(data_plotting$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_all <- data_plotting %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

# data_plotting_all$Date <- as.Date(paste(data_plotting_all$Year, data_plotting_all$Month, data_plotting_all$Day, sep = "-"))


#Plot data and change colors based on Zone:
viz_ch4_plot <- ggplot(data_plotting_all, aes(x = row_num, y = ch4_lin_flux.estimate, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "CH4", title = "CO2 Fluxes by Site") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_blank())

 viz_ch4_plot
 
 viz_co2_plot <- ggplot(data_plotting_all, aes(x = row_num, y = co2_lin_flux.estimate, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "CO2", title = "CO2 Fluxes by Site") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_blank())

 viz_co2_plot


```

## Boxplot summary by Date
```{r Overall Boxplot by Date, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_all <- data_plotting_all %>% 
    mutate(
    Zone_1 = case_when(
      Zone == "UP" ~ "Upland",
      Zone == "SWAMP" ~ "Swamp",
      Zone == "TR" ~ "Transition",
      Zone == "WC" ~ "Wetland",
      Zone == "SW" ~ "Surface Water",
      TRUE ~ Zone  # keep everything else unchanged
    )
  ) %>%
  mutate(Zone_1 = factor(Zone_1, levels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")))


##Summarize NPOC by Date, Site, and Zone
viz_ch4_boxplot <- ggplot(data_plotting_all, aes(x = as.character(Date), y = ch4_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 2) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CH4", title = "Summary Boxplot of CH4 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_ch4_boxplot


##Summarize TDN by Date, Site, and Zone
viz_co2_boxplot <- ggplot(data_plotting_all, aes(x = as.character(Date), y = co2_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 2) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CO2", title = "Summary Boxplot of CO2 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_co2_boxplot

```

## Boxplot summary by Site
```{r Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Summarize NPOC by Date, Site, and Zone
viz_ch4_summary_boxplot <- ggplot(data_plotting_all, aes(x = Zone_1, y = ch4_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 1) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CH4", title = "Summary Boxplot of CH4 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_ch4_summary_boxplot


##Summarize TDN by Date, Site, and Zone
viz_co2_summary_boxplot <- ggplot(data_plotting_all, aes(x = Zone_1, y = co2_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 1) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CO2", title = "Summary Boxplot of CO2 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_co2_summary_boxplot

```

##Calculate summary metrics
```{r Additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add to our collation codes a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, zone, depth for each year; also for each site zone, depth overall; and for each site and zone overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

##Metrics for each site and zone for all depths and years
summary <- data_plotting_all %>%
  group_by(Site, Zone_1) %>%
   reframe(
     min_ch4 = min(ch4_lin_flux.estimate, na.rm=TRUE),
     max_ch4 = max(ch4_lin_flux.estimate, na.rm=TRUE), 
     mean_ch4 = mean(ch4_lin_flux.estimate, na.rm=TRUE), 
     sd_ch4 = sd(ch4_lin_flux.estimate, na.rm=TRUE), 

     min_co2 = min(co2_lin_flux.estimate, na.rm=TRUE),
     max_co2 = max(co2_lin_flux.estimate, na.rm=TRUE), 
     mean_co2 = mean(co2_lin_flux.estimate, na.rm=TRUE), 
     sd_co2 = sd(co2_lin_flux.estimate, na.rm=TRUE))

```

## Summary Tables
```{r Summary table metrics, echo = FALSE, message = FALSE, dpi = 100}

library(kableExtra)

kable(
  summary[, c("Site", "Zone_1", "min_ch4",	"max_ch4",	"mean_ch4", "min_co2",	"max_co2",	"mean_co2")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Summary Statistics"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)


```
