---
title: "COMPASS_Synoptic_SoilFluxes_2025_CH4"
author: "Stephanie J. Wilson"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
---
## Goals of this code: 
1.	Reads in raw smart chamber
2.	If no metadatafile, write one out & fields will be: 
      Label, the IDs pulled from the label, 
      Plot *fixed so that if there are any issues in the label they are good here 
      Drop column – false as long as label looks right 
      Start time – would be initially populated by the start time in the OR blank 
      Observation length – initially populated by the obs length  OR blank 
      notes / comment 
3.	Code reads in the meta data 
    a.	Change drop column if fluxes need to be dropped
    b.	This is where the adjusted times will also be applied  
4.	Now you compute fluxes 
5.	Visualize and look for issues 
6.	If found, report to meta data file then go back to 3 and reanalyze 
7.	Apply algorithmic qaqc and generate flags 


## Set Up & Load Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(stringr)

```

## Pull in Smart Chamber Files from the Raw Data Folder
```{r, include=FALSE}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
zip_contents <- unzip("Raw Data.zip", list = TRUE)

# Keep only JSON files
json_files <- zip_contents$Name[grepl("json$", zip_contents$Name, ignore.case = TRUE)]

# Extract only the JSON files
unzip(
  "Raw Data.zip",
  files = json_files,
  exdir = tmp_dir
)

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "json$",
  full.names = TRUE,
  recursive = TRUE
)

head(file.list)

# Read in the data
alldat1 <- lapply(file.list, ffi_read_LIsmartchamber)

# Combine into one data.table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)


```

## Check if there is a meta-data file, if not, create a meta-data file with a unique Plot_ID 
```{r, echo=FALSE}

#create unique Plot ID for each measurement: 
    #this needs to be PLot ID so that later we ID plot from metadata and not the metadata row 
#detach("package:data.table", unload=TRUE)
alldat$month <- as.character(month(alldat$TIMESTAMP, label=TRUE, abbr=TRUE)) #will not run if you have data.table loaded 

alldat$year <- year(alldat$TIMESTAMP)

alldat <- alldat %>% 
  mutate(Plot_ID = str_c(pmin(month, RepNum), 
                      pmax(month, RepNum),  sep = '-'))
alldat <- alldat %>% 
  mutate(Plot_ID = str_c(pmin(Plot_ID, label), 
                     pmax(Plot_ID, label),  sep = '-'))  


# Path inside the zip
zip_file <- "Raw Data.zip"
csv_inside_zip <- "Raw Data/COMPASS_SmartChamber_Metadata_2025.csv"

# Make sure the destination folder exists
dir.create("Raw Data", showWarnings = FALSE)

# Extract only the metadata CSV if it doesn't already exist
if (!file.exists(csv_inside_zip)) {
  unzip(
    zip_file,
    files = csv_inside_zip,
    exdir = "."
  )
}

# Define the file path where you want to save the CSV file
file_path <- "Raw Data/COMPASS_SmartChamber_Metadata_2025.csv"

# Check if the CSV file exists
if (file.exists(file_path)) {
  # If it exists, read it back into R
  metadata <- read.csv(file_path)
  print("CSV file exists and has been read into the code.")
} else {
  # If it does not exist, create the CSV file
    #create a metadata file from the Plot IDs and pull timestart timestamps to put in the metadata and obs lengths 
  metadata <- alldat %>%
  group_by(Plot_ID) %>%
  summarise(
    Date = format(min(TIMESTAMP), "%Y-%m-%d"),
    Time_start = format(min(TIMESTAMP), "%H:%M:%S"),
    First_Timestamp = min(TIMESTAMP),  # Minimum (earliest) timestamp for each Plot_ID
    Last_Timestamp = max(TIMESTAMP),   # Maximum (latest) timestamp for each Plot_ID
    Total_Time = as.numeric(difftime(max(TIMESTAMP), min(TIMESTAMP), units = "secs")),  # Time difference in seconds
    ) %>%
  ungroup() %>%
  mutate(drop = FALSE) %>% # Add a 'drop' column with 'FALSE' for every row 
  #mutate(notes = " ")  # Add a 'drop' column with 'FALSE' for every row
  #mutate(Deadband = 5)
    mutate(Notes = "") #%>%
    #mutate(Plot = Plot_ID) #think need this because we need to make a plot column
  
  # Write df_unique to CSV
  write.csv(metadata, file = file_path, row.names = FALSE)
  
  print("CSV file does not exist. It has been created.")
}


```

## Mark Drop column in Metadata for Retaken Fluxes or other reasons to Drop 
```{r}

############## Check For Plot ID

##Sometimes there are other fluxes in the .json files that we don't want to calculate 
  #change drop column to TRUE if the plot doesn't include a site code in the Plot_ID 
metadata <- metadata %>%
  rename(Plot = Plot_ID) %>%
  mutate(
    drop = ifelse(
      !grepl("SWH|GCW|MSM|GWI|GCReW", Plot), 
      # Check if Plot_ID does not contain the specified strings
      TRUE,  # Set drop to TRUE if it doesn't contain any of the strings
      drop   # Keep the original value of drop if it contains one of the strings
    ) )

############## Check for retakes 

##Sometimes we have to retake a flux in the field due to issues or ebullition
##We need to change the drop column to reflect using only the latest flux
  #In order to we expect an "a" after the label, but sometimes this is not the case so we will fix that below and make it "a" 
metadata <- metadata %>%
    mutate(Plot = case_when(
    str_detect(Plot, "-retake") ~ str_replace(Plot, "-retake", "a"),
    str_detect(Plot, "-Retake") ~ str_replace(Plot, "-Retake", "a"),
    str_detect(Plot, "-b") ~ str_replace(Plot, "-b", "b"),
    TRUE ~ Plot  # Keeps the original value if none of the conditions match
  ))

#If the flux was taken more than once, flag the others so that we use the latest measurement 
# Extract the base plot (e.g., "GCW-TR-1" from "GCW-TR-1", "GCW-TR-1a", etc.)
metadata$base_plot <- str_extract(metadata$Plot,  "^[^-]+-[A-Za-z]+-[A-Za-z]+-[A-Za-z]+")

# Extract the replicate number (the number after the last hyphen)
metadata$replicate <- as.numeric(str_extract(metadata$Plot, "(?<=-)[0-9]+(?=[a-zA-Z]?$)"))

# Extract the suffix (if any, the letter after the replicate number)
metadata$suffix <- str_extract(metadata$Plot, "(?<=\\d)[a-zA-Z]$")

# If there's no suffix, set it to an empty string
metadata$suffix[is.na(metadata$suffix)] <- ""

# Sort by base plot, replicate number, and suffix (to keep highest suffix) - #####INSTEAD WE WANT TO CHANGE DROP TO TRUE HERE#######
metadata <- metadata %>%
  group_by(base_plot, replicate) %>%
  arrange(base_plot, replicate, desc(suffix), .by_group = TRUE) %>%
  mutate(drop = row_number() != 1) %>% #the entry with the latest suffix will keep false all others will be true
  #slice_head(n = 1) %>% #if you wanted to remove a the line rather than flag it 
  ungroup()

################# Check Observation lengths 

#if the observation length is over 5 mins then change drop column to true  
metadata <- metadata %>%
  mutate(drop = if_else(Total_Time > 300, TRUE, drop)) 

################# Remove columns with drop = TRUE
metadata <-  metadata %>%  
  filter(drop == "FALSE")


```

## Match up the metadata to the raw licor data   
```{r, include = FALSE}

#Remove any fluxes that have drop = TRUE in the metadata from the dataframe so we don't match them to raw data 
metadata <- metadata %>%
  filter(!drop)

#match the altered metadata to the concentration data with fluxfinder ffi_metadata_match
alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = alldat$TIMESTAMP,
  start_dates = metadata$Date,
  start_times = metadata$Time_start,
  obs_lengths = metadata$Total_Time) 
head(alldat)

#Add the Plot to the alldat dataframe 
alldat$Plot <- metadata$Plot[alldat$metadat_row] 


```

## Change the gas units from ppb to nmol  
```{r, include = FALSE}
#change the units
alldat$CH4_nmol <- ffi_ppb_to_nmol(alldat$ch4, 
                                volume = (alldat$TotalVolume / 1000000), # have to divide to get to m3
                                temp = alldat$chamber_t)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CH4_nmol_m2 <- alldat$CH4_nmol / 0.0318 #adjust for the area 


```

## Calculate fluxes
```{r, include = FALSE}

#need to create a new line that would say only calculate the flux if the drop column is false
    #before we add this we need to figure out how to drop except for the last flux taken 


#calculate fluxes for all other sites 
fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CH4_nmol_m2", 
                             dead_band = 5)
head(fluxes)

##Create a dataframe with averaged soil probe data per Plot and add to flux data
soildat <- alldat %>% 
    group_by(Plot) %>%
  dplyr::summarize(
     soilp_c = mean(soilp_c, na.rm=TRUE),
     soilp_m = mean(soilp_m, na.rm=TRUE),
     soilp_t = mean(soilp_t, na.rm=TRUE))

allflux <- left_join(fluxes, soildat, by = "Plot")

```


## Visualize fluxes and sort by r2 
```{r, echo = FALSE}

#take a look at the fluxes we calculated with R2 
ggplot(allflux, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + ylim(-50,5000) + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CH4 flux (nmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(allflux, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(allflux, lin_r.squared < .90 & abs(lin_flux.estimate) < 1, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(allflux, lin_r.squared < .90 & abs(lin_flux.estimate) > 5, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(allflux$lin_flux.estimate, 0.05) 
high <- quantile(allflux$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(allflux, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(allflux, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r}

tobechecked <- c(
)

colnames <- c("Plot")

#Loop to run five random plots subset above 
for(P in tobechecked){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), ch4)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }


#list of the ones that I changed and checked and looked good? 
fixedplots <- c("1-Sep-MSM-TR-1", "1-Sep-MSM-TR-2", "1-Sep-SWH-SWAMP-3", 
                "2-May-SWH-SWAMP-3a", "2-May-SWH-UP-3", "2-Sep-SWH-TR-6", 
                "1-Sep-GCW-TR-3", "1-Nov-SWH-SWAMP-1", "1-May-SWH-TR-4", 
                "1-May-SWH-TR-3", "1-Sep-SWH-TR-8", "1-May-SWH-TR-9", 
                "2-May-SWH-SWAMP-1a", "1-Mar-SWH-TR-3", "1-Mar-SWH-TR-5", 
                "1-May-SWH-SWAMP-4a", "1-Nov-SWH-SWAMP-3b", "1-Jul-SWH-SWAMP-4",
                "1-Jul-SWH-TR-10", "1-Jul-SWH-TR-8", "2-Jul-SWH-TR-6", 
                "2-May-SWH-TR-10", "2-May-SWH-TR-9", "2-Sep-SWH-TR-10a", 
                "2-Sep-SWH-TR-7", "2-Jul-SWH-TR-8", 
                "1-May-SWH-TR-5", "1-May-SWH-TR-6a", "1-Sep-GCW-TR-2", 
                "1-Sep-SWH-TR-5", "1-Sep-SWH-TR-6", "1-Sep-SWH-TR-7", 
                "1-Sep-SWH-TR-9", "2-Jul-SWH-TR-9", "2-Jul-SWH-UP-1", 
                "2-Jul-SWH-UP-3", "2-Jul-SWH-UP-4", "2-Jul-SWH-UP-5", 
                "2-Mar-SWH-TR-7a", "2-May-SWH-SWAMP-2", "2-May-SWH-TR-3", 
                "2-May-SWH-TR-4", "2-May-SWH-TR-7a", "2-May-SWH-TR-8a", 
                "2-Nov-SWH-SWAMP-1", "2-Nov-SWH-SWAMP-2", "2-Nov-SWH-SWAMP-3b", 
                "2-Nov-SWH-SWAMP-4a", "2-Nov-SWH-TR-9a", "2-Sep-GCW-TR-4", 
                "2-Sep-GCW-TR-5", "2-Sep-SWH-SWAMP-4", "2-Sep-SWH-TR-5", "2-Sep-SWH-TR-7", 
                "2-Mar-SWH-TR-5", "2-Sep-SWH-UP-4", "2-Sep-SWH-UP-1", "2-Sep-SWH-SWAMP-3", 
                "2-Sep-MSM-TR-1", "2-Sep-GCW-TR-2", "2-Nov-MSM-TR-5", "2-Nov-GWI-TR-1a", 
                "2-Nov-SWH-UP-4", "1-Jul-SWH-TR-5", "1-Jul-SWH-UP-2","1-Jun-GCW-TR-3",
                "1-Jun-GCW-TR-4", "1-Mar-SWH-TR-7a", "1-Mar-SWH-UP-1","1-May-SWH-SWAMP-3a", 
                "1-May-SWH-SWAMP-5", "1-May-SWH-TR-10", "1-May-SWH-TR-2", "1-May-SWH-UP-2", 
                "1-Nov-GCW-TR-1", "1-Nov-SWH-TR-5", "1-Nov-SWH-TR-6", "1-Nov-SWH-TR-9a", 
                "1-Nov-SWH-UP-2a", "1-Sep-GCW-TR-1", "1-Sep-GCW-TR-5", "1-Sep-MSM-TR-2", 
                "1-Sep-SWH-SWAMP-4", "1-Sep-SWH-SWAMP-5", "2-Jul-SWH-SWAMP-2", "2-Jul-SWH-TR-5", 
                "2-Jul-SWH-TR-8", "2-Jun-GCW-UP-1", "2-Mar-SWH-TR-1", "2-Mar-SWH-TR-2", 
                "2-Mar-SWH-TR-4", "2-Mar-SWH-UP-1", "2-May-SWH-SWAMP-3a", "2-May-SWH-UP-2")
  

```

\newpage

## Spot check a few fluxes with R2 > 0.9  
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), ch4)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################
```
\newpage

## Take a look at the  fluxes <0.9 and a flux estimate below 1
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot)

#Loop to run five random plots subset above 
for(P in low_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }


#print flag that says low r2, but low flux - checked 
allflux <- allflux %>%
  mutate(qaqc_flag = if_else(Plot %in% low_plots, "low r2, but also low flux - checked", NA_character_))


####################
```
\newpage

## Take a look at the bad fluxes <0.9 and have a high flux estimate (>1)
```{r, echo=FALSE}

print("See if we want to remove these fluxes or if we want to try and recalculate these ones")

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

#Loop to run five random plots subset above 
for(P in bad_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth()
  print(p1)
  }

####################

#add a flag for the fluxes that have ebullition and therefore not even a flux 
#allflux <- allflux %>%
    #mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% fluxes_weird, "low r2, high flux", qaqc_flag))


########## Ebullition 

#List the ones that have ebullition - assessed visually for now 
eb_flags <- c("2-Jul-GWI-TR-4")


#add a flag for the fluxes that have ebullition and therefore not even a flux 
allflux <- allflux %>%
    mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% eb_flags, "significant ebullition, flux compromised", qaqc_flag))


```

\newpage

## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

#Loop to run five random plots subset above 
for(P in tail_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Check all fluxes
```{#r, echo=FALSE}

##Put all fluxes together
fluxes_all <- as.data.frame(subset(allflux, select = Plot:TIMESTAMP_max))
all_plots <- unique(fluxes_all$Plot)

#Loop to run five random plots subset above
for(P in all_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth()
  print(p1)
  }

####################
```

## Flag fluxes that do not fit criteria, low R2 etc.  
```{r, results=FALSE}


########## Create a new data frame for output 

# previously removed rows that were present in fluxes_weird which have a low R2 and a flux higher than 1 (23 obs), but now they are flagged for ebullition
# flagged fluxes with low R squared and low flux
#allflux_filtered <- allflux[!(allflux$Plot %in% fluxes_weird$Plot), ]
allflux_filtered <- allflux


########## Poor fluxes

#List the ones that have odd/poor flux shapes - assessed visually for now 
poor_flags <- c("1-Sep-SWH-SWAMP-3", "1-Jul-SWH-TR-3", "1-Jul-SWH-TR-6", 
                "1-May-SWH-TR-7a", "1-Sep-SWH-TR-10a", "2-Sep-SWH-SWAMP-3", 
                "2-Mar-SWH-TR-3"
                )

#add a flag for the fluxes that are v-shaped and therefore not even a flux 
allflux_filtered <- allflux_filtered %>%
    mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% poor_flags, "poor flux, need to remove", qaqc_flag))


# 1-Sep-SWH-SWAMP-3 - messy flux - remove
# 1-Jul-SWH-TR-3 - messy flux - remove
# 1-Jul-SWH-TR-6 - messy flux - remove
# 1-May-SWH-TR-7a - messy flux - remove
# 1-Sep-SWH-TR-10a - messy flux - remove
# 2-Sep-SWH-SWAMP-3 - messy flux - remove
# 2-Mar-SWH-TR-3 - poor flux from ebullition and carryover - remove

# View the result
head(allflux_filtered)


```


## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(allflux_filtered, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(allflux_filtered, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(allflux_filtered, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere 
```{r, echo=FALSE}

#can remove base_plot column, the replicate column, and the suffix column
# allflux_filtered <- allflux_filtered[,-c(32:34)]

#create ID columns from the plots 
IDs <- data.frame(do.call('rbind', strsplit(as.character(allflux_filtered$Plot),'-',fixed=TRUE)))
colnames(IDs) <- c("RepNumber", "Month", "Site", "Zone", "Replicate")
head(IDs)

#combine these dataframes together 
allflux_filtered <- cbind(IDs, allflux_filtered)

#export as a csv to the folder that you want 
write.csv(allflux_filtered, file="Processed Data/COMPASS_Synoptic_CB_SmartChamber_Fluxes_Final_2025_CH4.csv", row.names = FALSE)           #Change file name


```


### ARCHIVE 

