---
title: "COMPASS Synoptic Surface Water Shakey GHG"
author: "2023 Samples"
date: "`r Sys.Date()`"
output:  
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information
```{r run information, include=TRUE}
cat("Run Information") #lets you know what section you're in
#set the run date & user name
sample_year <- "2023"
Sample_Run_Date <- "20240207"
# sample_month <- "May"
user <- "Max Wegner"

#record any notes about the run or anything other info here:
run_notes <- "All of CH4 checks were a bit high."

#identify the files you want to read in
#read in as a list to accommodate multiple runs in a month
GHG_files <- c("Raw Data/COMPASS_CBSYN_SW_GHG_2023.csv")


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Raw Data/COMPASS_Varian_QAQC_2023.csv"


# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/COMPASS_CBSYN_SW_GHG_2023_processed.csv"

#Gas in Water Calculations
# "TempC" column value in Celsius (equilibration temperature) and volume of gas (in liters) in syringe 
TempC = 25 #Equilibration Temperature (Celsius)

volume_g = 0.020 #volume of gas (in liters) in syringe when equilibrating
volume_w = 0.020 #volume of water (in liters) in syringe when equilibrating (should be equal to volume_g)



#Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

cat(run_notes)
```

##Set Up
```{r setup, include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  fuzzyjoin,
  ggplot2,
  ggpubr,
  stringr,
  here,
  data.table)

##Low & High Standard Curve Cutoffs
low_std_crv_cutoff <- 100
r2_cutoff <- 0.990

##Gas In Water Calculation
R <- 0.08206
CTK <- 273.15

#Metadata file path
Raw_Metadata = ("Raw Data/COMPASS_SynopticCB_PW_SampleLog_2023.csv")

##Detection Limits
#highest CH4 standard = 10000
#highest CO2 standard = 50000
adl_cutoff_ch4 <- 5000
adl_cutoff_co2 <- 50000
low_std_crv_area_cutoff = 71000
# bdl_cutoff <- 50

##Check Standard Concentration
#change this if check standard concentrations are different 
level_ch4 <- 5      #ppm
level_co2 <- 1000   #ppm
#Percentage of check standards that should be good
chks_flag = 60
blks_flag = 60



```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata
raw_metadata <- read.csv(Raw_Metadata)

##Create metadat sample ID to match raw data
metadata <- raw_metadata %>%
  filter(str_detect(Transect.Location, "Surface Water")) %>%
  mutate(Sample_ID = ifelse(Month < 10, paste0(Year, "0", Month, "_", Site, "_SW_", Lysimeter), 
                              paste0(Year, Month, "_", Site, "_SW_", Lysimeter))) #%>%

```

# Read in data files
```{r , include=FALSE}

#Read in Nov Data
raw_list <- map(GHG_files, read.csv)

# Combining Files
raw <- bind_rows(raw_list)

##Make sample ID
raw <- raw %>% 
  mutate(Month= match(Sample_Month, month.name)) %>% 
  mutate(Sample_ID = ifelse(Month < 10, paste0(Sample_Year, "0", Month, "_", Sample.ID), 
                              paste0(Sample_Year, Month, "_", Sample.ID))) %>% 
  mutate(Run_Date = Sample_Run_Date) #### All of these should be from the same run, even though it went into the next day. ####


```

# Assess standard curves
```{r , echo=FALSE}

#pull out methane standards
stds_ch4 <-  raw %>%  
  filter(str_detect(Sample_Type, "STD_CH4"))

#split into high curve and low curve for CH4
low_stds_ch4 <- stds_ch4  %>%  
  filter(!STD_Conc > low_std_crv_cutoff)

high_stds_ch4 <- stds_ch4  %>%  
  filter(STD_Conc >= low_std_crv_cutoff)



####High Curve for CH4####
CH4_High <- ggplot(high_stds_ch4, aes(STD_Conc, CH4_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High CH4 Std Curve")
CH4_High

# Convert 'run date' to a factor for grouping
high_stds_ch4_grouped <- high_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

split_data <- high_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
Slope_CH4_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "Slope_CH4_high",
                                        R2 = summary(lm(CH4_Area ~ STD_Conc, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ STD_Conc, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ STD_Conc, data = .))[1]))

Slope_CH4_raw <- bind_rows(Slope_CH4_raw_list)
  
Slope_CH4_High <- Slope_CH4_raw %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()


                              
####Low Curve for CH4####
CH4_Low <- ggplot(low_stds_ch4, aes(STD_Conc, CH4_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low CH4 Std Curve")
CH4_Low

low_CH4_lm <- lm(low_stds_ch4$CH4_Area ~ low_stds_ch4$STD_Conc)
low_cf <- coef(low_CH4_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_ch4_grouped <- low_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list_l <- map(split_data_l, ~ .x %>% mutate(Curve = "Slope_CH4_low",
                                        R2 = summary(lm(CH4_Area ~ STD_Conc, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ STD_Conc, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ STD_Conc, data = .))[1]))

Slope_CH4_raw_l <- bind_rows(Slope_CH4_raw_list_l)
  
Slope_CH4_Low <- Slope_CH4_raw_l %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()
 
 
 
####Pull out CO2 standards####
stds_co2 <- raw %>%  
  filter(str_detect(Sample_Type, "STD_CO2"))
 
#Curve for CO2
CO2_Curve <- ggplot(stds_co2, aes(STD_Conc, CO2_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "CO2 Std Curve")
CO2_Curve

CO2_lm <- lm(stds_co2$CO2_Area ~ stds_co2$STD_Conc)
CO2_cf <- coef(stds_co2)

 # Convert 'run date' to a factor for grouping
stds_co2_grouped <- stds_co2 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
split_data <- stds_co2_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  

#run regression and get coefficients by Run
Slope_CO2_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "Slope_CO2",
                                        R2 = summary(lm(CO2_Area ~ STD_Conc, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CO2_Area ~ STD_Conc, data = .))[2],
                                        Intercept = coef(lm(CO2_Area ~ STD_Conc, data = .))[1]))

Slope_CO2_raw <- bind_rows(Slope_CO2_raw_list)
  
Slope_CO2 <- Slope_CO2_raw %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()
 

Slopes <- rbind(Slope_CH4_Low, Slope_CH4_High, Slope_CO2)
Slopes

####Why are the intercepts not zero?####

#add new standard curve data to log
log <- read.csv(log_path)
head(log)
log <- log[, -c(1)]
log <- rbind(log, Slopes)

#remove possible duplicates from multiple QAQC attempts 
unique_log <- log %>% distinct()

Slopes_chk <- ggplot(log, aes(x = Run_Date, y = Slope, col = Curve)) +
  geom_point(size = 4) +
  # ylim(400, 800) +
  theme_bw()

Slopes_chk

write.csv(unique_log, log_path)

#check r2

ifelse(Slope_CH4_High$R2 < r2_cutoff, paste0("High_CH4_R2 is Bad","-", Slope_CH4_High$Run_Date), paste0("High_CH4_R2 is Good","-", Slope_CH4_High$Run_Date))
ifelse(Slope_CH4_Low$R2 < r2_cutoff, paste0("Low_CH4_R2 is Bad","-", Slope_CH4_Low$Run_Date), paste0("Low_CH4_R2 is Good","-", Slope_CH4_Low$Run_Date))
ifelse(Slope_CO2$R2 < r2_cutoff, paste0("CO2_R2 is Bad","-", Slope_CO2$Run_Date), paste0("CO2_R2 is Good","-", Slope_CO2$Run_Date))

High_CH4_R2_Flags <- Slope_CH4_High %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)
Low_CH4_R2_Flags <- Slope_CH4_Low %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)
CO2_R2_Flags <- Slope_CO2 %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)

Flags_df <- rbind(High_R2_Flags, Low_R2_Flags, CO2_R2_Flags)

```

## Now calculate the CH4 in ppm  
```{r Calculate Concentration, echo=FALSE}

##Calcaulte area of std curve cutoff
# low_std_crv_area_cutoff <- raw %>% 
#   filter(Sample.ID == "STD_100ppm_CH4") %>% 
#   pull(CH4_Area) %>%
#   mean(na.rm = TRUE)

#pull out Surface Water Samples
Samples <-  raw %>%  
  filter(str_detect(Type1, "SWGas")) %>%
  mutate(Curve = ifelse(CH4_Area >= low_std_crv_area_cutoff, "Slope_CH4_High", "Slope_CH4_Low"))
#head(Samples)

#Calculate CH4 concentrations in ppm
Samples$CH4_Conc_ppm <- ifelse(Samples$CH4_Area >= low_std_crv_area_cutoff, (Samples$CH4_Area-Slope_CH4_High$Intercept)/Slope_CH4_High$Slope, 
                              (Samples$CH4_Area-Slope_CH4_Low$Intercept)/Slope_CH4_Low$Slope )

#Calculate CO2 concentrations in ppm 
Samples$CO2_Conc_ppm <- ((Samples$CO2_Area-Slope_CO2$Intercept)/Slope_CO2$Slope) 

##########Add flags for any dilutions needed 
Samples$CH4_Flag <- ifelse(Samples$CH4_Conc_ppm > adl_cutoff_ch4, "Needs Dilution", "Within Range")
Samples$CO2_Flag <- ifelse(Samples$CO2_Conc_ppm > adl_cutoff_co2, "Needs Dilution", "Within Range")

#head(Samples)

```

# Aanlayze the Check Standards 
```{r Asses Check Standards, echo=FALSE}

chks <-  raw %>%
  filter(str_detect(Sample.ID, "CHKSTD"))

#Calculate CH4 concentrations in ppm
chks$CH4_Conc_ppm <- ifelse(chks$CH4_Area >= low_std_crv_area_cutoff, 
                      (chks$CH4_Area-Slope_CH4_High$Intercept)/Slope_CH4_High$Slope, 
                      (chks$CH4_Area-Slope_CH4_Low$Intercept)/Slope_CH4_Low$Slope )

#Calculate CO2 concentrations in ppm 
chks$CO2_Conc_ppm <- ((chks$CO2_Area-Slope_CO2$Intercept)/Slope_CO2$Slope)


#calculate percent difference of check standards
chks$Ch4_diff <- ((chks$CH4_Conc_ppm - level_ch4) / ((chks$CH4_Conc_ppm + level_ch4) /2)) * 100
chks$CH4_diff_flag <-  ifelse(chks$Ch4_diff < 10, 'YES', 'NO, rerun')

chks$CO2_diff <- ((chks$CO2_Conc_ppm - level_co2)/((chks$CO2_Conc_ppm + level_co2)/2)) * 100
chks$CO2_diff_flag <-  ifelse(chks$CO2_diff <10, 'YES', 'NO, rerun')


#calculate the percent of check standards that are within the range based on the flag
CH4_chks_percent <- (sum(chks$CH4_diff_flag == "YES") / nrow(chks)) * 100
CO2_chks_percent <- (sum(chks$CO2_diff_flag == "YES") / nrow(chks)) * 100

#report out if flags indicate need for rerun

if (CH4_chks_percent >= chks_flag) {
  message(">60% of CH4 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CH4 Check Standards are within range of expected concentration - REASSESS")
}

if (CO2_chks_percent >= chks_flag) {
  message(">60% of CO2 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CO2 Check Standards are within range of expected concentration - REASSESS")
}

#now plot the Ch4 concentrations and the CO2 concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

ch4_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CH4_Conc_ppm, fill = CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(padding = 0.25,preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "red", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title = "Check Stds: CH4") +
  theme(legend.position = "bottom") +  
  geom_hline(
    yintercept = level_ch4,
    linetype = "dashed",
    color = "black",
    linewidth = 1) +
  guides(fill = guide_legend(title = "% Difference <10%"))


co2_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CO2_Conc_ppm, fill = CO2_diff_flag)) +
  geom_bar(stat = 'identity', position = position_dodge2(padding = 0.25,preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "red", "YES" = "darkgreen")) +
  theme_classic() + 
  labs(x = " ", y = "Concentration (ppm)", title = "Check Stds: CO2") +
  theme(legend.position = "bottom") +  
  geom_hline(
    yintercept = level_co2,
    linetype = "dashed",
    color = "black",
    linewidth = 1) +
  guides(fill = guide_legend(title = "% Difference <10%"))

ggarrange(ch4_chk, co2_chk, nrow=1, ncol=2)

#Add Check Standard Flags

flag_chk <- ifelse(CH4_chks_percent >= chks_flag, "", "; CH4 Check Standards out Of Range")
flag_chk <- ifelse(CO2_chks_percent >= chks_flag, flag_chk, "; CO2 Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk))

```

# Analyze the Blanks 
```{r Analyze Blanks, echo=FALSE}
cat("Assess Blanks")

blks <- raw %>%
  filter(str_detect(Sample.ID, "Blank"))


#Pull out samples from df_all to calc quantile
samples_blanks <- raw %>%
  filter(Type1 == "SWGas")

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations
blk_flag_CH4 <- quantile(samples_blanks$CH4_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data
blks$CH4_diff_flag <-  ifelse(blks$CH4_Area <= blk_flag_CH4, 'YES', 'NO, rerun')

blk_flag_CO2 <- quantile(samples_blanks$CO2_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data
blks$CO2_diff_flag <-  ifelse(blks$CO2_Area <= blk_flag_CO2, 'YES', 'NO, rerun')


#calculate the percent of check standards that are within the range based on the flag
CH4_blks_percent <- (sum(blks$CH4_diff_flag == "YES") / nrow(blks)) * 100
CO2_blks_percent <- (sum(blks$CO2_diff_flag == "YES") / nrow(blks)) * 100

#report out if flags indicate need for rerun
if (CH4_blks_percent >= blks_flag) {
  message(">60% of CH4 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CH4 Blanks are lower 25% quartile of samples - REASSESS")
}

if (CO2_blks_percent >= blks_flag) {
  message(">60% of CO2 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CO2 Blanks are lower 25% quartile of samples - REASSESS")
}


#plotting the blanks compared to the lower 25% of conc (what the flag is)
CH4_blks_plot <-  ggplot(data = blks, aes(x = Run_Date, y = CH4_Area, fill = CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(padding = 0.25,preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CH4 Area", title = "CH4 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CH4),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))


CO2_blks_plot <-  ggplot(data = blks, aes(x = Run_Date, y = CO2_Area, fill = CO2_diff_flag)) +
  geom_bar(stat = 'identity', position = position_dodge2(padding = 0.25,preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CO2 Area", title = "CO2 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CO2),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "CO2 Blank Conc <25% Quartile Samples"))


ggarrange(CH4_blks_plot, CO2_blks_plot, nrow=1, ncol=2)


#find average of run blanks for flagging samples later
# Compute averages
blk_avg_CH4 <- mean(blks$CH4_Area, na.rm = TRUE)
blk_avg_CO2 <- mean(blks$CO2_Area, na.rm = TRUE)

# Create a data frame
blank_avgs <- data.frame(Test = c("CH4", "CO2"),
                         Blank_Mean_Area = c(blk_avg_CH4, blk_avg_CO2))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks <-  ifelse(CH4_blks_percent >= blks_flag,
                 "",
                 '; CH4 Blanks out of Range')
flag_blks <-  ifelse(CO2_blks_percent >= blks_flag,
                 flag_blks,
                 '; CO2 Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks))

```

# Dilution correct samples and Check Range
```{r, echo=FALSE}

####Are our exetainers for the SW GHG evacuated? ####

#Correct if volume other than 12mL of Sample was added
# Samples$CH4_Conc_ppm_dilcorr <- ((0.5 * Samples$CH4_Conc_ppm) * (12 + Samples$Sample_Volume)) / (Samples$Sample_Volume)


#multiply the concentration by the dilution factor
Samples$Dilution_Factor <- ifelse(is.na(Samples$Dilution_Factor), 1, Samples$Dilution_Factor)

Samples$CH4_Conc_ppm_dilcorr <- (Samples$CH4_Conc_ppm * Samples$Dilution_Factor)

Samples$CO2_Conc_ppm_dilcorr <- (Samples$CO2_Conc_ppm * Samples$Dilution_Factor)

#check results
#head(Samples)

#quick first look at the samples
ch4_samples_cor <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm_dilcorr, fill = CH4_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(padding = 0.25,preserve = "single"))+
  scale_fill_manual(values = c("green4", "darkred")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

co2_samples_cor <-  ggplot(data = Samples, aes(x = Sample_ID, y = CO2_Conc_ppm_dilcorr, fill = CO2_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(padding = 0.25,preserve = "single"))+
  scale_fill_manual(values = c("green4", "darkred")) +
  theme_classic() + labs(x = " ", y = "CO2 (ppm)", title = "CO2: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

ggarrange(ch4_samples_cor, co2_samples_cor, nrow=1, ncol=2)


```


# Calculate gas in water  
```{r, echo=FALSE}

# Convert ppm CH4 to total uM CH4 in the sample headspace
# use Ideal Gas Law:
#   hdsp umol CH4 = ppm CH4*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

####Is this correct?####
#Calculate CH4 uM from ppm 
Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w
Samples$CO2_Conc_uM <- (Samples$CO2_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w


#quick first look at the samples
ch4_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue",
    position = position_dodge2(padding = 0.25,preserve = "single"))+
  theme_classic() +
  labs(x = " ", y = "CH4 (uM)", title = "CH4 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

co2_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CO2_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue",
    position = position_dodge2(padding = 0.25,preserve = "single"))+
  theme_classic() +
  labs(x = " ", y = "CO2 (uM)", title = "CO2 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

ggarrange(ch4_samples, co2_samples, nrow=1, ncol=2)


```

# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

Samples <- Samples %>% mutate(Sample_ID = ifelse(Sample_ID %ilike% "SMARTX", Sample_ID, paste0("SMARTX_", Sample_ID)))

#check to see if all samples are present in the metadata
all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}


#merge metadata with sample run data
merged_data <- metadata %>%
  mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
  fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included?

#add flags

merged_2 <- merge(merged_data, Flags_df)
merged_data <-merged_2 %>% mutate(CH4_Flags = Flag)



df_all_clean <- merged_data

```

## Organize Data  
```{r, echo=FALSE}
#check results
#head(Samples)

#pull out what we need

df_all_clean_cols <- df_all_clean %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot,
    Depth_cm,
    CO2_Treatment,
    Temp_Treatment,
    CH4_Conc_ppm_dilcorr,
    CH4_Conc_uM,
    CH4_Flag,
    CH4_Flags,
    Run_Date
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = run_notes,# new column with notes about the run
    Analysis_rundate = Run_Date) 

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
    CH4_Conc_ppm = CH4_Conc_ppm_dilcorr,
    CH4_Conc_Flag = CH4_Flag,
    CH4_QAQC_Flags = CH4_Flags
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot,
    Depth_cm,
    CO2_Treatment,
    Temp_Treatment,
    CH4_Conc_ppm,
    CH4_Conc_uM,
    CH4_Conc_Flag,
    CH4_QAQC_Flags,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)

```


## Visualize Data
```{r Visualize Data C3, fig.height=10, echo=FALSE}
cat("Visualize Data")

### Ammonia 
CH4_forplot <- All_Clean_Data

#C3 plots
CH4_forplot_C3 <- CH4_forplot %>%
  filter(Nickname == "C3")

CH4_forplot_C3$Depth_cm <- 
  as.numeric(CH4_forplot_C3$Depth_cm)

#group the data for plotting
CH4_forplot_C3 <- CH4_forplot_C3 %>%
  group_by(Depth_cm) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup() %>%
  mutate(Treatment = ifelse(Temp_Treatment == 0, ifelse(CO2_Treatment == "Ambient", "0 C + aCO2", "0 C + eCO2"), ifelse(Temp_Treatment == 1.7, "1.7 C + aCO2", ifelse(Temp_Treatment == 3.4, "3.4 C + aCO2", ifelse(Temp_Treatment == 5.1, ifelse(CO2_Treatment == "Ambient", "5.1 C + aCO2", "5.1 C + eCO2"), NA)))))

#create plot
viz_CH4_plot_C3 <- ggplot(data = CH4_forplot_C3, aes(
  x = as.factor(Temp_Treatment),
  y = CH4_Conc_uM,
  fill = as.factor(Treatment),
  color = as.factor(Treatment)
)) +
  geom_bar(stat = "identity", position = position_dodge2(padding = 0.4, preserve = c("total", "single"), )) +
  facet_grid(as.numeric(Depth_cm) ~ ., scales = "free_x") +
  scale_fill_manual(
    values = c(
      "0 C + aCO2" = "blue",
      "1.7 C + aCO2" = "green",
      "3.4 C + aCO2" = "orange",
      "5.1 C + aCO2" = "red",
      "0 C + eCO2" = "white",
      "5.1 C + eCO2" = "white"
    ),
    name = "Treatment:",
    breaks = c("0 C + aCO2", "1.7 C + aCO2", "3.4 C + aCO2", "5.1 C + aCO2", "0 C + eCO2", "5.1 C + eCO2"),
    labels = c("0 C + aCO2", "1.7 C + aCO2", "3.4 C + aCO2", "5.1 C + aCO2", "0 C + eCO2", "5.1 C + eCO2")
  ) +
  scale_color_manual(
    values = c(
      "0 C + aCO2" = "blue",
      "1.7 C + aCO2" = "green",
      "3.4 C + aCO2" = "orange",
      "5.1 C + aCO2" = "red",
      "0 C + eCO2" = "blue",
      "5.1 C + eCO2" = "red"
    ),
    name = "Treatment:",
    breaks = c("0 C + aCO2", "1.7 C + aCO2", "3.4 C + aCO2", "5.1 C + aCO2", "0 C + eCO2", "5.1 C + eCO2"),
    labels = c("0 C + aCO2", "1.7 C + aCO2", "3.4 C + aCO2", "5.1 C + aCO2", "0 C + eCO2", "5.1 C + eCO2")
  ) +
  theme_classic() +
  labs(
    x = " ",
    y = "CH4 (uM)",
    title = "C3: Porewater CH4",
    fill = "CO2 Treatment",
    color = "+Temp. (C)"
  ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8),
      legend.key = element_rect("black")
    ) + 
  coord_flip()
print(viz_CH4_plot_C3)

```
\newpage
```{r Visualize Data C4, fig.height=10, echo=FALSE}
#C4 plots
CH4_forplot_C4 <- CH4_forplot %>%
  filter(Nickname == "C4")

CH4_forplot_C4$Depth_cm <-
  as.numeric(CH4_forplot_C4$Depth_cm)

#group the data for plotting
CH4_forplot_C4 <- CH4_forplot_C4 %>%
  group_by(Depth_cm) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_CH4_plot_C4 <- ggplot(data = CH4_forplot_C4, aes(
  x = as.factor(Temp_Treatment),
  y = CH4_Conc_uM,
  fill = as.factor(Temp_Treatment)
)) +
  geom_bar(stat = "identity", position = position_dodge2(padding = 0.25, preserve = "single"), ) +
  facet_grid(as.numeric(Depth_cm) ~ ., scales = "free_x") +
  facet_grid(Depth_cm ~ ., scales = "free_x") +
  scale_fill_manual(values = c(
    "0" = "blue",
    "1.7" = "green",
    "3.4" = "orange",
    "5.1" = "red"
  )) +
  
  theme_classic() +
  labs(x = " ",
       y = "CH4 (uM)",
       title = "C4: Porewater CH4",
       fill = "+Temp. (C)") +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    ),
    legend.key = element_rect("black")
    
  ) +
  coord_flip()
print(viz_CH4_plot_C4) 

```
\newpage
```{r Visualize Data by Rep C3,  fig.height=8.5, echo=FALSE}

#get replicate value and treatment info
CH4_forplot_C3 <- CH4_forplot_C3 %>%
  mutate(Replicate = str_sub(Plot, 2, 2)) %>%
  mutate(Treatment1 = ifelse(
    Temp_Treatment == 0,
    ifelse(CO2_Treatment == "Ambient", "+ 0 C", "eCO2 + 0 C"),
    ifelse(
      Temp_Treatment == 1.7,
      "+ 1.7 C",
      ifelse(
        Temp_Treatment == 3.4,
        "+ 3.4 C",
        ifelse(
          Temp_Treatment == 5.1,
          ifelse(CO2_Treatment == "Ambient", "+ 5.1 C", "eCO2 + 5.1 C"),
          NA
        )
      )
    )
  ))

#ensure points are in correct order
CH4_forplot_C3 <- CH4_forplot_C3[order(CH4_forplot_C3$Depth_cm), ]

#make plot
viz_CH4_plot_C3_2 <- ggplot(data = CH4_forplot_C3, 
                            aes(x = CH4_Conc_uM, y = Depth_cm, color = Replicate)) +
  geom_point(size = 3) +
  geom_path(size = 1) +
  scale_color_manual(values = c(
    "1" = "#FFD500",
    "2" = "#00D0FF",
    "3" = "#FF0075"
  )) +
  facet_grid(factor(
    Treatment1,
    levels = c("eCO2 + 0 C", "+ 0 C", "+ 1.7 C", "+ 3.4 C", "+ 5.1 C", "eCO2 + 5.1 C")
  ) ~ ., scales = "free_x") +
  labs(x = "CH4 (uM)", 
       y = "Depth (cm)", 
       title = "C3: Porewater CH4 by Replicate") +
  scale_y_reverse(breaks = c(10, 20, 40, 80, 120),
                  limits = c(125, 5)) +
  xlim(0,750) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_CH4_plot_C3_2

```
\newpage

```{r Visualize Data by Rep C4,  fig.height=5.7, echo=FALSE}

#get replicate value and treatment info
CH4_forplot_C4 <- CH4_forplot_C4 %>%
  mutate(Replicate = str_sub(Plot, 2, 2)) %>%
  mutate(Treatment1 = ifelse(
    Temp_Treatment == 0,
    ifelse(CO2_Treatment == "Ambient", "+ 0 C", "eCO2 + 0 C"),
    ifelse(
      Temp_Treatment == 1.7,
      "+ 1.7 C",
      ifelse(
        Temp_Treatment == 3.4,
        "+ 3.4 C",
        ifelse(
          Temp_Treatment == 5.1,
          ifelse(CO2_Treatment == "Ambient", "+ 5.1 C", "eCO2 + 5.1 C"),
          NA
        )
      )
    )
  ))

#make sure point are in correct order
CH4_forplot_C4 <- CH4_forplot_C4[order(CH4_forplot_C4$Depth_cm), ]

#make plot
viz_CH4_plot_C4_2 <- ggplot(data = CH4_forplot_C4, 
                            aes(x = CH4_Conc_uM, y = Depth_cm, color = Replicate)) +
  geom_point(size = 3) +
  geom_path(size = 1) +
  scale_color_manual(values = c(
    "4" = "#FFD500",
    "5" = "#00D0FF",
    "6" = "#FF0075"
  )) +
  facet_grid(factor(Treatment1, levels = c("+ 0 C", "+ 1.7 C", "+ 3.4 C", "+ 5.1 C")) ~
               ., scales = "free_x") +
  labs(x = "CH4 (uM)", 
       y = "Depth (cm)", 
       title = "C4: Porewater CH4 by Replicate") +
  scale_y_reverse(breaks = c(10, 20, 40, 80, 120),
                  limits = c(125, 5)) +
  xlim(0,750) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_CH4_plot_C4_2

```
##Export Data
```{r, echo=FALSE}
write.csv(All_Clean_Data, final_path)
#end
```
#end