---
title: "Microplate_FE_Code"
author: "Stephanie J. Wilson"
date: "2023-01-10"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## Information
```{r}
#######################################################################################
####### GENX2
####### Brackish
####### Data Analysis Code: Porewater iron 
####### MONTH: 
##############################################################################################


############################# Information ##################################
#Author: Alia N. Al-Haj
#Edited: 20241023

#Samples taken from porewater peepers
#Samples were filtered with 0.2 uM into HCl, kept on ice and then in fridge until analysis
#Field Protocol: 
#Samples Analyzed using GENX2 microplate method 
#Lab Protocol:
#
##Edited 20241104 by Melanie Giessner for
#high Fe curves typos
#graph read ability
#correcting the recombination of the high cv and okay data
```

## Code Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Date_Run = "20241231"
Run_by = "Melanie Giessner"
Project = "COMPASS"
Top_STD_Fe2_low = 60
Top_STD_Fe3_low = 60

Top_STD_Fe2_high = 500
Top_STD_Fe3_high = 100


library(ggplot2)
library(dplyr)
library(data.table)
library(matrixStats)
library(gridExtra)
library(ggpubr)
library(grid)
library(stringr)
library(raster)
library(plater)
#take out of sci notation
options(scipen = 999)
setwd("S:/GlobalChangeEco/Porewater_R scripts & Data/Data")
```

## R Markdown

#notes I don't have a system yet to compare old slopes to current slopes for that day 
#I don't have a system for eliminating wells if the sd is high - if you think one of your sample wells was off you would need to remove it manually... 

#Files Required: 
#csv of standards see template
#csv of samples see template


## Read in Standards and subset 
```{r}
stds <- read_plate(
  file = "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/Tidy Data/20241231_COMPASS_Fe_STD.csv",             # full path to the .csv file
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
)
str(stds)
head(stds)

#subset by FE_2 and FE_Total
FE_2_stds <- stds %>%  
  filter(str_detect(stds$IDs, "Fe_2"))  
head(FE_2_stds)

#subset by FE_2 and FE_Total
FE_Tot_stds <- stds %>%  
  filter(str_detect(IDs, "Fe_3"))  
head(FE_Tot_stds)

#subset Fe3 STD with in their linear range
FE_Tot_stds_abs1<- FE_Tot_stds
FE_Tot_stds_high <- FE_Tot_stds[!(FE_Tot_stds$Conc_uM > 100),]

#Subset the low standard curve from all 
FE_2_stds_low <- FE_2_stds[!(FE_2_stds$Conc_uM > 60),]
FE_Tot_stds_low <- FE_Tot_stds[!(FE_Tot_stds$Conc_uM > 60),]

#subset Fe2 abs2 to 100um 
FE_2_stds_high <-FE_2_stds
FE_2_stds_abs2<-FE_2_stds[!(FE_2_stds$Conc_uM > 100),]
FE_2_stds_diff<-FE_2_stds_abs2



```

## Calculate Reduction Efficiency 
```{r}

#pull out Fe_2 and Fe_3 standards at 60uM to check reduction efficiency

#subset by FE_2 and FE_Total
FE_2_60uM <- stds %>%  
  filter(str_detect(stds$IDs, "60 uM Fe_2 A"))  
head(FE_2_60uM)

#subset by FE_2 and FE_Total
FE_3_60 <- stds %>%  
  filter(str_detect(IDs, "60 uM Fe_3 A"))  
head(FE_3_60)


#create data frame with 0 rows and 3 columns
Fe_Red_eff <- data.frame(matrix(ncol = 0, nrow = 3))

Fe_Red_eff$ID <- seq.int(nrow(Fe_Red_eff))
Fe_Red_eff$FE2 <- (FE_2_60uM$Abs1)
Fe_Red_eff$FE3 <- FE_3_60$Abs2
Fe_Red_eff$Eff <- ((Fe_Red_eff$FE3)/Fe_Red_eff$FE2)*100
Fe_Red_eff$Eff1 <- ((Fe_Red_eff$FE3/0.8)/Fe_Red_eff$FE2)*100 #this one accounts for dilution of the reagents 
head(Fe_Red_eff)

#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 
redeffbar <- ggplot(data = Fe_Red_eff, aes(x = ID, y = Eff1)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("gray48")) + 
        theme_classic() + labs(x= "Sample ID", y="Reduction Efficiency (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=85, linetype="dashed", color = "black", size=1) + 
        geom_hline(yintercept=120, linetype="dashed", color = "black", size=1)

redeffbar

```

## Method Minimum Detection Limit (MDL) Calculation
I use the lowest standard to calculate MDL
How to calculate MDL https://www.epa.gov/sites/default/files/2016-12/documents/mdl-procedure_rev2_12-13-2016.pdf
Look at Table 1 on page 5
The Student's t-value used to calculate the method detection limit (MDL) is the one appropriate for a 99% confidence level and a standard deviation estimate with n-1 degrees of freedom

```{r}
#make a dataframe of just 0.6 uM Fe2 samples from standard dataframe
MDL_dat <- subset(FE_2_stds, Conc_uM == 0.6)

#calculate the standard deviation of abs 1
stddev <- sd(MDL_dat$Abs1)

#calculate n (number of 0.6 uM samples)
n <- nrow(MDL_dat)

#Identify the proper Student's T for n-1 from the table link above
StudentsT <- 2.528

#Multiply Standard deviation by Student's T
MDL <- stddev*StudentsT
MDL
```

#grab slope intercept from qaqc file
```{r}
#read in datafile with all the slopes
qlogO <- read.csv("S:/GlobalChangeEco/porewater_r scripts & Data/BGC_GCE_PorewaterAnalysis/Fe/Fe_STD_QAQC.csv")
qlog <- qlogO[,c(1:7)]
colnames(qlog) <- c("Date", "Project", "Curve", "R2", "Slope", "Intercept", "Top_STD")
head(qlog)


#Separate out by slope type for QAQC
qlog_Fe2_low <- subset(qlog, Curve == "FE_2_low")
qlog_Fe2_high <- subset(qlog, Curve == "FE_2_high")
qlog_FeTot_low <- subset(qlog, Curve == "FE_Tot_low")
qlog_FeTot_high <- subset(qlog, Curve == "FE_Tot_high")
qlog_Fe2_abs2 <- subset(qlog, Curve == "FE_2_abs2")
qlog_Fe3_abs1 <- subset(qlog, Curve == "FE_3_abs1")
qlog_Fe2_diff <- subset(qlog, Curve == "FE_2_diff")

# pull date run
qlog_Fe2_low_Date <- subset(qlog_Fe2_low, Date == Date_Run)
qlog_Fe2_high_Date <- subset(qlog_Fe2_high, Date == Date_Run)
qlog_FeTot_low_Date <- subset(qlog_FeTot_low, Date == Date_Run)
qlog_FeTot_high_Date <- subset(qlog_FeTot_high, Date == Date_Run)
qlog_Fe2_abs2_Date <- subset(qlog_Fe2_abs2, Date == Date_Run)
qlog_Fe3_abs1_Date <- subset(qlog_Fe3_abs1, Date == Date_Run)
qlog_Fe2_diff_Date <- subset(qlog_Fe2_diff, Date == Date_Run)

# get the slopes and intercepts from QAQC document
Fe2_low_intercept<- qlog_Fe2_low_Date$Intercept
Fe2_low_slope<- qlog_Fe2_low_Date$Slope
Fe2_high_intercept<- qlog_Fe2_high_Date$Intercept
Fe2_high_slope<- qlog_Fe2_high_Date$Slope
FeTot_low_intercept<- qlog_FeTot_low_Date$Intercept
FeTot_low_slope<- qlog_FeTot_low_Date$Slope
FeTot_high_intercept<- qlog_FeTot_high_Date$Intercept
FeTot_high_slope<- qlog_FeTot_high_Date$Slope
Fe2_abs2_intercept<- qlog_Fe2_abs2_Date$Intercept
Fe2_abs2_slope<- qlog_Fe2_abs2_Date$Slope
Fe3_abs1_intercept<- qlog_Fe3_abs1_Date$Intercept
Fe3_abs1_slope<- qlog_Fe3_abs1_Date$Slope
Fe2_diff_intercept<- qlog_Fe2_diff_Date$Intercept
Fe2_diff_slope<- qlog_Fe2_abs2_Date$Slope

```


## Read in Sample Data subset 
```{r}
setwd("S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe")

dat <- read_plate(
  file = "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/Tidy Data/20241231_COMPASS_Fe_plate8.csv",             # full path to the .csv file
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
)
str(dat)
head(dat)

## remove any lines that say stds in IDs
dat1 <- dat %>%  filter(!str_detect(IDs, "Fe_2"))
head(dat1)
dat1 <- dat1 %>%  filter(!str_detect(IDs, "Fe_3"))
head(dat1)
dat1 <- dat1 %>%  filter(!str_detect(IDs, "HCl"))
head(dat1)

#Now flag any Abs1 that are above the 60uM standard absorbance
#dat1$FE_2_Curve <- ifelse(dat1$Abs1 > mean(FE_2_60uM$Abs1), "High", "Within range")
#Now flag any Abs1 that are above the 500uM standard absorbance
FE_2_500uM<- FE_2_stds_high %>%  filter(str_detect(IDs, "500 uM Fe_2 A"))

dat1$FE_2_Curve <- ifelse(dat1$Abs1 > mean(FE_2_500uM$Abs1), "High", "Within range")

#and flag any that are below the MDL
dat1$FE_2_Curve <- ifelse(dat1$Abs1 < MDL, "Below MDL", dat1$FE_2_Curve)

head(dat1)

#Now flag any Abs2 that are above the 60uM standard absorbance
#dat1$FE_Tot_Curve <- ifelse(dat1$Abs2 > mean(FE_3_60$Abs2), "High", "Within range")
#Now flag any Abs2 that are above the 500uM standard absorbance
FE_2_100uM<- FE_2_stds %>%  filter(str_detect(IDs, "100 uM Fe_2 C"))

dat1$FE_Tot_Curve <- ifelse(dat1$Abs2 > mean(FE_2_100uM$Abs2), "High", "Within range")
#and flag any that are below the MDL
dat1$FE_Tot_Curve <- ifelse(dat1$Abs2 < MDL, "Below MDL", dat1$FE_Tot_Curve)


head(dat1)

#Make a list of samples to dilute and rerun
Fe_DiluteRerun <- subset(dat1, FE_2_Curve == "High"|FE_Tot_Curve == "High")

#Export the list of samples to rerun
write.csv(Fe_DiluteRerun, "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/QAQC'd Data/20241231_SamplesNeedDilution_plate8.csv")
```


## Check standards QAQC

```{r}
#Make sure check standards are not different from standard concentration
#subset Check Standards from the rest of the dataset
Fe2_std_Chk <- dat %>%  filter(str_detect(IDs, "Fe_2"))
head(Fe2_std_Chk)
FeTot_std_Chk <- dat %>%  filter(str_detect(IDs, "Fe_3"))
head(FeTot_std_Chk)


#Calculate Check standard Concentration
Fe2_std_Chk$Conc_uM <- (Fe2_std_Chk$Abs1-Fe2_low_intercept)/Fe2_low_slope
FeTot_std_Chk$Conc_uM <- (FeTot_std_Chk$Abs2-FeTot_low_intercept)/FeTot_low_slope


#Are the check standards significantly different from the standards?
#subset datasets for comparison
Fe2_std0.6 <- subset(FE_2_stds_low, Conc_uM == 0.6)
Fe2_Chkstd0.6 <- subset(Fe2_std_Chk, IDs %like% "0.6")
Fe2_std9 <- subset(FE_2_stds_low, Conc_uM == 9.0)
Fe2_Chkstd9 <- subset(Fe2_std_Chk, IDs %like% "9")
Fe2_std60 <- subset(FE_2_stds_low, Conc_uM == 60.0)
Fe2_Chkstd60 <- subset(Fe2_std_Chk, IDs %like% "60")

FeTot_std0.6 <- subset(FE_Tot_stds_low, Conc_uM == 0.6)
FeTot_Chkstd0.6 <- subset(FeTot_std_Chk, IDs %like% "0.6")
FeTot_std9 <- subset(FE_Tot_stds_low, Conc_uM == 9.0)
FeTot_Chkstd9 <- subset(FeTot_std_Chk, IDs %like% "9")
FeTot_std60 <- subset(FE_Tot_stds_low, Conc_uM == 60.0)
FeTot_Chkstd60 <- subset(FeTot_std_Chk, IDs %like% "60")


Fe2.t.test.std0.6 <- t.test(Fe2_std0.6$Abs1,Fe2_Chkstd0.6$Abs1,var.equal = T)
Fe2.t.test.std0.6

if(Fe2.t.test.std0.6$p.value > 0.05){
print("Continue")
} else {
print("Rerun plate")
}

Fe2.t.test.std9 <- t.test(Fe2_std9$Abs1,Fe2_Chkstd9$Abs1,var.equal = T)
Fe2.t.test.std9

if(Fe2.t.test.std9$p.value > 0.05){
print("Continue")
} else {
print("Rerun plate")
}

Fe2.t.test.std60 <- t.test(Fe2_std60$Abs1,Fe2_Chkstd60$Abs1,var.equal = T)
Fe2.t.test.std60

if(Fe2.t.test.std60$p.value > 0.05){
print("Continue")
} else {
print("Rerun plate")
}

FeTot.t.test.std0.6 <- t.test(FeTot_std0.6$Abs2,FeTot_Chkstd0.6$Abs2,var.equal = T)
FeTot.t.test.std0.6

if(FeTot.t.test.std0.6$p.value > 0.05){
print("Continue")
} else {
print("Rerun plate")
}

FeTot.t.test.std9 <- t.test(FeTot_std9$Abs2,FeTot_Chkstd9$Abs2,var.equal = T)
FeTot.t.test.std9

if(FeTot.t.test.std9$p.value > 0.05){
print("Continue")
} else {
print("Rerun plate")
}

FeTot.t.test.std60 <- t.test(FeTot_std60$Abs2,FeTot_Chkstd60$Abs2,var.equal = T)
FeTot.t.test.std60

if(FeTot.t.test.std60$p.value > 0.05){
print("Continue")
} else {
print("Rerun plate")
}
```

#No Matrix Checks
We do not use matrix checks for the Fe samples because our matrix checks are not made with Trace metal grade reagents. If we want to do this in the future, we need to use trace metal grade sodium chloride and sodium bicarbonate for making the matrix checks

## Subset Data and Calculate Concentrations 
```{r}
###Calculate concentrations of Fe(II)
dat1$FE_2_Conc <- ifelse(dat1$Abs1 > mean(Fe2_std60$Abs1), (dat1$Abs1-Fe2_high_intercept)/Fe2_high_slope, (dat1$Abs1-Fe2_low_intercept)/Fe2_low_slope)

head(dat1)
#Calculate the expected absorbance 2 for just Fe 2
#Expected_diff=m*conc+b
### abs2=abs1+diff
dat1$CalcAbs2<-dat1$Abs1+(dat1$FE_2_Conc*Fe2_diff_slope+Fe2_diff_intercept)

#subtract expected absorbance from actual absorbance this should leave you with the absorbance caused by the Fe3 reducing to Fe2
dat1$Abs2_diff<-dat1$Abs2-dat1$CalcAbs2

#using the slope from Fe 2 slope calculate Fe3 
dat1$FE_3_Conc <-ifelse(dat1$Abs2 > mean(FeTot_std60$Abs2), (dat1$Abs2_diff)/(FeTot_high_slope-Fe3_abs1_slope), (dat1$Abs2_diff)/(FeTot_low_slope-Fe3_abs1_slope))

#zeroAbs<-data.frame(IDs=dat$IDs,Conc_uM=dat$Conc_uM)
#zeroAbs$zero<- (dat1$CalcAbs2*Fe2_abs2_slope-dat1$Abs2*Fe3_abs1_slope)/(Fe2_abs2_slope*(Fe2_abs2_slope-Fe3_abs1_slope))
#zeroAbs<-zeroAbs %>%  filter(Conc_uM== 0.0)#Hcl with normal plate
#If samples have >45uM Fe3 recalculate Fe2 with Fe3 interference 
#have to do some tests with this
#dat1$FE_2_Conc_2<-ifelse(dat1$FE_3_Conc>45,((dat1$CalcAbs2*Fe2_abs2_slope-dat1$Abs2*Fe3_abs1_slope)/(Fe2_abs2_slope*(Fe2_abs2_slope-Fe3_abs1_slope)))-mean(zeroAbs$zero), dat1$FE_2_Conc)

head(dat1)
```


## Calculate dilutions and samples below the detection limit
```{r}

#Account for the dilution factor
dat1$FE_2_Conc_DilCorr <- dat1$FE_2_Conc*dat1$Dilutions

dat1$FE_3_Conc_DilCorr <- (dat1$FE_3_Conc*dat1$Dilutions)

head(dat1)

#Use ifelse to make any negative values equal to zero 
dat1$FE_2_Conc_Final <- ifelse(dat1$FE_2_Conc_DilCorr <0, 0, dat1$FE_2_Conc_DilCorr)

dat1$FE_3_Conc_Final <- ifelse(dat1$FE_3_Conc_DilCorr <0, 0, dat1$FE_3_Conc_DilCorr)

#Use ifelse to flag out of range values 
dat1$FE_2_Info <- ifelse(dat1$FE_2_Conc_DilCorr < MDL, 'bdl', ifelse( dat1$FE_2_Conc > Top_STD_Fe2_high,'high','within range'))

#Use ifelse to flag out of range values 
dat1$FE_3_Info <- ifelse(dat1$FE_3_Conc < MDL, 'bdl', ifelse( dat1$FE_3_Conc > Top_STD_Fe3_high,'high','within range'))

head(dat1)
```


## Calculate Averages across wells, std. dev, and cv.  
```{r}
head(dat1)

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat2 <- dat1 %>%
  group_by(IDs) %>%
  summarise(FE_2_mean = mean(FE_2_Conc_Final), FE_2_sd = sd(FE_2_Conc_Final), FE_2_cv = cv(FE_2_Conc_Final),
            FE_3_mean = mean(FE_3_Conc_Final), FE_3_sd = sd(FE_3_Conc_Final), FE_3_cv = cv(FE_3_Conc_Final),
            FE_2_flag = first(FE_2_Info), FE_3_flag = first(FE_3_Info), 
            Dilution = first(Dilutions))

head(dat2)

#plot data and sd's just to check and see what they look like - just a quick first look 
Fe2 <- ggplot(dat2, aes(x=IDs, y=FE_2_mean))+ 
  geom_point(size=4) +  theme_classic() + 
  labs(y="Fe(II) (uM)", x="Sample ID")+
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_2_mean-FE_2_sd,
                    ymax=FE_2_mean+FE_2_sd),width=0.3,position=position_dodge(.1))
Fe2

Fe3 <- ggplot(dat2, aes(x=IDs, y=FE_3_mean))+ 
  geom_point(size=4) +  theme_classic() + 
  labs(y="Fe(III) (uM)", x="Sample ID")+
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_3_mean-FE_3_sd,
                    ymax=FE_3_mean+FE_3_sd),width=0.3,position=position_dodge(.1))
Fe3

dat2$FE_2_cv_flag <- ifelse(dat2$FE_2_cv > 10, "High cv rerun", "ok")
dat2$FE_3_cv_flag <- ifelse(dat2$FE_3_cv > 10, "High cv rerun", "ok")

#Make a list of samples to rerun
HighCVSamples <- subset(dat2, FE_2_cv_flag == "High cv rerun"|FE_3_cv_flag == "High cv rerun")
HighCVSamplesFe2 <- subset(dat2, FE_2_cv_flag == "High cv rerun")
HighCVSamplesFe3 <- subset(dat2, FE_3_cv_flag == "High cv rerun")
```


## Remove bad reps
```{r}

#auto remove bad reps
dat_HCV <- subset(dat1, (IDs %in% HighCVSamples$IDs ))
head(dat_HCV)

#Columns 4 7 10
Column1= c("A04", "A07","A10", "B04", "B07", "B10" ,"C04" , "C07", "C10",
"D04", "D07", "D10", "E04","E07", "E10" ,"F04" , "F07" ,"F10", "G04","G07" ,"G10" ,"H04","H07","H10")

#Columns 5,8,11
Column2= c("A05", "A08","A11", "B05", "B08", "B11" ,"C05" , "C08", "C11",
"D05","D08", "D11", "E05","E08", "E11" ,"F05" , "F08", "F11" ,"G05","G08" ,"G11" ,"H05","H08", "H11")

#Columns 6,9,12
Column3= c("A06", "A09","A12", "B06", "B09" ,"B12" ,"C06" , "C09", "C12",
"D06","D09", "D12", "E06","E09", "E12" ,"F06" , "F09", "F12" ,"G06","G09" ,"G12" ,"H06","H09", "H12")


#delete Column one
dat_HCV1 <- subset(dat_HCV, !(Wells %in% Column1 ))
#delete Column two
dat_HCV2 <- subset(dat_HCV, !(Wells %in% Column2 ))
#delete Column three
dat_HCV3 <- subset(dat_HCV, !(Wells %in% Column3 ))

##Find CV for each Data set
#W/out column1
dat_HCV1 <- dat_HCV1 %>%
  group_by(IDs) %>%
  summarise(FE_2_mean = mean(FE_2_Conc_Final), FE_2_sd = sd(FE_2_Conc_Final), FE_2_cv = cv(FE_2_Conc_Final),
            FE_3_mean = mean(FE_3_Conc_Final), FE_3_sd = sd(FE_3_Conc_Final), FE_3_cv = cv(FE_3_Conc_Final),
            FE_2_flag = first(FE_2_Info),
            FE_3_flag = first(FE_3_Info),
            Dilution = first(Dilutions))
dat_HCV1$FE_2_cv_flag <- ifelse(dat_HCV1$FE_2_cv > 10, 'High CV', 'within range')
dat_HCV1$FE_3_cv_flag <- ifelse(dat_HCV1$FE_3_cv > 10, 'High CV', 'within range')
#W/out column2
dat_HCV2 <- dat_HCV2 %>%
  group_by(IDs) %>%
  summarise(FE_2_mean = mean(FE_2_Conc_Final), FE_2_sd = sd(FE_2_Conc_Final), FE_2_cv = cv(FE_2_Conc_Final),
            FE_3_mean = mean(FE_3_Conc_Final), FE_3_sd = sd(FE_3_Conc_Final), FE_3_cv = cv(FE_3_Conc_Final),
            FE_2_flag = first(FE_2_Info),
            FE_3_flag = first(FE_3_Info),
            Dilution = first(Dilutions))
dat_HCV2$FE_2_cv_flag <- ifelse(dat_HCV2$FE_2_cv > 10, 'High CV', 'within range')
dat_HCV2$FE_3_cv_flag <- ifelse(dat_HCV2$FE_3_cv > 10, 'High CV', 'within range')
#W/out column3
dat_HCV3 <- dat_HCV3 %>%
  group_by(IDs) %>%
  summarise(FE_2_mean = mean(FE_2_Conc_Final), FE_2_sd = sd(FE_2_Conc_Final), FE_2_cv = cv(FE_2_Conc_Final),
            FE_3_mean = mean(FE_3_Conc_Final), FE_3_sd = sd(FE_3_Conc_Final), FE_3_cv = cv(FE_3_Conc_Final),
            FE_2_flag = first(FE_2_Info),
            FE_3_flag = first(FE_3_Info),
            Dilution = first(Dilutions))
dat_HCV3$FE_2_cv_flag <- ifelse(dat_HCV3$FE_2_cv > 10, 'High CV', 'within range')
dat_HCV3$FE_3_cv_flag <- ifelse(dat_HCV3$FE_3_cv > 10, 'High CV', 'within range')

#find lowest CVs
dat_HCV1_Fe2 <- subset(dat_HCV1, dat_HCV1$FE_2_cv < dat_HCV2$FE_2_cv & dat_HCV1$FE_2_cv < dat_HCV3$FE_2_cv)
head(dat_HCV1_Fe2)
dat_HCV1_Fe3 <- subset(dat_HCV1, dat_HCV1$FE_3_cv < dat_HCV2$FE_3_cv & dat_HCV1$FE_3_cv < dat_HCV3$FE_3_cv)
head(dat_HCV1_Fe3)

dat_HCV2_Fe2 <- subset(dat_HCV2, dat_HCV2$FE_2_cv < dat_HCV1$FE_2_cv & dat_HCV2$FE_2_cv < dat_HCV3$FE_2_cv)
head(dat_HCV2_Fe2)
dat_HCV2_Fe3 <- subset(dat_HCV2, dat_HCV2$FE_3_cv < dat_HCV1$FE_3_cv & dat_HCV2$FE_3_cv < dat_HCV3$FE_3_cv)
head(dat_HCV2_Fe3)

dat_HCV3_Fe2 <- subset(dat_HCV3, dat_HCV3$FE_2_cv < dat_HCV1$FE_2_cv & dat_HCV3$FE_2_cv < dat_HCV2$FE_2_cv)
head(dat_HCV3_Fe2)
dat_HCV3_Fe3 <- subset(dat_HCV3, dat_HCV3$FE_3_cv < dat_HCV1$FE_3_cv & dat_HCV3$FE_3_cv < dat_HCV2$FE_3_cv)
head(dat_HCV3_Fe3)

#recombine data frames
HCV_Fe2_recalc <- rbind(dat_HCV1_Fe2, dat_HCV2_Fe2, dat_HCV3_Fe2)
HCV_Fe3_recalc <- rbind(dat_HCV1_Fe3, dat_HCV2_Fe3, dat_HCV3_Fe3)

#order the dataframes and make sure sample lists are the same size
HighCVSamples1<-HighCVSamples[order(HighCVSamples$IDs),]
HighCVSamplesFe2_1 <-subset(HighCVSamples1, (!IDs %in% HCV_Fe2_recalc$IDs ))
HighCVSamplesFe2_1 <- rbind(HCV_Fe2_recalc, HighCVSamplesFe2_1 )
HighCVSamplesFe2_1 <-HighCVSamplesFe2_1[order(HighCVSamplesFe2_1$IDs),]

HighCVSamples1<-HighCVSamples[order(HighCVSamples$IDs),]
HighCVSamplesFe3_1 <-subset(HighCVSamples1, (!IDs %in% HCV_Fe3_recalc$IDs ))
HighCVSamplesFe3_1 <- rbind(HighCVSamplesFe3_1, HCV_Fe3_recalc)
HighCVSamplesFe3_1 <-HighCVSamplesFe3_1[order(HighCVSamplesFe3_1$IDs),]


#correct mean, sd, cv for combined Fe2 and Fe3
HighCVSamples1$FE_2_mean <- ifelse(HighCVSamples$IDs %in% HighCVSamplesFe2$IDs, HighCVSamplesFe2_1$FE_2_mean,HighCVSamples$FE_2_mean)
HighCVSamples1$FE_2_sd <- ifelse(HighCVSamples$IDs %in% HighCVSamplesFe2$IDs, HighCVSamplesFe2_1$FE_2_sd, HighCVSamples$FE_2_sd )
HighCVSamples1$FE_2_cv <- ifelse(HighCVSamples$IDs %in% HighCVSamplesFe2$IDs, HighCVSamplesFe2_1$FE_2_cv,HighCVSamples$FE_2_cv )

HighCVSamples1$FE_3_mean <- ifelse(HighCVSamples$IDs %in% HighCVSamplesFe3$IDs, HighCVSamplesFe3_1$FE_3_mean,HighCVSamples1$FE_3_mean )
HighCVSamples1$FE_3_sd <- ifelse(HighCVSamples$IDs %in% HighCVSamplesFe3$IDs, HighCVSamplesFe3_1$FE_3_sd,HighCVSamples1$FE_3_sd )
HighCVSamples1$FE_3_cv <- ifelse(HighCVSamples$IDs %in% HighCVSamplesFe3$IDs, HighCVSamplesFe3_1$FE_3_cv,HighCVSamples1$FE_3_cv )

#rerun CV flag
HighCVSamples1$FE_2_cv_flag <- ifelse(HighCVSamples1$FE_2_cv > 10, "High cv rerun", "ok")
HighCVSamples1$FE_3_cv_flag <- ifelse(HighCVSamples1$FE_3_cv > 10, "High cv rerun", "ok")

#recombine corrected means with original data
dat2 <- subset(dat2, (!IDs %in% HighCVSamples1$IDs ))
dat2 <- rbind(HighCVSamples1, dat2)
head(dat2)

Fe2 <- ggplot(dat2, aes(x=IDs, y=FE_2_mean))+ 
  geom_point(size=4) +  theme_classic() + 
  labs(y="Fe(II) (uM)", x="Sample ID")+
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_2_mean-FE_2_sd,
                    ymax=FE_2_mean+FE_2_sd),width=0.3,position=position_dodge(.1))
Fe2

Fe3 <- ggplot(dat2, aes(x=IDs, y=FE_3_mean))+ 
  geom_point(size=4) +  theme_classic() + 
  labs(y="Fe(III) (uM)", x="Sample ID")+
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_3_mean-FE_3_sd,
                    ymax=FE_3_mean+FE_3_sd),width=0.3,position=position_dodge(.1))
Fe3

#Make a list of samples to rerun
HighCVSamples2 <- subset(dat2, FE_2_cv_flag == "High cv rerun"|FE_3_cv_flag == "High cv rerun")

#If samples with cv >10 rerun those samples 
#Export the list of samples to rerun
write.csv(HighCVSamples2, "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/QAQC'd Data/20241231_HighCVrerun_plate8.csv")
```


## Check the dups for QAQC 
```{r}

#Show me the data that we have from the calculations 
head(dat2)

#pull out any rows that have "d" in the SampleID column
dups <- dat2 %>%  
  filter(str_detect(IDs, "dup")) 
head(dups)

#remove these from dat2
dat3 <- dat2 %>%  
  filter(!str_detect(IDs, "dup")) %>%  
  filter(!str_detect(IDs, "spk")) 
head(dat3)

#remove the d from these IDs so we will have duplicates 
dups$IDs<-gsub(" dup","",as.character(dups$IDs))
dups <- dups[ ,-c(3,4,6:12)]
head(dups)
colnames(dups) <- c('IDs', 'FE_2_mean_dup', "FE_3_mean_dup")


#put it back together with the old data set and look for duplicates 
QAdups <- merge(dat3, dups)
head(QAdups)

QAdups$FE_2_dups_chk <- ((abs(QAdups$FE_2_mean-QAdups$FE_2_mean_dup))/((QAdups$FE_2_mean+QAdups$FE_2_mean_dup)/2))*100
QAdups$FE_2_dups_flag <-  ifelse(QAdups$FE_2_dups_chk <10, 'YES', 'NO, rerun')

QAdups$FE_3_dups_chk <- ((abs(QAdups$FE_3_mean-QAdups$FE_3_mean_dup))/((QAdups$FE_3_mean+QAdups$FE_3_mean_dup)/2))*100
QAdups$FE_3_dups_flag <-  ifelse(QAdups$FE_3_dups_chk <10, 'YES', 'NO, rerun')
head(QAdups)

#plot dups output as a bar graph to easily check - want any over 10% to be red need to work on this 
dupsbar_Fe2 <- ggplot(data = QAdups, aes(x = IDs, y = FE_2_dups_chk, fill=FE_2_dups_chk)) +
       geom_bar(stat = 'identity') + 
        scale_fill_gradient2(low='darkgreen', mid='darkgreen', high='darkgrey', space='Lab') + 
        theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
        theme(legend.position="none") +  geom_hline(yintercept=10, linetype="dashed", 
                color = "black", size=1)

dupsbar_Fe2

dupsbar_Fe3 <- ggplot(data = QAdups, aes(x = IDs, y = FE_3_dups_chk, fill=FE_3_dups_chk)) +
       geom_bar(stat = 'identity') + 
        scale_fill_gradient2(low='darkgreen', mid='darkgreen', high='darkgrey', space='Lab') + 
        theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
        theme(legend.position="none") +  geom_hline(yintercept=10, linetype="dashed", 
                color = "black", size=1)

dupsbar_Fe3


#check for any no's that would warrant reruns
Baddups <- subset(QAdups, FE_2_dups_flag == "NO, rerun"|FE_3_dups_flag == "NO, rerun")

#Export the list of samples to rerun
write.csv(Baddups, "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/QAQC'd Data/20241231_Baddups_plate8.csv")
```


## Check the spks for QAQC 
```{r}

#pull out any rows that have "spk" in the SampleID column
spks <- dat2 %>%  
  filter(str_detect(IDs, "spk")) 
head(spks)

#remove the spk from these IDs so we will have duplicates 
spks$IDs<-gsub(" spk","",as.character(spks$IDs))
spks <- spks[ ,-c(3:12)]
head(spks)
colnames(spks) <- c('IDs', 'FE_2_mean_spk')


#put it back together with the old data set and look for duplicates 
QAspks <- merge(dat3, spks)
head(QAspks)

#now we need to calculate the spike concentration and calculate the spike recovery 
#spike for these samples was 55uL of the 60uM standard
QAspks$FE_2_spk_Conc <- (60*(55/1000000))                        # this would be in umoles of FE in the spk 
QAspks$FE_Total_unspkd <- (QAspks$FE_2_mean/QAspks$Dilution)*(160/1000000) #gives us the total fe in the sample in umoles
QAspks$FE_Total_spkd <-    (QAspks$FE_2_mean_spk/QAspks$Dilution)*((160+55)/1000000)        ##total fe in spiked sample in umoles 
QAspks$FE_expctd_spkd <-  (QAspks$FE_Total_unspkd + QAspks$FE_2_spk_Conc)
QAspks$spk_recovery <-    (QAspks$FE_Total_spkd/QAspks$FE_expctd_spkd)*100
QAspks$FE_spks_flag <-  ifelse(QAspks$spk_recovery >80 & QAspks$spk_recovery < 120, 'YES', 'NO, rerun') 

head(QAspks)

#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 
spksbar <- ggplot(data = QAspks, aes(x = IDs, y = spk_recovery)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("gray48")) + 
        theme_classic() + labs(x= "Sample ID", y="Spike Recovery (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=80, linetype="dashed", color = "black", size=1) + 
        geom_hline(yintercept=120, linetype="dashed", color = "black", size=1)

spksbar

#check for any no's that would warrant reruns!
Badspks <- subset(QAspks, FE_spks_flag == "No, rerun")

#Export the list of samples to rerun
write.csv(Badspks, "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/QAQC'd Data/20241231_Badspks_plate8.csv")
```



## Export full data then just final data 
```{r}


#Read out the data with all flags 
head(dat3)
write.csv(dat3, file="S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/QAQC'd Data/COMAPSS_Fe_20241231_plate8_Flagdat.csv")

#Now take out everything but the means and remove any samples that did not pass QAQC 
head(dat3)
dat4 <- subset(dat3, FE_2_cv_flag == "ok" & FE_3_cv_flag == "ok")
dat4 <- dat4[ ,-c(3,4,6:12)]
head(dat4)
write.csv(dat4, file="S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/QAQC'd Data/COMPASS_Fe_20241231_plate8_shortdata.csv")


```

### END



