---
title: "Microplate_FE_Code"
author: "Stephanie J. Wilson"
date: "2023-01-10"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## Information
```{r}
#######################################################################################
####### GENX2
####### Brackish
####### Data Analysis Code: Porewater iron 
####### MONTH: 
##############################################################################################


############################# Information ##################################
#Author: Alia N. Al-Haj
#Edited: 20241023

#Samples taken from porewater peepers
#Samples were filtered with 0.2 uM into HCl, kept on ice and then in fridge until analysis
#Field Protocol: 
#Samples Analyzed using GENX2 microplate method 
#Lab Protocol:
#
##Edited 20241104 by Melanie Giessner for
#high Fe curves typos
#graph read ability
#correcting the recombination of the high cv and okay data
```

## Code Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Date_Run = "20241230"
Run_by = "Melanie Giessner"
Project = "COMPASS"
Top_STD_Fe2_low = 60
Top_STD_Fe3_low = 60

Top_STD_Fe2_high = 500
Top_STD_Fe3_high = 100


library(ggplot2)
library(dplyr)
library(data.table)
library(matrixStats)
library(gridExtra)
library(ggpubr)
library(grid)
library(stringr)
library(raster)
library(plater)

setwd("S:/GlobalChangeEco/Porewater_R scripts & Data/Data")
```

## R Markdown

#notes I don't have a system yet to compare old slopes to current slopes for that day 
#I don't have a system for eliminating wells if the sd is high - if you think one of your sample wells was off you would need to remove it manually... 

#Files Required: 
#csv of standards see template
#csv of samples see template


## Read in Standards and subset 
```{r}

stds <- read_plate(
  file = "S:/GlobalChangeEco/Porewater_R scripts & Data/Data/Fe/COMPASS/Synoptic/Tidy Data/20241230_COMPASS_Fe_STD2.csv",             # full path to the .csv file
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
)
str(stds)
head(stds)

#subset by FE_2 and FE_Total
FE_2_stds <- stds %>%  
  filter(str_detect(stds$IDs, "Fe_2"))  
head(FE_2_stds)

#subset by FE_2 and FE_Total
FE_Tot_stds <- stds %>%  
  filter(str_detect(IDs, "Fe_3"))  
head(FE_Tot_stds)

#subset Fe3 STD with in their linear range
FE_Tot_stds_abs1<- FE_Tot_stds
FE_Tot_stds_high <- FE_Tot_stds[!(FE_Tot_stds$Conc_uM > 100),]

#Subset the low standard curve from all 
FE_2_stds_low <- FE_2_stds[!(FE_2_stds$Conc_uM > 60),]
FE_Tot_stds_low <- FE_Tot_stds[!(FE_Tot_stds$Conc_uM > 60),]

#subset Fe2 abs2 to 100um 
FE_2_stds_high <-FE_2_stds
FE_2_stds_abs2<-FE_2_stds[!(FE_2_stds$Conc_uM > 100),]
FE_2_stds_diff<-FE_2_stds_abs2

```

## Plot standards 
```{r}
#Plot each of these and calculate the slope, intercept, and R2 
library(ggplot2)

#Low curve for Fe(II)
FE_2_low <- ggplot(FE_2_stds_low, aes(Conc_uM, Abs1)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Low Fe(II) Absorbance 1 Curve")
FE_2_low

FE_2_low_lm <- lm(FE_2_stds_low$Abs1 ~ FE_2_stds_low$Conc_uM)
summary(FE_2_low_lm)
cf <- coef(FE_2_low_lm)

#create data frame with 1 rows and 0 columns
Slopes1 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes1$Date <- Date_Run
Slopes1$Project <- Project
Slopes1$Curve <- "FE_2_low"
Slopes1$R2 <- summary(FE_2_low_lm)$adj.r.squared
Slopes1$Slope <- cf[2]
Slopes1$Intercept <- cf[1]
Slopes1$Top_STD <- Top_STD_Fe2_low
head(Slopes1)


#High curve for Fe2
 FE_2_high <- ggplot(FE_2_stds, aes(Conc_uM, Abs1)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("High Fe(II) Absorbance 1 Curve")
 FE_2_high
 
 FE_2_high_lm <- lm(FE_2_stds$Abs1 ~ FE_2_stds$Conc_uM)
 summary(FE_2_high_lm)
 cf <- coef(FE_2_high_lm)
 
#create data frame with 1 rows and 0 columns
 Slopes2 <- data.frame(matrix(ncol = 0, nrow = 1))
 Slopes2$Date <- Date_Run
 Slopes2$Project <- Project
 Slopes2$Curve <- "FE_2_high"
 Slopes2$R2 <- summary(FE_2_high_lm)$adj.r.squared
 Slopes2$Slope <- cf[2]
 Slopes2$Intercept <- cf[1]
 Slopes2$Top_STD <- Top_STD_Fe2_high
 head(Slopes2)
 

#Low curve for Fe(Total)
FE_Tot_low <- ggplot(FE_Tot_stds_low, aes(Conc_uM, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Low Fe(III) Absorbance 2 Curve")
FE_Tot_low

FE_Tot_low_lm <- lm(FE_Tot_stds_low$Abs2 ~ FE_Tot_stds_low$Conc_uM)
summary(FE_Tot_low_lm)
cf <- coef(FE_Tot_low_lm)

#create data frame with 1 rows and 0 columns
Slopes3 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes3$Date <- Date_Run
Slopes3$Project <- Project
Slopes3$Curve <- "FE_Tot_low"
Slopes3$R2 <- summary(FE_Tot_low_lm)$adj.r.squared
Slopes3$Slope <- cf[2]
Slopes3$Intercept <- cf[1]
Slopes3$Top_STD <- Top_STD_Fe3_low
head(Slopes3)


 #High curve for Fe(III)
 FE_Tot_high <- ggplot(FE_Tot_stds_high, aes(Conc_uM, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("High Fe(III) Absorbance 2 Curve")
FE_Tot_high
 
FE_Tot_high_lm <- lm(FE_Tot_stds_high$Abs2 ~ FE_Tot_stds_high$Conc_uM)
summary(FE_Tot_high_lm)
cf <- coef(FE_Tot_high_lm)
 
 #create data frame with 1 rows and 0 columns
Slopes4 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes4$Date <- Date_Run
Slopes4$Project <- Project
Slopes4$Curve <- "FE_Tot_high"
Slopes4$R2 <- summary(FE_Tot_high_lm)$adj.r.squared
Slopes4$Slope <- cf[2]
Slopes4$Intercept <- cf[1]
Slopes4$Top_STD <- Top_STD_Fe3_high
head(Slopes4)

#curve for Fe(II) Abs2
FE_2_abs2 <- ggplot(FE_2_stds_abs2, aes(Conc_uM, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(II) Absorbance 2 Curve")
FE_2_abs2

FE_2_stds_abs2_lm <- lm(FE_2_stds_abs2$Abs2 ~ FE_2_stds_abs2$Conc_uM)
summary(FE_2_stds_abs2_lm)
cf <- coef(FE_2_stds_abs2_lm)

#create data frame with 1 rows and 0 columns
Slopes5 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes5$Date <- Date_Run
Slopes5$Project <- Project
Slopes5$Curve <- "FE_2_abs2"
Slopes5$R2 <- summary(FE_2_stds_abs2_lm)$adj.r.squared
Slopes5$Slope <- cf[2]
Slopes5$Intercept <- cf[1]
Slopes5$Top_STD <- 100
head(Slopes5)

#curve for Fe(III) Abs1
FE_3_abs1 <- ggplot(FE_Tot_stds_abs1, aes(Conc_uM,Abs1 )) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(III) Absorbance 1 Curve")
FE_3_abs1

FE_3_abs1_lm <- lm(FE_Tot_stds_abs1$Abs1 ~ FE_Tot_stds_abs1$Conc_uM )
summary(FE_3_abs1_lm)
cf <- coef(FE_3_abs1_lm)

#create data frame with 1 rows and 0 columns
Slopes6 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes6$Date <- Date_Run
Slopes6$Project <- Project
Slopes6$Curve <- "FE_3_abs1"
Slopes6$R2 <- summary(FE_3_abs1_lm)$adj.r.squared
Slopes6$Slope <- cf[2]
Slopes6$Intercept <- cf[1]
Slopes6$Top_STD <- 500
head(Slopes6)

#This slope should be small e-6


#curve for Fe(II) Abs2-Abs1

FE_2_stds_diff$FE_2_Diff<-FE_2_stds_diff$Abs2-FE_2_stds_diff$Abs1

FE_2 <- ggplot(FE_2_stds_diff, aes(Conc_uM,FE_2_Diff)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(II) Absorbance Difference Curve")
FE_2

FE_2_diff_lm <- lm(FE_2_stds_diff$FE_2_Diff ~ FE_2_stds_diff$Conc_uM)
summary(FE_2_diff_lm)
cf <- coef(FE_2_diff_lm)

#create data frame with 1 rows and 0 columns
Slopes7 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes7$Date <- Date_Run
Slopes7$Project <- Project
Slopes7$Curve <- "FE_2_diff"
Slopes7$R2 <- summary(FE_2_diff_lm)$adj.r.squared
Slopes7$Slope <- cf[2]
Slopes7$Intercept <- cf[1]
Slopes7$Top_STD <- 100

head(Slopes7)


####
```
##Slope QAQC
```{r}
##Calculate cv for STDs
#low Fe_2 Curve
FE_2_stds_low1 <- FE_2_stds_low %>%
 group_by(IDs) %>%
 summarise(Abs1_mean = mean(Abs1), Abs1_sd = sd(Abs1), Abs1_cv = cv(Abs1), 
           Dilution = first(Dilutions))

head(FE_2_stds_low1)
##High Fe_2 Curve
FE_2_stds_high1 <- FE_2_stds_high %>%
 group_by(IDs) %>%
 summarise(Abs1_mean = mean(Abs1), Abs1_sd = sd(Abs1), Abs1_cv = cv(Abs1), 
    Dilution = first(Dilutions))

head(FE_2_stds_high1)

#low Fe_Tot Curve
FE_Tot_stds_low1 <- FE_Tot_stds_low %>%
  group_by(IDs) %>%
  summarise(Abs2_mean = mean(Abs2), Abs2_sd = sd(Abs2), Abs2_cv = cv(Abs2), 
         Dilution = first(Dilutions))

head(FE_Tot_stds_low1)

##high Fe_Tot Curve
 FE_Tot_stds_high1 <- FE_Tot_stds_high %>%
 group_by(IDs) %>%
 summarise(Abs2_mean = mean(Abs2), Abs2_sd = sd(Abs2), Abs2_cv = cv(Abs2), 
     Dilution = first(Dilutions))
 
head(FE_Tot_stds_high1)

#Fe_2 abs2 Curve
FE_2_stds_abs2_1 <- FE_2_stds_abs2 %>%
  group_by(IDs) %>%
  summarise(Abs2_mean = mean(Abs2), Abs2_sd = sd(Abs2), Abs2_cv = cv(Abs2), 
            Dilution = first(Dilutions))

head(FE_2_stds_abs2_1)


#Fe_2 diff Curve
FE_2_stds_diff_1 <- FE_2_stds_diff %>%
  group_by(IDs) %>%
  summarise(Abs_dif_mean = mean(FE_2_Diff), Abs_dif = sd(FE_2_Diff), Abs_dif_cv = cv(FE_2_Diff), 
            Dilution = first(Dilutions))
head(FE_2_stds_diff_1)


#Flag high cvs
FE_2_stds_low1$FE_2_cv_flag <- ifelse(FE_2_stds_low1$Abs1_cv > 10, 'High CV', 'within range')
head(FE_2_stds_low1)

FE_2_stds_high1$FE_2_cv_flag <- ifelse(FE_2_stds_high1$Abs1_cv > 10, 'High CV', 'within range')
head(FE_2_stds_high1)

FE_Tot_stds_low1$FE_Tot_cv_flag <- ifelse(FE_Tot_stds_low1$Abs2_cv > 10, 'High CV', 'within range')
head(FE_Tot_stds_low1)

FE_Tot_stds_high1$FE_Tot_cv_flag <- ifelse(FE_Tot_stds_high1$Abs2_cv > 10, 'High CV', 'within range')
head(FE_Tot_stds_high1)

FE_2_stds_abs2_1$FE_2_cv_flag <- ifelse(FE_2_stds_abs2_1$Abs2_cv > 10, 'High CV', 'within range')
head(FE_2_stds_abs2_1)

#flag high cv
FE_2_stds_diff_1$FE_2_cv_flag <- ifelse(FE_2_stds_diff_1$Abs_dif_cv > 10, 'High CV', 'within range')
head(FE_2_stds_diff_1)

#Was the standard curve ok for this plate? Do we need to rerun the plate? Continue means plate is ok. Rerun plate means the R2 is too low.
if(Slopes1$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
}

if(Slopes2$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
}

if(Slopes3$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
}

if(Slopes4$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
}

if(Slopes5$R2 > 0.985){
  print("Continue")
} else {
  print("Rerun plate")
}

if(Slopes7$R2 > 0.985){
  print("Continue")
} else {
  print("Rerun plate")
}

```

#remove high cv

```{r}
#FE_2_low
#manualy remove high cv points
#FE_2_stds_low <- subset(FE_2_stds_low, !(Wells %in% c( "A09","C08" )))
{
##Plot stds and calculate the slope, intercept, and R2 
FE_2_low <- ggplot(FE_2_stds_low, aes(Conc_uM, Abs1)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Low Fe(II) Absorbance 1 Curve")
print(FE_2_low)

FE_2_low_lm <- lm(FE_2_stds_low$Abs1 ~ FE_2_stds_low$Conc_uM)
summary(FE_2_low_lm)
cf <- coef(FE_2_low_lm)

#create data frame with 1 rows and 0 columns
Slopes1 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes1$Date <- Date_Run
Slopes1$Project <- Project
Slopes1$Curve <- "FE_2_low"
Slopes1$R2 <- summary(FE_2_low_lm)$adj.r.squared
Slopes1$Slope <- cf[2]
Slopes1$Intercept <- cf[1]
Slopes1$Top_STD <- Top_STD_Fe2_low
head(Slopes1)

##Calculate cv for STDs

FE_2_stds_low1 <- FE_2_stds_low %>%
  group_by(IDs) %>%
  summarise(Abs1_mean = mean(Abs1), Abs1_sd = sd(Abs1), Abs1_cv = cv(Abs1), 
           Dilution = first(Dilutions))

head(FE_2_stds_low1)

##Flag high cvs
FE_2_stds_low1$FE_2_cv_flag <- ifelse(FE_2_stds_low1$Abs1_cv > 10, 'High CV', 'within range')
print(FE_2_stds_low1)


#Was the standard curve ok for this plate? Do we need to rerun the plate? Continue means plate is ok. Rerun plate means the R2 is too low.
if(Slopes1$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
}
}

#FE_2_high
##manualy remove high cv points
FE_2_stds_high <- subset(FE_2_stds_high, !(Wells %in% c( "C11","A12")))
{ 
 #Plot stds and calculate the slope, intercept, and R2 
 FE_2_high <- ggplot(FE_2_stds_high, aes(Conc_uM, Abs1)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("High Fe(II) Absorbance 1 Curve")
 print(FE_2_high)
 
 FE_2_high_lm <- lm(FE_2_stds_high$Abs1 ~ FE_2_stds_high$Conc_uM)
 summary(FE_2_high_lm)
 cf <- coef(FE_2_high_lm)
 
 #create data frame with 1 rows and 0 columns
Slopes2 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes2$Date <- Date_Run
Slopes2$Project <- Project
Slopes2$Curve <- "FE_2_high"
Slopes2$R2 <- summary(FE_2_high_lm)$adj.r.squared
Slopes2$Slope <- cf[2]
Slopes2$Intercept <- cf[1]
Slopes2$Top_STD <- Top_STD_Fe2_high
head(Slopes2)
 
 
# #Calculate cv for STDs
 
 FE_2_stds_high1 <- FE_2_stds_high %>%
   group_by(IDs) %>%
   summarise(Abs1_mean = mean(Abs1), Abs1_sd = sd(Abs1), Abs1_cv = cv(Abs1), 
             Dilution = first(Dilutions))
 
 head(FE_2_stds_high1)
 
 #Flag high cvs
 FE_2_stds_high1$FE_2_cv_flag <- ifelse(FE_2_stds_high1$Abs1_cv > 10, 'High CV', 'within range')

 print(FE_2_stds_high1)

 #Was the standard curve ok for this plate? Do we need to rerun the plate? 
#Continue means plate is ok. Rerun plate means the R2 is too high.
 if(Slopes2$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
}
}

 #FE_Tot_low
##manualy remove high cv points
FE_Tot_stds_low <- subset(FE_Tot_stds_low, !(Wells %in% c( "F09")))
{ 
#Plot stds and calculate the slope, intercept, and R2 
FE_Tot_low <- ggplot(FE_Tot_stds_low, aes(Conc_uM, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Low Fe(III) Absorbance 2 Curve")
 print(FE_Tot_low)
 
 FE_Tot_low_lm <- lm(FE_Tot_stds_low$Abs2 ~ FE_Tot_stds_low$Conc_uM)
 summary(FE_Tot_low_lm)
 cf <- coef(FE_Tot_low_lm)
 
Slopes3 <- data.frame(matrix(ncol = 0, nrow = 1))
 Slopes3$Date <- Date_Run
 Slopes3$Project <- Project
 Slopes3$Curve <- "FE_Tot_low"
 Slopes3$R2 <- summary(FE_Tot_low_lm)$adj.r.squared
 Slopes3$Slope <- cf[2]
 Slopes3$Intercept <- cf[1]
 Slopes3$Top_STD <- Top_STD_Fe3_low
 head(Slopes2)

 #Calculate cv for STDs
 
 FE_Tot_stds_low1 <- FE_Tot_stds_low %>%
   group_by(IDs) %>%
   summarise(Abs2_mean = mean(Abs2), Abs2_sd = sd(Abs2), Abs2_cv = cv(Abs2), 
             Dilution = first(Dilutions))

head(FE_Tot_stds_low1)
 
 #Flag high cvs
FE_Tot_stds_low1$FE_Tot_cv_flag <- ifelse(FE_Tot_stds_low1$Abs2_cv > 10, 'High CV', 'within range')
 print(FE_Tot_stds_low1)
 
 
 #Was the standard curve ok for this plate? Do we need to rerun the plate? Continue means plate is ok. Rerun plate means the R2 is too low.
if(Slopes3$R2 > 0.985){
print("Continue")
} else {
print("Rerun plate")
 }
}

##FE_Tot_high
##manualy remove high cv points
FE_Tot_stds_high <- subset(FE_Tot_stds_high, !(Wells %in% c("F09")))
{ 
 #Plot stds and calculate the slope, intercept, and R2 
FE_Tot_high <- ggplot(FE_Tot_stds_high, aes(Conc_uM, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("High Fe(III) Absorbance 2 Curve")
 print(FE_Tot_high)
 
FE_Tot_high_lm <- lm(FE_Tot_stds_high$Abs2 ~ FE_Tot_stds_high$Conc_uM)
summary(FE_Tot_high_lm)
cf <- coef(FE_Tot_high_lm)
 
 #create data frame with 1 rows and 0 columns
Slopes4 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes4$Date <- Date_Run
Slopes4$Project <- Project
Slopes4$Curve <- "FE_Tot_high"
Slopes4$R2 <- summary(FE_Tot_high_lm)$adj.r.squared
Slopes4$Slope <- cf[2]
Slopes4$Intercept <- cf[1]
Slopes4$Top_STD <- Top_STD_Fe3_high
head(Slopes4)
 
 #Calculate cv for STDs
 
FE_Tot_stds_high1 <- FE_Tot_stds_high %>%
 group_by(IDs) %>%
  summarise(Abs2_mean = mean(Abs2), Abs2_sd = sd(Abs2), Abs2_cv = cv(Abs2), 
            Dilution = first(Dilutions))
 
head(FE_Tot_stds_high1)

 #Flag high cvs
FE_Tot_stds_high1$FE_Tot_cv_flag <- ifelse(FE_Tot_stds_high1$Abs2_cv > 10, 'High CV', 'within range')
print(FE_Tot_stds_high1)
 
 
#Was the standard curve ok for this plate? Do we need to rerun the plate? Continue means plate is ok. Rerun plate means the R2 is too high. 
if(Slopes4$R2 > 0.985)
{
print("Continue")
} else {
print("Rerun plate")
}
}

#Fe2 Absorbance 2
#manualy remove high cv points for Fe2 abs2
#FE_2_stds_abs2 <- subset(FE_2_stds_abs2, !(Wells %in% c( "A01" )))
{
##Plot stds and calculate the slope, intercept, and R2 
FE_2_abs2 <- ggplot(FE_2_stds_abs2, aes(Conc_uM, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(II) Absorbance 2 Curve")
print(FE_2_abs2)

FE_2_stds_abs2_lm <- lm(FE_2_stds_abs2$Abs2 ~ FE_2_stds_abs2$Conc_uM)
summary(FE_2_stds_abs2_lm)
cf <- coef(FE_2_stds_abs2_lm)

#create data frame with 1 rows and 0 columns
Slopes5 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes5$Date <- Date_Run
Slopes5$Project <- Project
Slopes5$Curve <- "FE_2_abs2"
Slopes5$R2 <- summary(FE_2_stds_abs2_lm)$adj.r.squared
Slopes5$Slope <- cf[2]
Slopes5$Intercept <- cf[1]
Slopes5$Top_STD <- 100
head(Slopes5)

##Calculate cv for STDs
#Fe_2 abs2 Curve
FE_2_stds_abs2_1 <- FE_2_stds_abs2 %>%
  group_by(IDs) %>%
  summarise(Abs2_mean = mean(Abs2), Abs2_sd = sd(Abs2), Abs2_cv = cv(Abs2), 
            Dilution = first(Dilutions))
head(FE_2_stds_abs2_1)

#flag high cv
FE_2_stds_abs2_1$FE_2_cv_flag <- ifelse(FE_2_stds_abs2_1$Abs2_cv > 10, 'High CV', 'within range')
print(FE_2_stds_abs2_1)


if(Slopes5$R2 > 0.985){
  print("Continue")
} else {
  print("Rerun plate")
}
}

#curve for Fe(III) Abs1
#Remove outliers
#FE_Tot_stds_abs1<- subset(FE_Tot_stds, !(Wells %in% c("E02","E09")))
{
FE_3_abs1 <- ggplot(FE_Tot_stds_abs1, aes(Conc_uM,Abs1 )) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle(" Fe(III) Absorbance 1 Curve")
print(FE_3_abs1)

FE_3_abs1_lm <- lm(FE_Tot_stds_abs1$Abs1 ~ FE_Tot_stds_abs1$Conc_uM )
summary(FE_3_abs1_lm)
cf <- coef(FE_3_abs1_lm)

#create data frame with 1 rows and 0 columns
Slopes6 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes6$Date <- Date_Run
Slopes6$Project <- Project
Slopes6$Curve <- "FE_3_abs1"
Slopes6$R2 <- summary(FE_3_abs1_lm)$adj.r.squared
Slopes6$Slope <- cf[2]
Slopes6$Intercept <- cf[1]
Slopes6$Top_STD <- 500
head(Slopes6)

#This slope should be small e-5 to e-6
}

#curve for Fe(II) Difference (Abs2-Abs1)
#manually remove points
FE_2_stds_diff <- subset(FE_2_stds_diff, !(Wells %in% c("B08","A08","C08","A07","A09","C10")))
{
FE_2_stds_diff$FE_2_Diff<-FE_2_stds_diff$Abs2-FE_2_stds_diff$Abs1


FE_2 <- ggplot(FE_2_stds_diff, aes(Conc_uM,FE_2_Diff)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+ggtitle(" Fe(II) Absorbance Difference Curve")
print(FE_2)

FE_2_diff_lm <- lm(FE_2_stds_diff$FE_2_Diff ~ FE_2_stds_diff$Conc_uM)
summary(FE_2_diff_lm)
cf <- coef(FE_2_diff_lm)

#create data frame with 1 rows and 0 columns
Slopes7 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes7$Date <- Date_Run
Slopes7$Project <- Project
Slopes7$Curve <- "FE_2_diff"
Slopes7$R2 <- summary(FE_2_diff_lm)$adj.r.squared
Slopes7$Slope <- cf[2]
Slopes7$Intercept <- cf[1]
Slopes7$Top_STD <- 100
head(Slopes7)

FE_2_stds_diff_1 <- FE_2_stds_diff %>%
  group_by(IDs) %>%
  summarise(Abs_dif_mean = mean(FE_2_Diff), Abs_dif = sd(FE_2_Diff), Abs_dif_cv = cv(FE_2_Diff), 
            Dilution = first(Dilutions))
head(FE_2_stds_diff_1)

#flag high cv
FE_2_stds_diff_1$FE_2_cv_flag <- ifelse(FE_2_stds_diff_1$Abs_dif_cv > 10, 'High CV', 'within range')
print(FE_2_stds_diff_1)

if(Slopes7$R2 > 0.985){
  print("Continue")
} else {
  print("Rerun plate")
}
}

#pull all the slopes together into one dataframe 
Slopes <- rbind(Slopes1, Slopes2, Slopes3, Slopes4, Slopes5, Slopes6, Slopes7)
print(Slopes)

```

# Checking STD Data against QAQC file

```{r}
#read in datafile with all the slopes
qlogO <- read.csv("S:/GlobalChangeEco/porewater_r scripts & Data/BGC_GCE_PorewaterAnalysis/Fe/Fe_STD_QAQC.csv")
qlogO <- qlogO[,c(1:7)]
colnames(qlogO) <- c("Date", "Project", "Curve", "R2", "Slope", "Intercept", "Top_STD")
head(qlogO)

#add data to file
qlog <- rbind(Slopes, qlogO)
head(qlog)


#Separate out by slope type for QAQC
qlog_Fe2_low <- subset(qlog, Curve == "FE_2_low")
qlog_Fe2_high <- subset(qlog, Curve == "FE_2_high")
qlog_FeTot_low <- subset(qlog, Curve == "FE_Tot_low")
qlog_FeTot_high <- subset(qlog, Curve == "FE_Tot_high")
qlog_Fe2_abs2 <- subset(qlog, Curve == "FE_2_abs2")
qlog_Fe3_abs1 <- subset(qlog, Curve == "FE_3_abs1")
qlog_Fe2_diff <- subset(qlog, Curve == "FE_2_diff")

# pull date run
qlog_Fe2_low_Date <- subset(qlog_Fe2_low, Date == Date_Run)
qlog_Fe2_high_Date <- subset(qlog_Fe2_high, Date == Date_Run)
qlog_FeTot_low_Date <- subset(qlog_FeTot_low, Date == Date_Run)
qlog_FeTot_high_Date <- subset(qlog_FeTot_high, Date == Date_Run)
qlog_Fe2_abs2_Date <- subset(qlog_Fe2_abs2, Date == Date_Run)
qlog_Fe3_abs1_Date <- subset(qlog_Fe3_abs1, Date == Date_Run)
qlog_Fe2_diff_Date <- subset(qlog_Fe2_diff, Date == Date_Run)
```

#Check Slopes against QAQC file
```{r}
##plot the slopes to make sure there are no crazy outliers 
#Fe2 low curve
slope_Fe2_low <- ggplot(data=qlog_Fe2_low, aes(x=Date, y=Slope)) +
           geom_hline(yintercept= (mean(qlog_Fe2_low$Slope)+ (2*sd(qlog_Fe2_low$Slope))), linetype="dashed", color = "red", size=2)+
            geom_hline(yintercept= (mean(qlog_Fe2_low$Slope)- (2*sd(qlog_Fe2_low$Slope))), linetype="dashed", color = "red", size=2)+
            geom_point(aes(size=3)) + 
            theme_classic() + ylim(0.001, 0.004) + 
           theme(legend.position="none") + 
           ggtitle("Fe2_low Slopes")+
  guides(x = guide_axis(angle = 70))
  
slope_Fe2_low
#Fe2 high curve
 slope_Fe2_high <- ggplot(data=qlog_Fe2_high, aes(x=Date, y=Slope)) +
            geom_hline(yintercept= (mean(qlog_Fe2_high$Slope)+ (2*sd(qlog_Fe2_high$Slope))), linetype="dashed", color = "red", size=2)+
             geom_hline(yintercept= (mean(qlog_Fe2_high$Slope)- (2*sd(qlog_Fe2_high$Slope))), linetype="dashed", color = "red", size=2)+
             geom_point(aes(size=3)) + 
             theme_classic() + ylim(0.0, 0.003) + 
            theme(legend.position="none") + 
            ggtitle("Fe2_high Slopes")+
  guides(x = guide_axis(angle = 70))
   
slope_Fe2_high
#Fe3 low curve
slope_FeTot_low <- ggplot(data=qlog_FeTot_low, aes(x=Date, y=Slope)) +
           geom_hline(yintercept= (mean(qlog_FeTot_low$Slope)+ (2*sd(qlog_FeTot_low$Slope))), linetype="dashed", color = "red", size=2)+
            geom_hline(yintercept= (mean(qlog_FeTot_low$Slope)- (2*sd(qlog_FeTot_low$Slope))), linetype="dashed", color = "red", size=2)+
            geom_point(aes(size=3)) + 
            theme_classic() + ylim(0.0, 0.0075) + 
           theme(legend.position="none") + 
           ggtitle("FeTot_low Slopes")+
  guides(x = guide_axis(angle = 70))
  
slope_FeTot_low
#Fe3 high curve
 slope_FeTot_high <- ggplot(data=qlog_FeTot_high, aes(x=Date, y=Slope)) +
            geom_hline(yintercept= (mean(qlog_FeTot_high$Slope)+ (2*sd(qlog_FeTot_high$Slope))), linetype="dashed", color = "red", size=2)+
             geom_hline(yintercept= (mean(qlog_FeTot_high$Slope)- (2*sd(qlog_FeTot_high$Slope))), linetype="dashed", color = "red", size=2)+
             geom_point(aes(size=3)) + 
             theme_classic() + ylim(0.0, 0.003) + 
            theme(legend.position="none") + 
            ggtitle("FeTot_high Slopes")+
  guides(x = guide_axis(angle = 70))
   
 slope_FeTot_high
#Fe2 abs2
 slope_Fe2_abs2 <- ggplot(data=qlog_Fe2_abs2, aes(x=Date, y=Slope)) +
  geom_hline(yintercept= (mean(qlog_Fe2_abs2$Slope)+ (2*sd(qlog_Fe2_abs2$Slope))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe2_abs2$Slope)- (2*sd(qlog_Fe2_abs2$Slope))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.002, 0.006) + 
  theme(legend.position="none") + 
  ggtitle("Fe2_abs2 Slopes")+
  guides(x = guide_axis(angle = 70))

slope_Fe2_abs2
#Fe3 abs1
 slope_Fe3_abs1 <- ggplot(data=qlog_Fe3_abs1, aes(x=Date, y=Slope)) +
  geom_hline(yintercept= (mean(qlog_Fe3_abs1$Slope)+ (2*sd(qlog_Fe3_abs1$Slope))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe3_abs1$Slope)- (2*sd(qlog_Fe3_abs1$Slope))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0, 0.00003) + 
  theme(legend.position="none") + 
  ggtitle("Fe3_abs1 Slopes")+
  guides(x = guide_axis(angle = 70))

slope_Fe3_abs1
#Fe2 difference
 slope_Fe2_diff <- ggplot(data=qlog_Fe2_diff, aes(x=Date, y=Slope)) +
  geom_hline(yintercept= (mean(qlog_Fe2_diff$Slope)+ (2*sd(qlog_Fe2_diff$Slope))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe2_diff$Slope)- (2*sd(qlog_Fe2_diff$Slope))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.001, 0.002) + 
  theme(legend.position="none") + 
  ggtitle("Fe2_diff Slopes")+
  guides(x = guide_axis(angle = 70))

slope_Fe2_diff

if (qlog_Fe2_low_Date$Slope[1] > (mean(qlog_Fe2_low$Slope)+ (2*sd(qlog_Fe2_low$Slope)))| qlog_Fe2_low_Date$Slope[1] < (mean(qlog_Fe2_low$Slope)- (2*sd(qlog_Fe2_low$Slope)))) {
print("rerun Plate")
} else {
print("continue")
}

 if (qlog_Fe2_high_Date$Slope[1] > (mean(qlog_Fe2_high$Slope)+ (2*sd(qlog_Fe2_high$Slope)))| qlog_Fe2_high_Date$Slope[1] < (mean(qlog_Fe2_high$Slope)- (2*sd(qlog_Fe2_high$Slope)))) {
 print("rerun Plate")
 } else {
 print("continue")
 }

if (qlog_FeTot_low_Date$Slope[1] > (mean(qlog_FeTot_low$Slope)+ (2*sd(qlog_FeTot_low$Slope)))| qlog_FeTot_low_Date$Slope[1] < (mean(qlog_FeTot_low$Slope)- (2*sd(qlog_FeTot_low$Slope)))) {
print("rerun Plate")
} else {
print("continue")
}

 if (qlog_FeTot_high_Date$Slope[1] > (mean(qlog_FeTot_high$Slope)+ (2*sd(qlog_FeTot_high$Slope)))| qlog_FeTot_high_Date$Slope[1] < (mean(qlog_FeTot_high$Slope)- (2*sd(qlog_FeTot_high$Slope)))) {
 print("rerun Plate")
 } else {
 print("continue")
 }

if (qlog_Fe2_abs2_Date$Slope[1] > (mean(qlog_Fe2_abs2$Slope)+ (2*sd(qlog_Fe2_abs2$Slope)))| qlog_Fe2_abs2_Date$Slope[1] < (mean(qlog_Fe2_abs2$Slope)- (2*sd(qlog_Fe2_abs2$Slope)))) {
  print("rerun Plate")
} else {
  print("continue")
}

if (qlog_Fe3_abs1_Date$Slope[1] > (mean(qlog_Fe3_abs1$Slope)+ (2*sd(qlog_Fe3_abs1$Slope)))| qlog_Fe3_abs1_Date$Slope[1] < (mean(qlog_Fe3_abs1$Slope)- (2*sd(qlog_Fe3_abs1$Slope)))) {
  print("rerun Plate")
} else {
  print("continue")
}

if (qlog_Fe2_diff_Date$Slope[1] > (mean(qlog_Fe2_diff$Slope)+ (2*sd(qlog_Fe2_diff$Slope)))| qlog_Fe2_diff_Date$Slope[1] < (mean(qlog_Fe2_diff$Slope)- (2*sd(qlog_Fe2_diff$Slope)))) {
  print("rerun Plate")
} else {
  print("continue")
}
#Rerun if outside of red lines
```
# Check Intercepts against QAQC file

```{r}
##plot the intercepts to make sure there are no crazy outliers 
#Fe2 low
int_Fe2_low <- ggplot(data=qlog_Fe2_low, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog_Fe2_low$Intercept)+ (2*sd(qlog_Fe2_low$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe2_low$Intercept)- (2*sd(qlog_Fe2_low$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.03,0.05) + 
  theme(legend.position="none")+ 
  ggtitle("Fe2_low Intercepts")+
  guides(x = guide_axis(angle = 70))

int_Fe2_low
#Fe2 high
 int_Fe2_high <- ggplot(data=qlog_Fe2_high, aes(x=Date, y=Intercept)) +
   geom_hline(yintercept= (mean(qlog_Fe2_high$Intercept)+ (2*sd(qlog_Fe2_high$Intercept))), linetype="dashed", color = "red", size=2)+
   geom_hline(yintercept= (mean(qlog_Fe2_high$Intercept)- (2*sd(qlog_Fe2_high$Intercept))), linetype="dashed", color = "red", size=2)+
   geom_point(aes(size=3)) + 
   theme_classic() + ylim(0.03,0.06) + 
   theme(legend.position="none")+ 
   ggtitle("Fe2_high Intercepts")+
  guides(x = guide_axis(angle = 70))
 
 int_Fe2_high
#fe3 low
int_FeTot_low <- ggplot(data=qlog_FeTot_low, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog_FeTot_low$Intercept)+ (2*sd(qlog_FeTot_low$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_FeTot_low$Intercept)- (2*sd(qlog_FeTot_low$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.03,0.05) + 
  theme(legend.position="none")+ 
  ggtitle("FeTot_low Intercepts")+
  guides(x = guide_axis(angle = 70))

int_FeTot_low
#fe3 high
 int_FeTot_high <- ggplot(data=qlog_FeTot_high, aes(x=Date, y=Intercept)) +
   geom_hline(yintercept= (mean(qlog_FeTot_high$Intercept)+ (2*sd(qlog_FeTot_high$Intercept))), linetype="dashed", color = "red", size=2)+
   geom_hline(yintercept= (mean(qlog_FeTot_high$Intercept)- (2*sd(qlog_FeTot_high$Intercept))), linetype="dashed", color = "red", size=2)+
   geom_point(aes(size=3)) + 
   theme_classic() + ylim(0.03,0.05) + 
   theme(legend.position="none")+ 
   ggtitle("FeTot_high Intercepts")+
  guides(x = guide_axis(angle = 70))
 
 int_FeTot_high
#Fe2 abs2
 int_Fe2_abs2 <- ggplot(data=qlog_Fe2_abs2, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog_Fe2_abs2$Intercept)+ (2*sd(qlog_Fe2_abs2$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe2_abs2$Intercept)- (2*sd(qlog_Fe2_abs2$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.03,0.05) + 
  theme(legend.position="none")+ 
  ggtitle("Fe2_abs2 Intercepts")+
  guides(x = guide_axis(angle = 70))

int_Fe2_abs2
#Fe3 abs1
 int_Fe3_abs1 <- ggplot(data=qlog_Fe3_abs1, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog_Fe3_abs1$Intercept)+ (2*sd(qlog_Fe3_abs1$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe3_abs1$Intercept)- (2*sd(qlog_Fe3_abs1$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.03,0.05) + 
  theme(legend.position="none")+ 
  ggtitle("Fe3_abs1 Intercepts")+
  guides(x = guide_axis(angle = 70))

int_Fe3_abs1
#Fe2 diff
 int_Fe2_diff <- ggplot(data=qlog_Fe2_diff, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog_Fe2_diff$Intercept)+ (2*sd(qlog_Fe2_diff$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_hline(yintercept= (mean(qlog_Fe2_diff$Intercept)- (2*sd(qlog_Fe2_diff$Intercept))), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0,0.01) + 
  theme(legend.position="none")+ 
  ggtitle("Fe2_diff Intercepts")+
  guides(x = guide_axis(angle = 70))

int_Fe2_diff

if (qlog_Fe2_low_Date$Intercept[1] > (mean(qlog_Fe2_low$Intercept)+ (2*sd(qlog_Fe2_low$Intercept)))| qlog_Fe2_low_Date$Intercept[1] < (mean(qlog_Fe2_low$Intercept)- (2*sd(qlog_Fe2_low$Intercept)))) {
print("rerun Plate")
} else {
print("continue")
}

 if (qlog_Fe2_high_Date$Intercept[1] > (mean(qlog_Fe2_high$Intercept)+ (2*sd(qlog_Fe2_high$Intercept)))| qlog_Fe2_high_Date$Intercept[1] < (mean(qlog_Fe2_high$Intercept)- (2*sd(qlog_Fe2_high$Intercept)))) {
 print("rerun Plate")
 } else {
 print("continue")
 }

if (qlog_FeTot_low_Date$Intercept[1] > (mean(qlog_FeTot_low$Intercept)+ (2*sd(qlog_FeTot_low$Intercept)))| qlog_FeTot_low_Date$Intercept[1] < (mean(qlog_FeTot_low$Intercept)- (2*sd(qlog_FeTot_low$Intercept)))) {
print("rerun Plate")
} else {
print("continue")
}

 if (qlog_FeTot_high_Date$Intercept[1] > (mean(qlog_FeTot_high$Intercept)+ (2*sd(qlog_FeTot_high$Intercept)))| qlog_FeTot_high_Date$Intercept[1] < (mean(qlog_FeTot_high$Intercept)- (2*sd(qlog_FeTot_high$Intercept)))) {
 print("rerun Plate")
 } else {
 print("continue")
 }

if (qlog_Fe2_abs2_Date$Intercept[1] > (mean(qlog_Fe2_abs2$Intercept)+ (2*sd(qlog_Fe2_abs2$Intercept)))| qlog_Fe2_abs2_Date$Intercept[1] < (mean(qlog_Fe2_abs2$Intercept)- (2*sd(qlog_Fe2_abs2$Intercept)))) {
  print("rerun Plate")
} else {
  print("continue")
}

if (qlog_Fe3_abs1_Date$Intercept[1] > (mean(qlog_Fe3_abs1$Intercept)+ (2*sd(qlog_Fe3_abs1$Intercept)))| qlog_Fe3_abs1_Date$Intercept[1] < (mean(qlog_Fe3_abs1$Intercept)- (2*sd(qlog_Fe3_abs1$Intercept)))) {
  print("rerun Plate")
} else {
  print("continue")
}

if (qlog_Fe2_diff_Date$Intercept[1] > (mean(qlog_Fe2_diff$Intercept)+ (2*sd(qlog_Fe2_diff$Intercept)))| qlog_Fe2_diff_Date$Intercept[1] < (mean(qlog_Fe2_diff$Intercept)- (2*sd(qlog_Fe2_diff$Intercept)))) {
  print("rerun Plate")
} else {
  print("continue")
}
```
# Check R2s against QAQC file

```{r}
#plot the R2s to make sure there are no crazy outliers 
#Fe2 low
Rsq1_Fe2_low <- ggplot(data=qlog_Fe2_low, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.96, 1.01) + 
  theme(legend.position="none")+ 
  ggtitle("Fe2_low R2s")+
  guides(x = guide_axis(angle = 70))

Rsq1_Fe2_low
#Fe2 high
 Rsq1_Fe2_high <- ggplot(data=qlog_Fe2_high, aes(x=Date, y=R2)) +
   geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
   geom_point(aes(size=3)) + 
   theme_classic() + ylim(0.96, 1.01) + 
   theme(legend.position="none")+ 
   ggtitle("Fe2_high R2s")+
  guides(x = guide_axis(angle = 70))

 Rsq1_Fe2_high
#Fe3 low
Rsq1_FeTot_low <- ggplot(data=qlog_FeTot_low, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.96, 1.01) + 
  theme(legend.position="none")+ 
  ggtitle("FeTot_low R2s")+
  guides(x = guide_axis(angle = 70))

Rsq1_FeTot_low
#Fe3 high
 Rsq1_FeTot_high <- ggplot(data=qlog_FeTot_high, aes(x=Date, y=R2)) +
   geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
   geom_point(aes(size=3)) + 
   theme_classic() + ylim(0.96, 1.01) + 
   theme(legend.position="none")+ 
   ggtitle("FeTot_high R2s")+
  guides(x = guide_axis(angle = 70))
 
 Rsq1_FeTot_high
#Fe2 abs2 
 Rsq1_Fe2_abs2 <- ggplot(data=qlog_Fe2_abs2, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.96, 1.01) + 
  theme(legend.position="none")+ 
  ggtitle("Fe2_abs2 R2s")+
  guides(x = guide_axis(angle = 70))

Rsq1_Fe2_abs2
#Fe3 abs1
 Rsq1_Fe3_abs1 <- ggplot(data=qlog_Fe3_abs1, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0, 1.0) + 
  theme(legend.position="none")+ 
  ggtitle("Fe3_abs1 R2s")+
  guides(x = guide_axis(angle = 70))

Rsq1_Fe3_abs1
#Fe2 diff
 Rsq1_Fe2_diff <- ggplot(data=qlog_Fe2_diff, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", size=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.96, 1.01) + 
  theme(legend.position="none")+ 
  ggtitle("Fe2_diff R2s")+
  guides(x = guide_axis(angle = 70))

Rsq1_Fe2_diff
```

## Calculate Reduction Efficiency 
```{r}
#pull out Fe_2 and Fe_3 standards at 60uM to check reduction efficiency

#subset by FE_2 and FE_Total
FE_2_60uM <- stds %>%  
  filter(str_detect(stds$IDs, "60 uM Fe_2 A"))  
head(FE_2_60uM)

#subset by FE_2 and FE_Total
FE_3_60 <- stds %>%  
  filter(str_detect(IDs, "60 uM Fe_3 A"))  
head(FE_3_60)


#create data frame with 0 rows and 3 columns
Fe_Red_eff <- data.frame(matrix(ncol = 0, nrow = 3))

Fe_Red_eff$ID <- seq.int(nrow(Fe_Red_eff))
Fe_Red_eff$FE2 <- (FE_2_60uM$Abs1)
Fe_Red_eff$FE3 <- FE_3_60$Abs2
Fe_Red_eff$Eff <- ((Fe_Red_eff$FE3)/Fe_Red_eff$FE2)*100
Fe_Red_eff$Eff1 <- ((Fe_Red_eff$FE3/0.8)/Fe_Red_eff$FE2)*100 #this one accounts for dilution of the reagents 
head(Fe_Red_eff)

#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 
redeffbar <- ggplot(data = Fe_Red_eff, aes(x = ID, y = Eff1)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("gray48")) + 
        theme_classic() + labs(x= "Sample ID", y="Reduction Efficiency (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=85, linetype="dashed", color = "black", size=1) + 
        geom_hline(yintercept=120, linetype="dashed", color = "black", size=1)

redeffbar

```

## Method Minimum Detection Limit (MDL) Calculation
I use the lowest standard to calculate MDL
How to calculate MDL https://www.epa.gov/sites/default/files/2016-12/documents/mdl-procedure_rev2_12-13-2016.pdf
Look at Table 1 on page 5
The Student's t-value used to calculate the method detection limit (MDL) is the one appropriate for a 99% confidence level and a standard deviation estimate with n-1 degrees of freedom

```{r}
#make a dataframe of just 0.6 uM Fe2 samples from standard dataframe
MDL_dat <- subset(FE_2_stds, Conc_uM == 0.6)

#calculate the standard deviation of abs 1
stddev <- sd(MDL_dat$Abs1)

#calculate n (number of 0.6 uM samples)
n <- nrow(MDL_dat)

#Identify the proper Student's T for n-1 from the table link above
StudentsT <- 2.528

#Multiply Standard deviation by Student's T
MDL <- stddev*StudentsT
MDL
```
#save QAQC Data to QAQC File

```{r}
#save QAQC Data to QAQC File
#write.csv(qlog, file="S:/GlobalChangeEco/porewater_r scripts & Data/BGC_GCE_PorewaterAnalysis/Fe/Fe_STD_QAQC.csv", row.names=FALSE)
```



```

### END



