---
title: "COMPASS_Synoptic_SmartChamberFluxes_2024_CO2"
author: "Stephanie J. Wilson"
date: "2025-01-30"
output: html_document
---
## Goals of this code: 
1.	Reads in raw smart chamber
2.	If no metadatafile, write one out & fields will be: 
Label 
the IDs pulled from the label, 
Plot *fixed so that if there are any issues in the label they are good here 
Drop column – false as long as label looks right 
adjust time – would be initially populated by the start time in the OR blank 
adjust observation length – initially populated by the obs length  OR blank 
notes / comment 
3.	Code reads in the meta data and applies it to the data 
a.	i.e. drops fluxes that have the drop column TRUE or FLAG this 
b.	adjusting times as needed 
4.	Now you compute fluxes 
5.	Visualize and look for issues 
6.	If found, report to meta data file then go back to 3 and reanalyze 
7.	Apply algorithmic qaqc and generate flags 


## Set Up & Load Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(stringr)

```

## Pull in Smart Chamber Files from the Raw Data folder
```{r}

setwd("Raw Data/")

#list all the files in the directory
L <- list.files(path = "Raw Data/")
head(L)

#create a list of the files there 
file.list <- dir(pattern = "json$", full.names = FALSE)

#Read in the data 
alldat1 <- lapply(file.list, ffi_read_LIsmartchamber)

#combind into one dataframe
alldat <- data.table::rbindlist(alldat1, fill=TRUE)


```

## Create a unique Plot_ID, check if there is a meta-data file, if not, create a meta-data file
```{r}

############### only run this first part if you don't already have a meta-data file 
        #because if you do then all the data for each one will have a plot and you won't be able to change start times


#create unique Plot ID for each measurement: 
#detach("package:data.table", unload=TRUE)
alldat$month <- as.character(month(alldat$TIMESTAMP, label=TRUE, abbr=TRUE)) #will not run if you have data.table loaded 

alldat$year <- year(alldat$TIMESTAMP)

alldat <- alldat %>% 
  mutate(Plot_ID = str_c(pmin(month, RepNum), 
                      pmax(month, RepNum),  sep = '-'))
alldat <- alldat %>% 
  mutate(Plot_ID = str_c(pmin(Plot_ID, label), 
                     pmax(Plot_ID, label),  sep = '-'))  
#this needs to be PLot ID so that later we ID plot from metadata and not the metadata row 



# Define the file path where you want to save the CSV file
file_path <- "COMPASS_SmartChamber_Metadata_2024.csv"

# Check if the CSV file exists
if (file.exists(file_path)) {
  # If it exists, read it back into R
  metadata <- read.csv(file_path)
  print("CSV file exists and has been read into the code.")
} else {
  # If it does not exist, create the CSV file
    #create a metadata file from the Plot IDs 
      # pull timestart timestamps to put in the metadata and obs lengths 
  metadata <- alldat %>%
  group_by(Plot_ID) %>%
  summarise(
    Date = format(min(TIMESTAMP), "%Y-%m-%d"),
    Time_start = format(min(TIMESTAMP), "%H:%M:%S"),
    First_Timestamp = min(TIMESTAMP),  # Minimum (earliest) timestamp for each Plot_ID
    Last_Timestamp = max(TIMESTAMP),   # Maximum (latest) timestamp for each Plot_ID
    Total_Time = as.numeric(difftime(max(TIMESTAMP), min(TIMESTAMP), units = "secs")),  # Time difference in seconds
    ) %>%
  ungroup() %>%
  mutate(drop = FALSE) %>% # Add a 'drop' column with 'FALSE' for every row 
  #mutate(notes = " ")  # Add a 'drop' column with 'FALSE' for every row
  #mutate(Deadband = 5)
    mutate(Notes = "") #%>%
    #mutate(Plot = Plot_ID) #think need this because we need to make a plot column
  
  # Write df_unique to CSV
  write.csv(metadata, file = file_path, row.names = FALSE)
  
  print("CSV file does not exist. It has been created.")
}


```

## Mark Drop column in Metadata for Retaken Fluxes or other reasons to Drop 
```{r}

############## Check For Plot ID

##Sometimes there are other fluxes in the .json files that we don't want to calculate 
  #change drop column to TRUE if the plot doesn't include a site code in the Plot_ID 
metadata <- metadata %>%
  mutate(
    drop = ifelse(
      !grepl("SWH|GCW|MSM|GWI|GCReW", Plot), 
      # Check if Plot_ID does not contain the specified strings
      TRUE,  # Set drop to TRUE if it doesn't contain any of the strings
      drop   # Keep the original value of drop if it contains one of the strings
    ) )

############## Check for retakes 

##Sometimes we have to retake a flux in the field due to issues or ebullition
##We need to change the drop column to reflect using only the latest flux
  #In order to we expect an "a" after the label, but sometimes this is not the case so we will fix that below and make it "a" 
metadata <- metadata %>%
    mutate(Plot = case_when(
    str_detect(Plot, "-retake") ~ str_replace(Plot, "-retake", "a"),
    str_detect(Plot, "-Retake") ~ str_replace(Plot, "-Retake", "a"),
    str_detect(Plot, "-b") ~ str_replace(Plot, "-b", "b"),
    str_detect(Plot, "1-May-GCW-TR-3-A") ~ str_replace(Plot, "1-May-GCW-TR-3-A", "1-May-GCW-TR-3a"),
    TRUE ~ Plot  # Keeps the original value if none of the conditions match
  ))

#If the flux was taken more than once, flag the others so that we use the latest measurement 
# Extract the base plot (e.g., "GCW-TR-1" from "GCW-TR-1", "GCW-TR-1a", etc.)
metadata$base_plot <- str_extract(metadata$Plot,  "^[^-]+-[A-Za-z]+-[A-Za-z]+-[A-Za-z]+")

# Extract the replicate number (the number after the last hyphen)
metadata$replicate <- as.numeric(str_extract(metadata$Plot, "(?<=-)[0-9]+(?=[a-zA-Z]?$)"))

# Extract the suffix (if any, the letter after the replicate number)
metadata$suffix <- str_extract(metadata$Plot, "(?<=\\d)[a-zA-Z]$")

# If there's no suffix, set it to an empty string
metadata$suffix[is.na(metadata$suffix)] <- ""

# Sort by base plot, replicate number, and suffix (to keep highest suffix) - #####INSTEAD WE WANT TO CHANGE DROP TO TRUE HERE#######
metadata <- metadata %>%
  group_by(base_plot, replicate) %>%
  arrange(base_plot, replicate, desc(suffix), .by_group = TRUE) %>%
  mutate(drop = row_number() != 1) %>% #the entry with the latest suffix will keep false all others will be true
  #slice_head(n = 1) %>% #if you wanted to remove a the line rather than flag it 
  ungroup()

################# Check Observation lengths 

#if the observation length is over 5 mins then change drop column to true  
metadata <- metadata %>%
  mutate(drop = if_else(Total_Time > 300, TRUE, drop)) 




```



##Dont need to pull L1 temp data or any other temps like we do for macrophyte because the smartchamber records the temp

## Match up the metadata to the rest of the data   
```{r}

#match the altered metadata to the concentration data with fluxfinder ffi_metadata_match
alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = alldat$TIMESTAMP,
  start_dates = metadata$Date,
  start_times = metadata$Time_start,
  obs_lengths = metadata$Total_Time) 
head(alldat)

#alldat$Deadband <- metadata$Deadband[alldat$metadat_row] #flux calcs function cannot handle a column as a deadband and that is a problem to fix
alldat$Plot <- metadata$Plot[alldat$metadat_row] 

#alldat$Drop <- metadata$Drop[alldat$metadat_row] #don't need this for now because we remove the fluxes that were retaken below 

```

## Change the units from ppm to umol  
```{r}
#change the units
alldat$CO2_umol <- ffi_ppm_to_umol(alldat$co2, 
                                volume = (alldat$TotalVolume / 1000000), # have to divide to get to m3
                                temp = alldat$chamber_t)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CO2_umol_m2 <- alldat$CO2_umol / 0.0318 #adjust for the area 


```

## Calculate fluxes
```{r}

#need to create a new line that would say only calculate the flux if the drop column is false
    #before we add this we need to figure out how to drop except for the last flux taken 


#calculate fluxes for all other sites 
fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CO2_umol_m2", 
                             dead_band = 5)
head(fluxes)

allflux <- fluxes


```

## Remove fluxes that were retaken - this works, but I think we want to do this earlier with the drop column???   
```{#r}

#so first we need to make it so that the fluxes that were retaken have similar names or some way to identify them 
  # if they were retaken it will either have a letter after it or a retake written 


#remove any lines that don't have a site code in the Plot ID 
fluxes <-  fluxes %>%  
  filter(str_detect(Plot, "SWH|MSM|GWI|GCW|GCReW"))

#if they say retake then we want to replace that with an "a" and if it has -A then we also want to remove that and make it "a" 
fluxes <- fluxes %>%
  mutate(Plot = if_else(str_detect(Plot, "-retake"), 
                        str_replace(Plot, "-retake", "a"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-Retake"), 
                        str_replace(Plot, "-Retake", "a"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-b"), 
                        str_replace(Plot, "-b", "b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "1-May-GCW-TR-3-A"), 
                        str_replace(Plot, "1-May-GCW-TR-3-A", "1-May-GCW-TR-3a"), 
                        Plot))


# Extract the base plot (e.g., "GCW-TR-1" from "GCW-TR-1", "GCW-TR-1a", etc.)
fluxes$base_plot <- str_extract(fluxes$Plot,  "^[^-]+-[A-Za-z]+-[A-Za-z]+-[A-Za-z]+")

# Extract the replicate number (the number after the last hyphen)
fluxes$replicate <- as.numeric(str_extract(fluxes$Plot, "(?<=-)[0-9]+(?=[a-zA-Z]?$)"))

# Extract the suffix (if any, the letter after the replicate number)
fluxes$suffix <- str_extract(fluxes$Plot, "(?<=\\d)[a-zA-Z]$")

# If there's no suffix, set it to an empty string
fluxes$suffix[is.na(fluxes$suffix)] <- ""

#create ID columns from the plots 
IDs <- data.frame(do.call('rbind', strsplit(as.character(fluxes$Plot),'-',fixed=TRUE)))
colnames(IDs) <- c("RepNumber", "Month", "Site", "Zone", "Replicate") 
head(IDs)

#combine these dataframes together 
allflux <- cbind(IDs, fluxes)

# Sort by base plot, replicate number, and suffix (to keep highest suffix)
allflux <- allflux %>%
  group_by(base_plot, replicate) %>%
  arrange(Month, Site, Zone, replicate, desc(suffix), .by_group = TRUE) %>%
  slice_head(n = 1) %>%
  ungroup()



```

## Visualize fluxes and sort by r2 
```{r}

#take a look at the fluxes we calculated with R2 
ggplot(allflux, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + ylim(-5,100) + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CO2 flux (mmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(allflux, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate > 1, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(allflux$lin_flux.estimate, 0.05) 
high <- quantile(allflux$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(allflux, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(allflux, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r}

tobechecked <- c("2-May-SWH-TR-6a", "2-Sep-SWH-TR-1a"                 )

colnames <- c("Plot")

#Loop to run five random plots subset above 
for(P in tobechecked){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), co2)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

#r1 <- subset(alldat, Plot ==  "2-May-GWI-TR-4b")
#p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), ch4)) + geom_point() + 
 # scale_x_datetime(breaks = "15 sec") + 
 # theme_classic() +
 # ylab(expression(paste( CH [4], " (ppb)"))) +
 # xlab("Time") + labs(title = "plot name above") + 
  #geom_smooth()
#print(p1)


#list of the ones that I changed and checked and looked good? 
fixedplots <- c("2-Jul-GWI-UP-3", "2-Aug-SWH-TR-6a", "1-Oct-SWH-TR-9", "1-Sep-SWH-TR-9",
                "2-May-GWI-TR-4b", "1-Sep-SWH-UP-2", "2-Oct-SWH-TR-1a", "2-Oct-SWH-UP-2", "2-Nov-SWH-SWAMP-2a",
                "2-Oct-GWI-UP-3", "1-Aug-SWH-TR-1", "2-May-SWH-SWAMP-2", "2-Nov-SWH-SWAMP-1", "2-Aug-SWH-TR-1",
                 "2-Aug-SWH-SWAMP-2", "1-Sep-SWH-TR-8b", "1-Oct-SWH-TR-10", "1-Oct-SWH-SWAMP-1a", "1-Nov-SWH-TR-4a", 
                "1-Nov-SWH-SWAMP-2a", "1-Nov-SWH-SWAMP-1",  "1-May-SWH-TR-4a", "1-May-SWH-SWAMP-2", "1-May-GCW-TR-4", 
                "1-Jun-SWH-TR-8b", "2-Jun-SWH-TR-3a", "1-Aug-SWH-TR-8a", "1-Aug-SWH-SWAMP-2", "2-May-GWI-UP-5", "1-Nov-SWH-UP-2",
                "2-Aug-SWH-TR-2c", "1-Oct-SWH-SWAMP-3a", "1-Jun-SWH-UP-4", "1-May-SWH-TR-5a", 
                 "2-Nov-SWH-SWAMP-5a", "2-Aug-SWH-TR-9a", "2-Sep-GCW-TR-5", "2-Aug-SWH-SWAMP-1a", "2-Oct-SWH-SWAMP-2", 
                "1-Aug-SWH-TR-2c", "2-Oct-GCW-TR-4", "2-Aug-SWH-SWAMP-1a", "1-Nov-SWH-UP-4", 
                "1-Jun-SWH-TR-5" , "1-Aug-SWH-TR-5", "2-Oct-GCW-UP-4", "1-Jul-SWH-UP-3", "1-Oct-SWH-SWAMP-4", "2-Sep-SWH-TR-2a", 
                 "2-Jun-SWH-TR-1a", "2-Jun-SWH-SWAMP-1", "2-Jul-SWH-SWAMP-2a", "2-Jul-SWH-SWAMP-1a", "1-Oct-SWH-TR-2a", "2-Sep-GCW-UP-5", 
                 "2-Jun-GCW-UP-5", "2-Aug-GCW-UP-5", "2-Jul-GCW-UP-1", "2-Oct-SWH-TR-3a", "1-Aug-SWH-TR-7a", "2-Nov-SWH-UP-1", "2-Sep-SWH-UP-1",
                "2-Jun-GCW-TR-3", "1-Sep-SWH-TR-3a", "2-Sep-SWH-TR-5", "2-Sep-GWI-UP-4", "2-Oct-SWH-SWAMP-1a", "2-May-GWI-TR-1b", "2-Jul-GCW-TR-2", 
                 "2-Jun-GCW-UP-3", "1-Jul-SWH-UP-2", "2-Jun-GCW-TR-1", "1-Nov-SWH-TR-7a", "2-Nov-SWH-TR-10", "2-Nov-SWH-SWAMP-4",
                 "1-Nov-SWH-TR-8", "1-Apr-MSM-UP-3", "1-Sep-SWH-SWAMP-3", "1-Apr-MSM-TR-1", "2-Jun-SWH-UP-5", "1-Oct-SWH-SWAMP-5a")
  

```


## Spot check a few fluxes with R2 > 0.9  
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), co2)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
}

####################
```

## Take a look at the  fluxes <0.9 and a flux estimate below 5 
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot)

#Loop to run five random plots subset above 
for(P in low_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO2 [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
}

#print flag that says low r2, but low flux - checked 
allflux <- allflux %>%
  mutate(qaqc_flag = if_else(Plot %in% low_plots, "low r2, but also low flux - checked",
                             NA_character_))

####################

```

## Take a look at the bad fluxes <0.9 and have a high flux estimate 
# See if we want to remove these fluxes or if we want to try and recalculate these ones? 
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

#Loop to run five random plots subset above 
for(P in bad_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################

########## Ebulition 

#List the ones that have ebulition - assessed visually for now 
eb_flags <- c("2-Nov-SWH-TR-9a", "1-Nov-SWH-TR-9a", "2-Sep-SWH-TR-7a", "2-Sep-SWH-TR-7a", 
              "2-Jul-SWH-TR-9d", "1-Jul-SWH-TR-7b", "2-Oct-SWH-TR-3a", "2-Nov-SWH-TR-4a", 
              "2-Aug-SWH-TR-8a", "1-Oct-SWH-TR-8a")


#add a flag for the fluxes that have ebullition and therefore not even a flux 
allflux <- allflux %>%
    mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% eb_flags, "significant ebulition, flux compromised", qaqc_flag))


```


## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

#Loop to run five random plots subset above 
for(P in tail_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Remove any fluxes that do not fit criteria 
```{r, echo=FALSE}

# Create a new dataframe by removing rows that were present in fluxes_weird which have a low R2 and a flux higher than 1 (23 obs)
allflux_filtered <- allflux[!(allflux$Plot %in% fluxes_weird$Plot), ]

# View the result
head(allflux_filtered)


```


## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(allflux_filtered, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(allflux_filtered, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(allflux_filtered, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere 
```{r, echo=FALSE}

#can remove base_plot column, the replicate column, and the suffix column
allflux_filtered <- allflux_filtered[,-c(32:34)]

#export as a csv to the folder that you want 
write.csv(allflux_filtered, file="Processed Data/COMPASS_Synoptic_CB_SmartChamber_Fluxes_Final_2024_CO2.csv")           #Change file name


```



