---
title: "COMPASS_Synoptic_MacrophyteFluxes_2024_CO2"
author: "Stephanie J. Wilson"
date: "2024-11-26"
output: pdf_document
---
## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
```

## Pull in Raw Datafiles
```{r}
#read in files
file.list <- dir(path = "Raw Data/", pattern = "txt$", full.names = TRUE)
head(file.list)

#Read in the data 
alldat1 <- lapply(file.list, ffi_read_LI7810)

alldat <- data.table::rbindlist(alldat1, fill=TRUE)

```

## Read in metadata file and match to raw data
```{r}

#read in metadatfile
md <- "Raw Data/COMPASS_ChamberFlux_Metadata_2024.csv"
#metadat <- read_excel(md)
metadat <- read.csv(md) 

head(metadat)  

#can remove any row that have NA's in the date column 
metadat <- metadat %>%
  filter(!is.na(Date))

head(metadat) 

```

## Read in Air Temp Data from Air Temp Code # use after L1 release
```{r}
# Air temp 

#setwd("S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Fluxes/Macrophyte GHG Fluxes/Flux Calculations/Air Temp Data")
#download the L1 dataset from: https://drive.google.com/drive/folders/1RP-hxmjAYNGzTfCNpYbMN6U_zbZmEWHM

# ##Read in L1 Sensor Data from S Drive and only save necessary data
# # Path to zip file
# L1 <- "S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Synoptic L1 Sensor/L1_20241107.zip"
# 
# # List CSV files inside the zip
# files <- unzip(L1, list = TRUE)$Name
# files <- files[grepl("\\.csv$", files)]
# 
# results <- list()
# 
# for (f in files) {
#   message(f)
# 
#   selected_research_names <- c("wx_tempavg15")
# 
#   # Read CSV directly from zip
#   dat <- read.csv(unz(L1, f))
# 
#   temp <- subset(dat, Instrument == "ClimaVUE50")
# 
#   gw <- temp %>%
#     filter(research_name %in% selected_research_names) %>%
#     select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
# 
#   results[[f]] <- gw
# }
# 
# #we need to bind these together
# dat <- do.call(rbind, results)
# rownames(dat) <- NULL
# dat1 <- dat %>% filter(!(F_OOB == 1 & F_OOS == 1))
# head(dat1)
# 
# #convert to wide format
# temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
# head(temps)
# 
# write.csv(temps, file="Raw Data/L1_temps.csv")

temps <- read.csv("Raw Data/L1_temps.csv")


```

## Match up the L1 temp data and add it to the metadata dataframe
```{r}
#Going to need to use site and timestamp in the metadata in order to find a date 

#first we need to make a timestamp column in metadat so that we can 
metadat$TIMESTAMP <- ymd_hms(paste(metadat$Date, metadat$Time_start))

#make sure time stamps are in the correct format 
metadat$TIMESTAMP <- as.POSIXct(metadat$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")
temps$TIMESTAMP <- as.POSIXct(temps$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")

#time rounding step 
metadat$DateTime_round <- round_date(metadat$TIMESTAMP, "15 mins")
temps$TIMESTAMP_round <- round_date(temps$TIMESTAMP, "15 mins")
  
#ensure there is only one data observation per hour 
temps_clean <-   temps %>% 
    group_by(TIMESTAMP_round, Site) %>%
    summarize(wx_tempavg15 = mean(wx_tempavg15)) 

#join them 
metadat_L1temp <- metadat %>% 
  left_join(temps_clean, by=c("Site"="Site", "DateTime_round"="TIMESTAMP_round"))



```

## Match metadata to raw data
```{r}

#match the metadata to the concentration data 
alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = alldat$TIMESTAMP,
  start_dates = metadat$Date,
  start_times = metadat$Time_start,
  obs_lengths = metadat$Obs_length) 
head(alldat)

#Set plot, temp, volume, etc. here 
alldat$Plot <- metadat$Plot[alldat$metadat_row]
alldat$ID <- metadat$Plot_ID[alldat$metadat_row]
alldat$Volume <- metadat$Volume[alldat$metadat_row]
alldat$Hand_Temp <- metadat$Temp_ByHand[alldat$metadat_row] #found online from weather 
alldat$L1_Temp <- metadat_L1temp$wx_tempavg15[alldat$metadat_row] #calculate  temp for the measurement time from L1
alldat$Site <- metadat$Site[alldat$metadat_row]
metadat$metadat_row <- seq_len(nrow(metadat))

#We need one more line that creates a column for final temp and takes the L1 data but if there isn't L1 data it will pull in another datatype 
# in this case we will use the by hand data i guess? but we could look to using some sort of NOAA dataset.... that we would need to match 

alldat <- alldat %>%
  mutate(final_temp = coalesce(L1_Temp, Hand_Temp))

#remove the rows that don't match to metadata 
#alldat <- alldat %>%
 # filter(!is.na(metadat_row))

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(!is.na(CH4))

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(CH4 < 0)

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(!is.na(CH4))

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(CH4 < 0)

```

## Calculate fluxes 
```{r}
#change the units
alldat$CO2_umol <- ffi_ppm_to_umol(alldat$CO2, 
                                volume = alldat$Volume, # m3
                                temp = alldat$final_temp)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CO2_umol_m2 <- alldat$CO2_umol / 0.16

fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CO2_umol_m2", 
                             dead_band = 10) #, c0_band = 5?
head(fluxes)

#take a look at the fluxes we calculated with R2 
ggplot(fluxes, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CO2 flux (nmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(fluxes, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(fluxes, lin_r.squared < .90, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(fluxes$lin_flux.estimate, 0.05) 
high <- quantile(fluxes$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(fluxes, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(fluxes, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## Spot check a few fluxes with R2 > 0.9
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

####################
```

## Take a look at the bad fluxes <0.9 
# If a flux was started to early/late, change start time or Obs_length in metadata file and note n the note column
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

#Loop to run five random plots subset above 
for(P in bad_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

####################

#Notes on the bad ones 

#	GCW WC 3 202405 – non-linear and wobbly but I think it's okay?  
#	GWI TR 1 202404 – low and okay 
######	GWI TR 1 202405 – Includes something weird at the beginning and needs a look 
#	GWI TR 1 202408 – low and weird, but okay 
#	GWI TR 1 202411 – low and okay 
#	GWI TR 2 202404 – low uptake and okay 
#	GWI TR 2 202405 – low and okay 
######	GWI TR 2 202406 – includes a big jump that looks like chamber opening
#	GWI TR 2 202408 – low and okay 
#	GWI TR 4 202404 – low and okay 
#	GWI TR 4 202405 – low and okay 
#	GWI TR 4 202408 – low and okay 
#	GWI TR 4 202409 – low and okay 
#	GWI TR 4 202411 – low and okay 
#######	GWI WC 1 202406 – includes chamber opening - need to change observation length or start time 
#	GWI WC 1 202408 – low and okay
#	GWI WC 2 202411 – low and okay
# MSM TR 1 202405 - low and okay 
# MSM TR 1 202407 - low and okay 
# MSM TR 1 202410 - low and okay, but looks weird like an M 
# MSM TR 2 202404 - low and okay 
# MSM TR 2 202405 - low and okay 
# MSM TR 2 202409 - low and okay 
# MSM TR 2 202411 - low and okay 
# MSM TR 3 202411 - low and okay 
# MSM TR 4 202411 - low and okay 
# MSM TR 5 202404 - low and okay 
# MSM TR 5 202407 - low and okay 
# MSM TR 5 202409 - low and okay 
# MSM TR 5 202411 - low and okay 
# MSM WC 1 202411 - low and okay 
# MSM WC 4 202411 - low and okay
# MSM WC 5 202411 - low and okay
####### SWH WC 1 202407 - really wonky flux, needs checked 
####### SWH WC 2 202407 - really wonky flux, needs checked 
####### SWH WC 3 202407 - really wonky flux, needs checked 

###########################


```

## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

#Loop to run five random plots subset above 
for(P in tail_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(fluxes, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(fluxes, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(fluxes, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Pull together dataframe and export for use elsewhere 
```{r, echo=FALSE}

#head(fluxes1)

#make new columns with the plot; zone; replicate 
IDs <- data.frame(do.call('rbind', strsplit(as.character(fluxes$Plot),'_',fixed=TRUE)))
colnames(IDs) <- c("Site",  "Zone", "Replicate", "Date") #, "RHZ", "RHZ_Rep")
head(IDs)

#rejoin them to the dataframe
allflux <- cbind(IDs, fluxes)

##Add in notes from metadata
metadat_notes <- metadat[, c("Plot", "Notes")]
allflux_notes <- allflux %>% 
  left_join(metadat_notes, by=c("Plot"="Plot"))

#export as a csv to the folder that you want 
# setwd("S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Fluxes/Macrophyte GHG Fluxes/Final Data")        #Change wd 

write.csv(allflux_notes, file="Processed Data/COMPASS_Synoptic_CB_Macrophyte_Fluxes_Final_2024_CO2.csv")           #Change file name

```

# If you need to look at fluxes from a certain day to see how the concentrations before and after look:
# this doesn't work unless you reload and don't filter out the non matched data because you want to see what is before and after 
# and we take that out in the filter step *** 
```{r, echo=FALSE}

#SWH 2024-07 because there are one or two that look funky  
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
         ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(200, 5250) +
      xlim(c(as.POSIXct('2024-07-24 06:50:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-07-24 08:30:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")

print(p)

#GWI 2024-06 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-06-27 04:55:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-06-27 10:15:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

#GCW 2024-06 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-06-19 02:55:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-06-19 7:15:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

#MSM 2024-06 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-06-18 05:45:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-06-18 8:35:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

#SWH 2024-06 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-06-26 02:45:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-06-26 12:15:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

#GWI 2024-07 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-07-25 05:45:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-07-25 8:15:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

#GWI 2024-10 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-10-23 05:15:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-10-23 10:15:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

#SWH 2024-10 because there are a few that are odd and wonky 
p <- ggplot(alldat, aes(TIMESTAMP, CO2, color = Plot)) + geom_point() + theme_classic() + 
  ylab(expression(paste( CO [2], " (ppmv)"))) + ylim(100, 1500) +
  xlim(c(as.POSIXct('2024-10-24 04:35:00', format = "%Y-%m-%d %H:%M:%S"),
         as.POSIXct('2024-10-24 8:45:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
print(p)

```

