---
title: "Synoptic CB: Macrophyte Fluxes"
author: "2025 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

output_dir: "To Be Reviewed/PDF"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(openxlsx)
```

#Read in and collate all the files in the Processed Data folder 
```{r, include = FALSE}

ch4 <- read.csv("Processed Data/COMPASS_Synoptic_CB_Macrophyte_Fluxes_Final_2025_CH4.csv")
ch4$X <- NULL
ch4 <- ch4 %>%
  rename_with(
    .cols = 7:(ncol(.) - 4),
    .fn = ~paste0("ch4_", .)
  ) %>%
  rename(ch4_qaqc_flag = qaqc_flag)

co2 <- read.csv("Processed Data/COMPASS_Synoptic_CB_Macrophyte_Fluxes_Final_2025_CO2.csv")
co2$X <- NULL
co2 <- co2 %>%
  rename_with(
    .cols = 7:(ncol(.) - 4),
    .fn = ~paste0("co2_", .)
  ) %>%
  rename(co2_qaqc_flag = qaqc_flag)

# Combine all files
all_dat <- merge(ch4, co2, by = c("Plot", "Site", "Zone", "Replicate", "Date", "TIMESTAMP", "TIMESTAMP_min", "TIMESTAMP_max", "Notes"))

##Rename Zone
all_dat$Zone <- gsub("W", "WC", all_dat$Zone)
all_dat$Plot <- gsub("_W_", "_WC_", all_dat$Plot)

```


##Pull in Hydrago probe data
```{r, echo = FALSE}

hydrago <- read.csv("S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Soil Conductivity Probe/Raw Data Files/2025/COMPASS_Synoptic_2025_ALL_SoilCond_Data.csv")

# colnames(hydrago)

## Remove columns that have replicates in flux data
hydrago <- hydrago %>% 
  select(-c("Memo", "Site", "Zone", "Type", "Replicate", "Month", "Month_Name", "Date"))


##Find fluxes that are missing hydrago data: 
all_present <- all(all_dat$Plot %in% hydrago$Plot)

if (all_present) {
  message("All flux IDs are present in hydrago data")
} else {
  message("Some flux IDs are missing from hydrago data")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(all_dat$Plot, hydrago$Plot)
  print(missing_ids)
}

# ##The other way
# ##Find hydrago that are missing fluxes: 
# all_present1 <- all(hydrago$Plot %in% all_dat$Plot)
# 
# if (all_present1) {
#   message("All hydrago IDs are present in flux data")
# } else {
#   message("Some hydrago IDs are missing from flux data")
#   
#   # Optional: Which ones are missing?
#   missing_ids1 <- setdiff(hydrago$Plot, all_dat$Plot)
#   print(missing_ids1)
# }

##Print missing IDs into a dataframe
missing_ids <- data.frame(missing_ids)
write.csv(missing_ids, "fluxes_missing_hydrago_data_2025.csv", row.names = FALSE)

##Combine with flux data
final_dat1 <- merge(all_dat, hydrago, by = c("Plot"), all = TRUE)

```


##pull TEROS data for the zone during the time of the fluxes and put it into another column so that we can use that too.
```{r}
# Air temp 

##Read in L1 Sensor Data from S Drive and only save necessary data
# Path to zip file
L1 <- "S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Synoptic L1 Sensor/L1 2025 Sensor Data v2-1.zip"

# List CSV files inside the zip
files <- unzip(L1, list = TRUE)$Name
files <- files[grepl("\\.csv$", files)]

results <- list()

for (f in files) {
  message(f)
# f = "L1 2025 Sensor Data/GWI_2025/GWI_W_20250101-20250630_wx-tempavg15_L1_v2-0.csv"
  selected_research_names <- c("wx-tempavg15")

  # Read CSV directly from zip
  dat <- read.csv(unz(L1, f))

  temp <- subset(dat, Instrument == "ClimaVUE50" | Instrument == "CS EE181")

  gw <- temp %>%
    filter(research_name %in% selected_research_names) %>%
    select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)

  results[[f]] <- gw
}

#we need to bind these together
dat <- do.call(rbind, results)
rownames(dat) <- NULL
dat1 <- dat %>% filter((F_OOB != 1 & F_OOS != 1))
head(dat1)

#convert to wide format
temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
head(temps)

write.csv(temps, file="Raw Data/L1_temps.csv")

temps <- read.csv("Raw Data/L1_temps.csv")

##Read in raw temperature data
files <- dir(path = "Raw Data/", pattern = "_wx-tempavg15.*\\.csv$", full.names = TRUE)
results <- lapply(files, read.csv) 
dat <- do.call(rbind, results)
rownames(dat) <- NULL
dat <- dat %>%
  select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
dat1 <- dat %>% filter((F_OOB != 1 & F_OOS != 1))
head(dat1)

#convert to wide format
temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
head(temps)

```

##Remove bad fluxes and calculate how many there were
```{r, echo=FALSE}
##Remove any samples with "remove" in the qaqc_flag columns
final_dat <- final_dat1 %>% 
  filter(!grepl("remove", ch4_qaqc_flag, ignore.case = TRUE)) %>% 
  filter(!grepl("remove", co2_qaqc_flag, ignore.case = TRUE)) 

##Calculate how many bad fluxes in each category
##Percent fluxes removed: 
#Total
remove_dat <- all_dat %>% 
  filter(grepl("remove", ch4_qaqc_flag, ignore.case = TRUE) | grepl("remove", co2_qaqc_flag, ignore.case = TRUE))
perc_removed = nrow(remove_dat)/nrow(all_dat)*100

#Negative CO2
neg_co2 <- all_dat %>% 
  filter(grepl("negative CO2", co2_qaqc_flag, ignore.case = TRUE))
perc_neg_co2 = nrow(neg_co2)/nrow(all_dat)*100

#Unusually high CH4/CO2 values
unusual <- all_dat %>% 
  filter(grepl("Unusual", ch4_qaqc_flag, ignore.case = TRUE))
perc_unusual = nrow(unusual)/nrow(all_dat)*100

#Poor/messy flux
messy <- all_dat %>% 
  filter(grepl("messy", ch4_qaqc_flag, ignore.case = TRUE) | grepl("poor", co2_qaqc_flag, ignore.case = TRUE))
perc_messy = nrow(messy)/nrow(all_dat)*100

##Table with  percentage and type of issue

cat(paste0("Number flux measurements removed: ", nrow(remove_dat)))
cat(paste0("Number negative CO2 fluxes removed: ", nrow(neg_co2)))
cat(paste0("Number unusually high starting value fluxes removed: ", nrow(unusual)))
cat(paste0("Number messy fluxes removed: ", nrow(messy)))

cat(paste0("Percentage flux measurements removed: ", round(perc_removed, digits = 2), "%"))
cat(paste0("Percentage negative CO2 fluxes removed: ", round(perc_neg_co2, digits = 2), "%"))
cat(paste0("Percentage unusually high starting value fluxes removed: ", round(perc_unusual, digits = 2), "%"))
cat(paste0("Percentage messy fluxes removed: ", round(perc_messy, digits = 2), "%"))


```

## Write out files
```{r write out combined data and metadata, include=FALSE}

#write out a csv of all the data to the main folder: 
write.csv(final_dat, "COMPASS_SynopticCB_MC_Fluxes_2025.csv")

```


## All Data by Year, Site and Zone
```{r Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting <- final_dat

#Order by Zone from Upland to Surface Water
data_plotting$Zone = factor(data_plotting$Zone, levels = c("UP", "SWAMP", "TR", "WC", "SW"))
data_plotting <- data_plotting[order(data_plotting$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting$Site = factor(data_plotting$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting <- data_plotting[order(data_plotting$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_all <- data_plotting %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

# data_plotting_all$Date <- as.Date(paste(data_plotting_all$Year, data_plotting_all$Month, data_plotting_all$Day, sep = "-"))


#Plot data and change colors based on Zone:
viz_ch4_plot <- ggplot(data_plotting_all, aes(x = row_num, y = ch4_lin_flux.estimate, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "CH4", title = "CO2 Fluxes by Site") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_blank())

 viz_ch4_plot
 
 viz_co2_plot <- ggplot(data_plotting_all, aes(x = row_num, y = co2_lin_flux.estimate, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "CO2", title = "CO2 Fluxes by Site") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_blank())

 viz_co2_plot


```

## Boxplot summary by Date
```{r Overall Boxplot by Date, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_all <- data_plotting_all %>% 
    mutate(
    Zone_1 = case_when(
      Zone == "UP" ~ "Upland",
      Zone == "SWAMP" ~ "Swamp",
      Zone == "TR" ~ "Transition",
      Zone == "WC" ~ "Wetland",
      Zone == "SW" ~ "Surface Water",
      TRUE ~ Zone  # keep everything else unchanged
    )
  ) %>%
  mutate(Zone_1 = factor(Zone_1, levels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")))


##Summarize NPOC by Date, Site, and Zone
viz_ch4_boxplot <- ggplot(data_plotting_all, aes(x = as.character(Date), y = ch4_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 2) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CH4", title = "Summary Boxplot of CH4 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_ch4_boxplot


##Summarize TDN by Date, Site, and Zone
viz_co2_boxplot <- ggplot(data_plotting_all, aes(x = as.character(Date), y = co2_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 2) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CO2", title = "Summary Boxplot of CO2 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_co2_boxplot

```

## Boxplot summary by Site
```{r Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Summarize NPOC by Date, Site, and Zone
viz_ch4_summary_boxplot <- ggplot(data_plotting_all, aes(x = Zone_1, y = ch4_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 1) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CH4", title = "Summary Boxplot of CH4 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_ch4_summary_boxplot


##Summarize TDN by Date, Site, and Zone
viz_co2_summary_boxplot <- ggplot(data_plotting_all, aes(x = Zone_1, y = co2_lin_flux.estimate, fill = Zone_1)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales = "free", nrow = 1) +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  ), labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")) +
  theme_classic() +
  labs(x = " ", y = "CO2", title = "Summary Boxplot of CO2 by Site and Zone") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  #scale_x_discrete(labels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water"))+ 

viz_co2_summary_boxplot

```

##Calculate summary metrics
```{r Additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add to our collation codes a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, zone, depth for each year; also for each site zone, depth overall; and for each site and zone overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

##Metrics for each site and zone for all depths and years
summary <- data_plotting_all %>%
  group_by(Site, Zone_1) %>%
   reframe(
     min_ch4 = min(ch4_lin_flux.estimate, na.rm=TRUE),
     max_ch4 = max(ch4_lin_flux.estimate, na.rm=TRUE), 
     mean_ch4 = mean(ch4_lin_flux.estimate, na.rm=TRUE), 
     sd_ch4 = sd(ch4_lin_flux.estimate, na.rm=TRUE), 

     min_co2 = min(co2_lin_flux.estimate, na.rm=TRUE),
     max_co2 = max(co2_lin_flux.estimate, na.rm=TRUE), 
     mean_co2 = mean(co2_lin_flux.estimate, na.rm=TRUE), 
     sd_co2 = sd(co2_lin_flux.estimate, na.rm=TRUE))

```

## Summary Tables
```{r Summary table metrics, echo = FALSE, message = FALSE, dpi = 100}

library(kableExtra)

kable(
  summary[, c("Site", "Zone_1", "min_ch4",	"max_ch4",	"mean_ch4", "min_co2",	"max_co2",	"mean_co2")],  
  format = "latex",
  booktabs = TRUE,
  caption = "Summary Statistics"
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)


```
