---
title: "COMPASS_Synoptic_MacrophyteFluxes_2025_CH4"
author: "Stephanie J. Wilson"
date: "2025-11-26"
output: pdf_document
---
## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
```

## Pull in Raw Datafiles
```{r}
##Read in 2026 Louise data and only save necessary data (9/29)
#read in files
# file.list.Louise <- dir(path = "S:/Biogeochemistry/LICOR/Louise Data Backup 20260121", pattern = "data$", full.names = TRUE)
# head(file.list.Louise)
# 
# # #Read in the data 
# alldat1.Louise <- lapply(file.list.Louise, ffi_read_LI7810)
# alldat.Louise <- data.table::rbindlist(alldat1.Louise, fill=TRUE)
# 
# #Clip data to only 9/29/2025
# # Define the start and end of the specific date
# start_datetime <- as.POSIXct("2025-09-29 00:00:00")
# end_datetime <- as.POSIXct("2025-09-29 23:59:59") # Or the start of the next day
# 
# alldat.Louise.Sept <- alldat.Louise %>%
#   filter(TIMESTAMP >= start_datetime & TIMESTAMP <= end_datetime)
# 
# #Export Louise data already read in - can read in later and just cbind it on 
# write.csv(alldat.Louise.Sept, file="Raw Data/Louise_20250929_preprocessed.csv", row.names = FALSE)

# alldat$metadat_row <- as.numeric(alldat$metadat_row)
# sort(unique(alldat$metadat_row))
##We were missing 16-25 (Gru 5/13 data) and 121-125 (9/29 Louise) and 131-166 (11/4-11/13 Simon)

#Read in the data only for November - data was regustered on the Licor as UTC when it was still US/Eastern time. Need to force the data to be recognized as US/Eastern Time to line up with metadata because otherwise the data will be forced into UTC and times will be 7 hours off. 
file.list.nov <- dir(path = "Raw Data/", pattern = "(202511.*\\.txt$)", full.names = TRUE)
file.list.nov
alldat1.nov <- lapply(file.list.nov, ffi_read_LI7810)
alldat.raw.nov <- data.table::rbindlist(alldat1.nov, fill=TRUE)

#Convert to US/Eastern TZ
alldat.raw.nov$TZ = "US/Eastern"
alldat.raw.nov$TIMESTAMP <- with_tz(alldat.raw.nov$TIMESTAMP, tz = "UTC")
alldat.raw.nov$TIMESTAMP <- force_tz(alldat.raw.nov$TIMESTAMP, "US/Eastern")


# read in all other files
# read in files
file.list <- dir(path = "Raw Data/", pattern = "txt$", full.names = TRUE)
head(file.list)
file.list <- file.list[!file.list %in% file.list.nov]
file.list

#Read in the data
alldat1 <- lapply(file.list, ffi_read_LI7810)
alldat.raw <- data.table::rbindlist(alldat1, fill=TRUE)

##Read in Louise data from 9/29
Louise.Sept <- read.csv("Raw Data/Louise_20250929_preprocessed.csv")
Louise.Sept$TIMESTAMP <- as.POSIXct(Louise.Sept$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")

##Combine all other data with Simon's November 2025 and Louise's Sept 29 2025 data
alldat <- rbind(alldat.raw, alldat.raw.nov, Louise.Sept)

```

## Read in metadata file and match to raw data
```{r}

#read in metadatfile
md <- "Raw Data/COMPASS_ChamberFlux_Metadata_2025a.csv"
#metadat <- read_excel(md)
metadat <- read.csv(md) 

head(metadat)  

#can remove any row that have NA's in the date column 
metadat <- metadat %>%
  filter(!is.na(Date))

head(metadat) 

```

## Read in Air Temp Data from Air Temp Code # use after L1 release
```{r}
# Air temp 

#setwd("S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Fluxes/Macrophyte GHG Fluxes/Flux Calculations/Air Temp Data")
#download the L1 dataset from: https://drive.google.com/drive/folders/1RP-hxmjAYNGzTfCNpYbMN6U_zbZmEWHM

##Read in L1 Sensor Data from S Drive and only save necessary data
# Path to zip file
# L1 <- "S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Synoptic L1 Sensor/L1 2025 Sensor Data.zip"
# 
# # List CSV files inside the zip
# files <- unzip(L1, list = TRUE)$Name
# files <- files[grepl("\\.csv$", files)]
# 
# results <- list()
# 
# for (f in files) {
#   message(f)
# # f = "L1 2025 Sensor Data/GWI_2025/GWI_W_20250101-20250630_wx-tempavg15_L1_v2-0.csv"
#   selected_research_names <- c("wx-tempavg15")
# 
#   # Read CSV directly from zip
#   dat <- read.csv(unz(L1, f))
# 
#   temp <- subset(dat, Instrument == "ClimaVUE50" | Instrument == "CS EE181")
# 
#   gw <- temp %>%
#     filter(research_name %in% selected_research_names) %>%
#     select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
# 
#   results[[f]] <- gw
# }
# 
# #we need to bind these together
# dat <- do.call(rbind, results)
# rownames(dat) <- NULL
# dat1 <- dat %>% filter((F_OOB != 1 & F_OOS != 1))
# head(dat1)
# 
# #convert to wide format
# temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
# head(temps)
# 
# write.csv(temps, file="Raw Data/L1_temps.csv")
# 
# temps <- read.csv("Raw Data/L1_temps.csv")

##Read in raw temperature data
files <- dir(path = "Raw Data/", pattern = "_wx-tempavg15.*\\.csv$", full.names = TRUE)
results <- lapply(files, read.csv) 
dat <- do.call(rbind, results)
rownames(dat) <- NULL
dat <- dat %>%
  select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
dat1 <- dat %>% filter((F_OOB != 1 & F_OOS != 1))
head(dat1)

#convert to wide format
temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
head(temps)

```

## Match up the L1 temp data and add it to the metadata dataframe
```{r}
#Going to need to use site and timestamp in the metadata in order to find a date 

#first we need to make a timestamp column in metadat so that we can 
metadat$TIMESTAMP <- mdy_hms(paste(metadat$Date, metadat$Time_start))

#make sure time stamps are in the correct format 
metadat$TIMESTAMP <- as.POSIXct(metadat$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")
temps$TIMESTAMP <- as.POSIXct(temps$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")

#time rounding step 
metadat$DateTime_round <- round_date(metadat$TIMESTAMP, "15 mins")
temps$TIMESTAMP_round <- round_date(temps$TIMESTAMP, "15 mins")
  
#ensure there is only one data observation per hour 
temps$`wx_tempavg15` <- temps$`wx-tempavg15`
temps_clean <- temps %>% 
    group_by(TIMESTAMP_round, Site) %>%
    summarize(wx_tempavg15 = mean(wx_tempavg15))

#join them 
metadat_L1temp <- metadat %>% 
  left_join(temps_clean, by=c("Site"="Site", "DateTime_round"="TIMESTAMP_round"))



```

## Match metadata to raw data
```{r}

#match the metadata to the concentration data 
alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = alldat$TIMESTAMP,
  start_dates = metadat$Date,
  start_times = metadat$Time_start,
  obs_lengths = metadat$Obs_length) 
head(alldat)


#Set plot, temp, volume, etc. here 
alldat$Plot <- metadat$Plot[alldat$metadat_row]
alldat$ID <- metadat$Plot1[alldat$metadat_row]
alldat$Volume <- metadat$Volume[alldat$metadat_row]
alldat$Hand_Temp <- metadat$Temp_ByHand[alldat$metadat_row] #found online from weather 
alldat$L1_Temp <- metadat_L1temp$wx_tempavg15[alldat$metadat_row] #calculate  temp for the measurement time from L1
alldat$Site <- metadat$Site[alldat$metadat_row]
metadat$metadat_row <- seq_len(nrow(metadat))

#We need one more line that creates a column for final temp and takes the L1 data but if there isn't L1 data it will pull in another datatype 
# in this case we will use the by hand data i guess? but we could look to using some sort of NOAA dataset.... that we would need to match 

alldat <- alldat %>%
  mutate(final_temp = coalesce(L1_Temp, Hand_Temp))

unique(alldat$Plot)

#remove the rows that don't match to metadata 
#alldat <- alldat %>%
 # filter(!is.na(metadat_row))

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(!is.na(CH4))

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(CH4 < 0)

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(!is.na(CH4))

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(CH4 < 0)

```

## Calculate fluxes 
```{r}
#change the units
alldat$CH4_nmol <- ffi_ppb_to_nmol(alldat$CH4, 
                                volume = alldat$Volume, # m3
                                temp = alldat$final_temp)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CH4_nmol_m2 <- alldat$CH4_nmol / 0.16

fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CH4_nmol_m2", 
                             dead_band = 10) #, c0_band = 5? 
head(fluxes)

#take a look at the fluxes we calculated with R2 
ggplot(fluxes, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CH4 flux (nmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(fluxes, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(fluxes, lin_r.squared < .90, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(fluxes$lin_flux.estimate, 0.05) 
high <- quantile(fluxes$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(fluxes, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(fluxes, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## Check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r}

tobechecked <- c("")

colnames <- c("Plot")

#Loop to run plots above 
for(P in tobechecked){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#list of the ones that I changed and checked and looked good? 
fixedplots <- c()
  

```


## Spot check a few fluxes with R2 > 0.9
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }


```

## Take a look at the bad fluxes <0.9 
# If a flux was started too early/late, change start time or Obs_length in metadata file and note n the note column
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

#Loop to run five random plots subset above 
for(P in bad_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }


#Notes on the bad ones 

##Fluxes to change: 
# GCW_TR_4_202509 - added 5 sec to start time to avoid chamber closure
# MSM_TR_3_202509 - reduced obs length to 4 minutes to avoid ebullition 
# SWH_TR_1_202507 - changed obs length to 3 minutes to avoid ebullition 
# SWH_TR_3_202507 - changed obs length to 4 minutes to avoid ebullition 
# SWH_TR_4_202509 - increase start time to 10:43:30 to avoid ebullition
# SWH_TR_5_202509 - reduced obs length to 3 minutes to avoid ebullition 
# SWH_TR_3_202509 - reduced obs length to 4 minutes to avoid ebullition 
# SWH_TR_4_202507 - increase start time by 30 seconds
# GCW_TR_2_202511 - reduce obs length to 3 minutes to avoid ebullition
# MSM_TR_3_202511 - increase start time to 15:10:30
# SWH_TR_3_202507 - increase start time to 15:03
# MSM_TR_5_202511 - increase start time to 15:28, reduce obs length to 4 minutes
# MSM_W_2_202509 - increase start time to 13:08, reduce obs length to 2.5 minutes -***still need to reduce obs length to 2 minutes!

#SWH_TR_3_202509 - asked Steph, she said to keep it

##Fluxes to remove
# SWH_TR_5_202507 - extremely high CH4 values (450,000 ppb), remove 
# SWH_TR_4_202507 - extremely high CH4 values (150,000 ppb), remove 

# MSM_TR_4_202509 - flux compromised by ebullition, remove
# GCW_TR_5_202509 - ebullition, poor flux
# GCW_W_1_202509 - ebullition, poor flux
# MSM_TR_3_202509 - ebullition, poor flux
# SWH_TR_2_202509 - ebullition, poor flux
# SWH_TR_4_202509 - ebullition, poor flux
# SWH_TR_5_202509 - poor flux
# SWH_TR_1_202509 - ebullition, poor flux
# GCW_TR_1_202509 - poor flux
# GCW_TR_2_202509 - poor flux
# GCW_TR_5_202507 - ebullition, poor flux
# SWH_TR_3_202509 - remove, poor flux

##Added notes to metadata for the ones I fixed

####Add qaqc flags####

fluxes$qaqc_flag <- NA

##Fluxes with crazy high CH4 values
high_CH4_flags <- c(
  "SWH_TR_5_202507", 
  "SWH_TR_4_202507")

fluxes <- fluxes %>%
    mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% high_CH4_flags, "Unusually high CH4 values, remove", qaqc_flag))

#Fluxes that need to be removed
remove_flags <- c(
  "MSM_TR_4_202509", 
  "GCW_TR_5_202509", 
  "GCW_W_1_202509", 
  "MSM_TR_3_202509", 
  "SWH_TR_2_202509", 
  "SWH_TR_4_202509", 
  "SWH_TR_5_202509", 
  "SWH_TR_1_202509", 
  "GCW_TR_1_202509", 
  "GCW_TR_2_202509", 
  "GCW_TR_5_202507")

fluxes <- fluxes %>%
    mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% remove_flags, "messy flux, remove", qaqc_flag))

#Any other fluxes that had low R squared but are okay otherwise
fluxes <- fluxes %>%
    mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% bad_plots, "low r squared, but checked and flux is okay", qaqc_flag))



```

## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

#Loop to run five random plots subset above 
for(P in tail_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################
# SWH_TR_5_202507


```

## Plot all fluxes
```{r, echo=FALSE}

#ffi_qaqc(fluxes, Plot, output_file = "qaqc.html", output_dir = getwd(), open_output = TRUE)

### pull a random sample of good fluxes to check and plot 
all_plots <- unique(fluxes$Plot)

#Loop to for all plots
for(P in all_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################


```

## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(fluxes, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(fluxes, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(fluxes, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Pull together dataframe and export for use elsewhere 
```{r, echo=FALSE}

#head(fluxes1)

#make new columns with the plot; zone; replicate 
IDs <- data.frame(do.call('rbind', strsplit(as.character(fluxes$Plot),'_',fixed=TRUE)))
colnames(IDs) <- c("Site",  "Zone", "Replicate", "Date") #, "RHZ", "RHZ_Rep")
head(IDs)

#rejoin them to the dataframe
allflux <- cbind(IDs, fluxes)

##Add in notes from metadata
metadat_notes <- metadat[, c("Plot", "Notes")]
allflux_notes <- allflux %>% 
  left_join(metadat_notes, by=c("Plot"="Plot"))

#export as a csv to the folder that you want 
# setwd("S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Fluxes/Macrophyte GHG Fluxes/Final Data")        #Change wd 

write.csv(allflux_notes, file="Processed Data/COMPASS_Synoptic_CB_Macrophyte_Fluxes_Final_2025_CH4.csv")           #Change file name


```

##Histogram of CH4 concentrations by Site
```{r}

alldat_plot <- subset(alldat, !is.na(Site))
alldat_plot <- subset(alldat_plot, Plot != "SWH_TR_5_202507")
alldat_plot <- subset(alldat_plot, Plot != "SWH_TR_4_202507")

hist <- ggplot(alldat_plot, aes(x = CH4)) +
  geom_histogram(bins = 50, color = "black", fill = "skyblue", na.rm = TRUE) +
  facet_wrap(~ Site, scales = "free") + # Creates separate plots for each group
  labs(title = "Methane spans by site for 2025", x = "CH4 Conc (ppb)", y = "Frequency")+ 
  theme_classic()
hist

ggsave("CH4_span.png", 
       plot = hist, 
       width = 8,
       height = 4,
       units = "in",
       dpi = 300)

##Do the same thing but average CH4 per flux measurement taken
#ensure there is only one data observation per flux measurement 
alldat_plot_means <- alldat_plot %>% 
    group_by(Plot, Site) %>%
    summarize(CH4_avg = mean(CH4)) 


hist_avgs <- ggplot(alldat_plot_means, aes(x = CH4_avg)) +
  geom_histogram(bins = 50, color = "black", fill = "skyblue", na.rm = TRUE) +
  facet_wrap(~ Site, scales = "free") + # Creates separate plots for each group
  labs(title = "Avg methane per flux measurement by site for 2025", x = "CH4 Conc (ppb)", y = "Frequency")+ 
  theme_classic()
hist_avgs

ggsave("CH4_span_avgs.png", 
       plot = hist_avgs, 
       width = 8,
       height = 4,
       units = "in",
       dpi = 300)

```



# If you need to look at fluxes from a certain day to see how the concentrations before and after look:
# this doesn't work unless you reload and don't filter out the non matched data because you want to see what is before and after 
# and we take that out in the filter step *** 
```{r, echo=FALSE}

# #GWI 2025-05 because there are one or two that look funky  
# p <- ggplot(alldat, aes(TIMESTAMP, CH4, color = Plot)) + geom_point() + theme_classic() + 
#          ylab(expression(paste( CH [4], " (ppmv)"))) + ylim(1000, 70250) +
#       xlim(c(as.POSIXct('2025-07-23 10:00:00', format = "%Y-%m-%d %H:%M:%S"),
#          as.POSIXct('2025-07-23 12:30:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
# 
# print(p)
# 
# #GWI 2025-06 because there are a few that are odd and wonky 
# p <- ggplot(alldat, aes(TIMESTAMP, CH4, color = Plot)) + geom_point() + theme_classic() + 
#   ylab(expression(paste( CH [4], " (ppmv)"))) + #ylim(1800, 3000) +
#   xlim(c(as.POSIXct('2025-06-27 08:45:00', format = "%Y-%m-%d %H:%M:%S"),
#          as.POSIXct('2025-06-27 13:40:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
# print(p)
# 
# #SWH
# p <- ggplot(alldat, aes(TIMESTAMP, CH4, color = Plot)) + geom_point() + theme_classic() + 
#   ylab(expression(paste( CH [4], " (ppmv)"))) + ylim(-1000,50000) +
#   xlim(c(as.POSIXct('2023-11-08 12:30:00', format = "%Y-%m-%d %H:%M:%S"),
#          as.POSIXct('2023-11-08 14:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme(legend.position="none")
# print(p)
# 
# #GWI
# p <- ggplot(alldat, aes(TIMESTAMP, CH4, color = Plot)) + geom_point() + theme_classic() + 
#   ylab(expression(paste( CH [4], " (ppmv)"))) +  ylim(2050, 2500) +
#   xlim(c(as.POSIXct('2023-11-09 09:40:00', format = "%Y-%m-%d %H:%M:%S"),
#          as.POSIXct('2023-11-09 12:40:00', format = "%Y-%m-%d %H:%M:%S")))+ theme(legend.position="none")
# print(p)
```

