---
title: "Synoptic CB: Manual Redox Probes"
author: "2023 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
  toc: true
number_sections: true

output_dir: "To Be Reviewed/PDF"
---
  
# Measurements taken with SWAP Redox Probes & Calomel Ref probe
# Protocol: https://docs.google.com/document/d/
# 1dL2ZPhLFBAZJmVBCfKxup_KOYY96qA5IWfJmQen50pY/edit
# Units = mV 

##Add Required Packages
```{r packages, include = FALSE}

#Packages that are required 
lapply(c(
  "dplyr", "ggplot2", "ggpubr", "stringr",
  "purrr", "tidyverse", "here", "broom", "tibble",
  "googledrive", "googlesheets4", "data.table", 
  "matrixStats", "gridExtra", "grid", "tidyverse", "knitr"), 
  library, character.only = TRUE)

```

##Setup
```{r setup}

###### File Names - PLEASE CHANGE 
#file path and name for raw summary data file 
dat <- read.csv("Raw Data/COMPASS_Synoptic_CB_Redox_2023_Raw.csv")
soildat <- read.csv("Raw Data/CB_Soil_pH.csv")

#file path and name of processed data file 
processed_file_name = "Processed Data/COMPASS_Synoptic_CB_Redox_2023_processed.csv" 

# ##Fix Sample Zone and Site naming
# soildat <- soildat %>%
#   mutate(Zone = case_when(Site == "SWH" & Zone == "Upland" ~ "Swamp", 
#                           Site == "SWH" & Zone == "Upland-Control" ~ "Upland",
#                           TRUE ~ Zone), 
#          Site = ifelse(Site == "Gcrew", "GCW", Site))

##Fix Site naming
dat <- dat %>%
  mutate(Zone = case_when(Site == "SWH" & Zone == "Upland" ~ "Swamp", 
                          Site == "SWH" & Zone == "Upland Control" ~ "Upland",
                          Site == "SWH" & Zone == "Upland-Control" ~ "Upland",
                          TRUE ~ Zone), 
         Site = ifelse(Site == "Gcrew", "GCW", Site), 
         Month = ifelse(Month == "Arpil", "April", Month))

dat$Site <- gsub(" ", "", dat$Site) ## Remove space in GWI site name

##Add column for SWH topography
dat <- dat %>%
  mutate(Topography = case_when(Site == "SWH" & Zone == "Transition-Hummock" ~ "Hummock", 
                                Site == "SWH" & Zone == "Transition-Hollow" ~ "Hollow", 
                                TRUE ~ NA), 
         Zone = case_when(Site == "SWH" & Zone == "Transition-Hummock" ~ "Transition",
                          Site == "SWH" & Zone == "Transition-Hollow" ~ "Transition", 
                          TRUE ~ Zone))

##Check redox depths of 30 cm
dat30 <- dat %>% 
  filter(Depth == 30)

##Std error function
std.error <- function (x, na.rm=FALSE){
  sqrt(var(x, na.rm = na.rm) / sum(!is.na(x)))}

```

## Adjust Raw Readings by pH
## Need to adjust for pH: 
# redox = redoxi + [(ph - 5)59 + 224] This is the equation from Megonigal, 
# Patrick, & Faulkner 1993
# They used a calomel electrode and used 224 to adjust to H2, 
# we use a KCl ref probe, so we need to add 202 instead 
```{r adjust for pH, message = FALSE, warning = FALSE}

##Remove unneccesary column in redox data
prelong <- dat %>% 
  select (-Campaign) %>% 
  mutate(Notes = NA, 
         Time_24hr = NA, 
         Time_Zone = NA)

#switch redox data to long format 
datlong <-  as.data.frame(melt(setDT(prelong), id.vars = c("Date", "Month", "Year", "Site", "Zone", "Depth", "Time_24hr", "Time_Zone", "Topography", "Notes"), 
                               variable.name = "RP_mV"))

# ##Add in soil pH data
# all_corr <- merge(datlong, soildat, by = c("Site", "Zone"), all = TRUE)
# 
# ##Adjust redox for pH
# all_corr <- all_corr %>% 
#   mutate(value_corr = value + ((pH - 5)*59 + 202))

##Average across all 5-6 measurements for each sampling Site, Zone, Depth, Month, Year
red1 <- datlong %>%
  group_by(Date, Month, Year, Site, Zone, Depth, Time_24hr, Time_Zone, Topography, Notes) %>%
  summarise(Date = mean(Date), 
            Avg_mV = mean(value, na.rm=TRUE), 
            Std_dv = sd(value, na.rm=TRUE), 
            Std_err = std.error(value, na.rm=TRUE), 
            Notes = first(Notes)
            )


# head(red1)

```

##Write Out Adjusted & Avg'd Data
```{r write out processed data, include = FALSE}

#write a .csv of the new data frame 
write.csv(red1, processed_file_name, row.names = FALSE) 

```

## Vizualize Data
```{r visualize data, echo = FALSE, warning = FALSE, message = FALSE}

#Plot samples to get a first look at redox (sanity check)
data_plotting <- red1 %>% 
  mutate(Zone = case_when(Site == "SWH" & Zone == "Transition" & Topography == "Hummock" ~ "Transition-Hummock",
                          Site == "SWH" & Zone == "Transition" & Topography == "Hollow" ~ "Transition-Hollow",
                          TRUE ~ Zone))

#Order by Zone from Upland to Surface Water
data_plotting$Zone = factor(data_plotting$Zone, levels = c("Upland",  "Swamp", "Transition", "Transition-Hummock", "Transition-Hollow", "Wetland"))
data_plotting <- data_plotting[order(data_plotting$Zone), ]

#group the data for plotting
data_plotting <- data_plotting %>%
  group_by(Site) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

#Plot all data and change colors based on zone:
viz_redox_plot <- ggplot(data_plotting, aes(x = row_num, y = Avg_mV, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
  facet_grid(~ Site, scales="free_x") +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Swamp" = "darkgrey",
    "Transition" = "#FFBC42",
    "Transition-Hollow" = "darkorange",
    "Transition-Hummock" = "#FFBC42",
    "Wetland" = "#419973"
  )) +
  theme_classic() +
  theme(axis.text.x = element_blank(), 
        legend.position = "bottom") +
  labs(x = " ", y = "Redox Potential (mV)", title = "All Redox Data") +
  scale_x_discrete(drop = TRUE)

viz_redox_plot

#Plot averaged data by depth and change colors based on zone:
red <- datlong %>%
  mutate(Zone = case_when(Site == "SWH" & Zone == "Transition" & Topography == "Hummock" ~ "Transition-Hummock",
                          Site == "SWH" & Zone == "Transition" & Topography == "Hollow" ~ "Transition-Hollow",
                          TRUE ~ Zone)) %>% 
  group_by(Site, Zone, Depth) %>%
  summarise(Date = mean(Date), Avg_mV = mean(value, na.rm=TRUE), Std_dv = sd(value, na.rm=TRUE), Std_err = std.error(value, na.rm=TRUE))

viz_redox_depth <- ggplot(red, aes(x=Depth, y= Avg_mV, col=Zone, fill=Zone))+
  geom_point(size=3) +  
  geom_line(linewidth=1.5) +
  scale_x_reverse() + 
  # ylim(c(-150, 700)) +
  coord_flip() + 
  theme_classic() +
  facet_grid(~factor(Site, levels=c("MSM", "GWI", "GCW", "SWH"))) +
  labs(x="Depth (cm)", y="Redox Potential (mV)", title ="Averaged Redox Data") +
  scale_colour_manual(values = c(
    "Upland" = "#20063B",
    "Swamp" = "darkgrey",
    "Transition" = "#FFBC42",
    "Transition-Hollow" = "darkorange",
    "Transition-Hummock" = "#FFBC42",
    "Wetland" = "#419973"
  )) +
  theme(panel.spacing = unit(.5, "lines"),
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        strip.background = element_rect(color = "black", size = 1),
        plot.margin = margin(t=0.1, r=0.1, l=0.1, b=0.1, "cm"),
        plot.title = element_text(hjust = 0.0),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.text=element_text(size=12), 
        legend.title=element_blank(),
        legend.position = "bottom")+
  geom_errorbar(aes(ymin=Avg_mV-Std_err, ymax=Avg_mV+Std_err),width=0.3,position=position_dodge(.1))

viz_redox_depth

```


