---
title: "COMPASS_SynopticCB_Collate_Iron_2024_Data"
author: "Stephanie J. Wilson and Zoe Read"
date: "2025-12-04"
output: html_document
---

##Run Notes
```{r}
run_notes = "Sample IDs for May 2024 MSM seem to have already been fixed, since there is no MSM_202405_UP_LysA_10cm sample ID."
```


##Read in packages
```{r packages, include = FALSE}
#packages
library(ggplot2)
library(dplyr)
library(data.table)
library(matrixStats)
library(gridExtra)
library(ggpubr)
library(grid)
library(stringr)
library(plater)
library(raster)
library(tidyverse)
library(readxl)
library(dplyr)

```

#Read in and collate all the files in the Processed Data folder 
```{r}
files <- list.files(path = "QAQC'd Data", pattern = "\\_shortdata.csv$", full.names = TRUE, recursive =  TRUE)

read_csv_filename <- function(filename) {
  # Read the CSV file
  data <- read.csv(filename)
  # Add a new column 'SourceFile' with only the filename (using basename)
  data$SourceFile <- basename(filename)
  # Return the modified data frame
  return(data)
}

# 3. Use lapply to apply the function to all files
list_of_dfs <- lapply(files, read_csv_filename)

# 4. Combine the list of data frames into a single data frame
#    This uses base R's do.call and rbind
all_data <- do.call(rbind, list_of_dfs)

##Extract Analysis date from file name
all_data <- all_data %>%
  mutate(Analysis_rundate = str_extract(SourceFile, "\\d{8}"))

all_data <- all_data %>%
  mutate(Plate = str_extract(SourceFile, "Plate[0-9]+"))

all_data <- all_data[, -which(names(all_data) == "X")]
all_data <- all_data[, -which(names(all_data) == "SourceFile")]

head(all_data)
```

##Read in and Fix Sample IDs
```{r Read in and Fix Sample IDs, include = FALSE}
#Read in Sample IDs
SampleIDs <- read.csv("COMPASS_CBSyn_PW_Fe_2024_SampleList.csv", skip = 1)
```

##Convert Sample Numbers to Sample IDs
```{r Convert Sample Numbers to Sample IDs, include = FALSE}
##Merge with data
Sample_data_IDs <- all_data %>%
  left_join(SampleIDs, by = c("IDs" = "Sample.No."))

head(Sample_data_IDs)

#Remove Part of Sample ID we don't need
Sample_data_IDs$Sample.ID <- gsub("_PW_MonMon_FE_", "_", Sample_data_IDs$Sample.ID)
Sample_data_IDs$Sample.ID <- gsub("_MonMon_FE_", "_", Sample_data_IDs$Sample.ID)

head(Sample_data_IDs)

##Separate sample IDs 
Sample_data_IDs <- Sample_data_IDs %>%
    separate(
    col = "Sample.ID",
    sep = "_",
    into = c("Site", "Zone", "Date", "Replicate", "Depth"),
    remove = FALSE) 

Sample_data_IDs$`Sample.ID` <- NULL

##Make full sample IDs
##Depths for SW samples should be 0 
Sample_data_IDs$Depth <-  ifelse((Sample_data_IDs$Zone == "SW"), "0cm", Sample_data_IDs$Depth)

#Create IDs from what was collected for comparison later
Sample_data_IDs <- Sample_data_IDs %>%
  mutate(Sample_ID = paste(Site, Date, Zone, Replicate, Depth, sep = "_"))

#Change the SW lines because they don't have lysimeters or a depth  
Sample_data_IDs <- Sample_data_IDs %>%
  mutate(
    Sample_ID = if_else(
      Zone == "SW",
      # Modify the string:
      Sample_ID %>%
        str_replace("_LysA", "_A") %>%                    # Replace "_LysA" with "_A"
        str_replace("_LysB", "_B") %>%                    # Replace "_LysB" with "_B"
        str_replace("_LysC", "_C") %>%                    # Replace "_LysC" with "_C"
        str_replace("_0cm$", ""),                          # Remove trailing "_0cm"
      Sample_ID  # else keep original
    )
  )

#Remove Lys from Replicate Column
Sample_data_IDs <- Sample_data_IDs %>%
  mutate(Replicate = gsub("Lys", "", as.character(Replicate)))%>%
  mutate(Replicate = gsub("lys", "", as.character(Replicate)))

```


#Note from Zoe and Steph: It looks like we did not get any samples for MSM_202405_UP_LysA_10cm. According to the corrections on the scanned datasheets, it looks like for Upland, B samples were accidentally put in the A vials, C samples were put in B vials, and A samples were put in C vials. So, I'm thinking I can check if any analytes have MSM_202405_UP_LysA_10cm, then all the samples need to be corrected A -->B, B --> C and C-->A. 
```{r Check and fix MSM Up Lys sample IDs, echo = FALSE}

string_to_find <- "MSM_202405_UP_LysA_10cm"
found <- any(sapply(Sample_data_IDs, function(x) any(grepl(string_to_find, x))))

# Print the result
if (found) {
  print(paste("The string '", string_to_find, "' was found in the dataframe.", sep = ""))
} else {
  print(paste("The string '", string_to_find, "' was not found in the dataframe.", sep = ""))
}

```

##Pull in Metadata
```{r Pull in Metadata, include = FALSE}

Raw_Metadata = "COMPASS_SynopticCB_PW_SampleLog_2024.csv"

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata)

#make a new columns in the metadata with important info:
metadata <- raw_metadata %>%
  mutate(Depth = paste0(Depth_cm, "cm")) %>%
  mutate(LysID = paste0("Lys", Lysimeter)) %>%
  mutate(YearMonth = sprintf("%d%02d", Year, Month))  %>%
  mutate(Zone = case_when(
    `Transect.Location` == "Transition" ~ "TR",
    `Transect.Location` == "Wetland"    ~ "WC",
    `Transect.Location` == "Upland"     ~ "UP",   
    `Transect.Location` == "Surface Water" ~ "SW",
    TRUE                 ~ `Transect.Location`    # keep original value if no match
  ))

#Create IDs from what was collected for comparison later
metadata <- metadata %>%
  mutate(Iron_ID = ifelse(SO4 == "x",
                                     paste(Site,
                                           YearMonth,
                                           Zone,
                                           LysID,
                                           Depth,
                                           sep = "_"),
                                     NA),  )

#Change the SW lines because they don't have lysimeters or a depth  
metadata <- metadata %>%
  mutate(
    Iron_ID = if_else(
      Zone == "SW",
      # Modify the string:
      Iron_ID %>%
        str_replace("_LysA", "_A") %>%                    # Replace "_LysA" with "_A"
        str_replace("_LysB", "_B") %>%                    # Replace "_LysB" with "_B"
        str_replace("_LysC", "_C") %>%                    # Replace "_LysC" with "_C"
        str_replace("_0cm$", ""),                          # Remove trailing "_0cm"
      Iron_ID  # else keep original
    )
  )

#Take out the columns and rows that are not relevant
iron_metadata <- metadata %>%
  dplyr::select(Iron_ID, Year, Month, Day, YearMonth, Site, Zone, Lysimeter, LysID, Depth_cm, 
         Depth, Time..24hr., Time.Zone_EDT.EST, Field.Notes, ) %>%        
  # only keep specific columns
  filter(!is.na(Iron_ID) & Iron_ID != "")  # remove missing/blank Cl_SO4_ID rows

head(iron_metadata)

```

##Check to see if samples run match metadata & merge info
```{r Check to see if samples run match metadata & merge info, echo = FALSE}

#check to see if all samples are present in the metadata 
Sample_data_IDs$Sample_ID <- toupper(Sample_data_IDs$Sample_ID)
iron_metadata$Iron_ID <- toupper(iron_metadata$Iron_ID)

all_present <- all(Sample_data_IDs$Sample_ID %in% iron_metadata$Iron_ID)

if (all_present) {
  message("All sample IDs are present in metadata.")
} else {
  message("Some sample IDs are missing from metadata.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(Sample_data_IDs$Sample_ID, iron_metadata$Iron_ID)
  print(missing_ids)
}

# missing_ids <- data.frame(missing_ids)
# 
# ##Export missing IDs to csv
# write.csv(missing_ids, "20250529_COMPASS_Iron_missing_IDs.csv")

##Remove cm units from Depth
Sample_data_IDs$Depth_cm <- gsub("cm", "", Sample_data_IDs$Depth)

##Select only necessary columns
all_data_selected <- Sample_data_IDs %>%
  dplyr::select(-c("RUN.", "Month", "Y.N", "Depth"))

all_data_selected$Depth_cm <- as.integer(all_data_selected$Depth_cm)

##Select only necessary columns
iron_metadata_selected <- iron_metadata %>%
  dplyr::select(Iron_ID, Year, Month, Day, Time..24hr., Time.Zone_EDT.EST,  Field.Notes) 

#merge metadata with sample run data 
merged_data <- all_data_selected %>%
  left_join(iron_metadata_selected, by = c("Sample_ID" = "Iron_ID"))

```

## Visualize Data by Plot   
```{r Visualize Data, echo=FALSE, warning=FALSE}

#Plot samples to get a first look at concentrations (sanity check)
data_plotting <- merged_data

#Order by Zone from Upland to Surface Water
data_plotting$Zone = factor(data_plotting$Zone, levels = c("UP",  "SWAMP","TR", "WC", "SW"))
data_plotting <- data_plotting[order(data_plotting$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting$Site = factor(data_plotting$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting <- data_plotting[order(data_plotting$Site), ]

#group the data for plotting
data_plotting <- data_plotting %>%
  group_by(Site) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

#Plot data and change colors based on Zone:
viz_Iron_plot <- ggplot(data_plotting, aes(x = row_num, y = FE_3_mean, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
  facet_wrap(~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "SWAMP" = "darkgrey",
    "TR" = "#FFBC42",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  theme(axis.text.x = element_blank()) +
  labs(x = " ", y = "Iron II (uM)", title = "Samples: Iron") +
  scale_x_discrete(drop = TRUE)+
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.1))

viz_Iron_plot

```

##Export Combined Data
```{r Export Combined Data, include = FALSE}

#Prepare data to be exported - if there is anything else to add 
#Add any necessary identifiers to the samples  ### VERY IMPORTANT AND STANDARD FOR PROJECT ####
  #example read in sample IDs list and merge 
  #create required ID columns in R, etc. 
final_data <- merged_data %>% 
  mutate(
    Project = "COMPASS: Synoptic",   # new column with same value on every row
    Region = "CB", 
    Run_notes  = run_notes
  ) 

colnames(final_data)

final_data <- final_data %>%
  rename(
    Time = Time..24hr.,
    Time_Zone = Time.Zone_EDT.EST,
    Field_notes = Field.Notes
    # add more rename pairs as needed
  ) %>%
  dplyr::select(
    Project, Region, Site, Zone, Replicate, Depth_cm, 
         Sample_ID, Year, Month, Day, Time, Time_Zone,
         FE_3_mean, Analysis_rundate, Field_notes, Run_notes
        # list columns in the order you want them
  )

head(final_data)

#Write out final data frame 
write.csv(final_data, "COMPASS_Iron_202410_Plates_Combined_Data.csv")

# #Write out samples that still have high CVs
# write.csv(HighCV, "COMPASS_Iron_202410_Plates_Combined_HighCVs.csv")
# 
# #Write out ADL samples
# write.csv(ADL_not_fixed, "COMPASS_Iron_202410_Plates_Combined_ADL.csv")
# 
# #Write out BDL samples
# write.csv(BDL_not_fixed, "COMPASS_Iron_202410_Plates_Combined_BDL.csv")


```


##END