---
title: "FeCode_updated2025_20260226"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
library(purrr)
library(tidyverse)
library(here)
library(data.table)
library(matrixStats)
library(gridExtra)
library(grid)
library(plater)
library(raster)
library(knitr)  
library(readxl)

#take out of sci notation
options(scipen = 999)

```

# Things that need to be changed
```{r Information to be changed}

Date_Run = "20260226"
plates<- c("Plate1", "Plate2")
Std_plates<- c("STD")
Run_by = "Zoe Read"#Instrument user
Script_run_by ="Zoe Read"#Code user
Project = "COMPASS_Synoptic_CB"
Experiment="2025"
Run_notes=""#any notes from run

#Stds that should be excluded
stds_to_remove<-c("")
```
## File paths 
```{r file paths, echo=FALSE}
#read in all the Tidy data files
folder_path_tidydata <- paste0("Tidy Data")
file_paths_STDs <- list.files(path = folder_path_tidydata, full.names = TRUE,pattern =paste0(Date_Run,".+.STD"))
head(file_paths_STDs)
# file_paths_tidydata <- list.files(path = folder_path_tidydata, full.names = TRUE,pattern =paste0(Date_Run,".+.Plate"))

file_paths_tidydata <- list.files(path = folder_path_tidydata, full.names = TRUE,pattern ="Plate")

head(file_paths_tidydata)

#file path name for processed data
summarizeddata_path=paste0("QAQC'd Data/",Project,"_FE_",Date_Run,"_Summary_Data.csv")
fulldata_path=paste0("QAQC'd Data/",Project,"_FE_",Date_Run,"_Data.csv")
shortdata_path=paste0("QAQC'd Data/",Project,"_FE_",Date_Run,"_short_Data.csv")
out_of_Range=paste0("QAQC'd Data/",Project,"_FE_",Date_Run,"_out_of_range_Data.csv")
high_CV=paste0("QAQC'd Data/",Project,"_FE_",Date_Run,"_high_CV_Data.csv")
#QAQC log path
log_path="Fe_STD_QAQC.csv"

```

```{r Set Up Code - constants and QAQC cutoffs, include=FALSE}

#Flag cutoffs
  r2_cutoff = 0.985            #this is the level below which we want to rerun or consider a curve 
  cv_flag_stds = 10           #this is the maximum cv allowed for standards
  p_value_chkstds = 0.05      #the p-value for the t-test between check  and top std must be greater than this
  cv_flag_sample = 10         #this is the maximum cv allowed for samples
  dups_perc_diff = 15.5       #this is the maximum percent difference allowed between duplicates
  high_recovery_cutoff = 120  #this is the maximum percent recovery allowed in spiked samples
  low_recovery_cutoff = 80    #this is the minimum percent recovery allowed in spiked samples

# Chk Standard concentrations - Update if running different standard curve:
# standard units are in uM


#Spike concentration calc
  #spike for these samples was 55 uL of the 60uM standard
  Con1 <- 1000000               #conversion factor value for spike volumes (uL -> L)
  spk_std <- 60                # uM Fe2- standard used
  spkvol <- 55                  # uL volume of spike added
  spk_Conc <- spk_std*(spkvol/Con1)   # this would be in umoles of FE in the spk
  sample_vol <- 160             # the sample volume without the spike is 160 uL

#Top standard Concentration- Update if running different standard curve: 
Top_STD_Fe2_low = 60
Top_STD_Fe3_low = 60

Top_STD_Fe2_high = 500
Top_STD_Fe3_high = 100
Top_STD_Fe2_Abs2_high=100

```

```{r read in Std curves,echo=FALSE,include=FALSE}

#read in the csv files
stds<-read_plates(
  files = "Tidy Data/20260226_COMPASS_Synoptic_CB_Fe_STD.csv",#list of all file paths
  plate_names = Std_plates ,  #list of plate names          
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
) 

stds <- stds %>%
  rename(Dilution = Dilutions)

head(stds)
  
```

```{r read in Data,echo=FALSE,include=FALSE}

#read in the csv files
dat<-read_plates(
  files = file_paths_tidydata,#list of all file paths
  plate_names = plates ,  #list of plate names          
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
) 

dat <- dat %>%
  rename(Dilution = Dilutions)

str(dat)
head(dat)
dat$QAQC_flag<-""
  
```

```{r Remove high cv stds,echo=FALSE,fig.keep='none',include=FALSE}
stds_all<- stds[!is.na(stds$IDs),]
stds_all_fixed<-stds_all

#Calculate cv for STDs
stds1_all <- stds_all %>%
  group_by(IDs) %>%
  summarise(Abs1_mean = mean(na.omit(Abs1)), Abs1_sd = sd(na.omit(Abs1)), Abs1_cv = cv(na.omit(Abs1)),
              Abs2_mean = mean(na.omit(Abs2)), Abs2_sd = sd(na.omit(Abs2)), Abs2_cv = cv(na.omit(Abs2)),
              Dilution = first(Dilution),
              Conc_uM = first(Conc_uM),
              Plate = first(Plate))
#flag high cv stds
stds1_all$Abs1_cv_flag <- ifelse(stds1_all$Abs1_cv > cv_flag_stds, 'High CV', 'Within range')
stds1_all$Abs2_cv_flag <- ifelse(stds1_all$Abs2_cv > cv_flag_stds, 'High CV', 'Within range')
#make dataframes for the fixed data
stds1_all_fixed<-stds1_all
stds_Abs1_HCV_fixed<-data.frame(matrix(ncol = 12, nrow =0 ))
colnames(stds_Abs1_HCV_fixed)<-colnames(stds1_all)
  
#filter High CV Samples for Abs1
stds_Abs1_all_HCV <- stds1_all %>%  
    filter(str_detect(Abs1_cv_flag, "High CV"))

#this if loop will keep the code from breaking if there are no high cv points
if(!nrow(stds_Abs1_all_HCV)==0){  
    
    #make a loop to filter each high cv sample
    for(x in 1:nrow(stds_Abs1_all_HCV)){
      sing_ID<- stds_all_fixed %>% subset(IDs == stds_Abs1_all_HCV$IDs[x])
      #make a dataframe for cvs
      cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
      
      #make a loop to find the cv for each deleted point
      for(i in 1:nrow(sing_ID)){
        sing_ID1<-sing_ID[-i,] 
        cv_trial[i]<-cv(sing_ID1$Abs1)
      }
      
      # delete the point that gives the lowest cv
      sing_ID[which.min(cv_trial),c(3)]=NA
      stds_Abs1_HCV_fixed<-rbind(stds_Abs1_HCV_fixed,sing_ID)
  }
    #recalculate the average and cv  
    stds1_Abs1_all_fixed <- stds_Abs1_HCV_fixed %>%
      group_by(IDs) %>%
      summarise(Abs1_mean = mean(na.omit(Abs1)), Abs1_sd = sd(na.omit(Abs1)), Abs1_cv = cv(na.omit(Abs1)),
                Abs2_mean = mean(na.omit(Abs2)), Abs2_sd = sd(na.omit(Abs2)), Abs2_cv = cv(na.omit(Abs2)),
                Dilution = first(Dilution),
                Conc_uM = first(Conc_uM),
                Plate = first(Plate))
    #Flag high cvs
    stds1_Abs1_all_fixed$Abs1_cv_flag <- ifelse(stds1_Abs1_all_fixed$Abs1_cv > cv_flag_stds, 'High CV', 'Within range')
    stds1_Abs1_all_fixed$Abs2_cv_flag <- ifelse(stds1_Abs1_all_fixed$Abs2_cv > cv_flag_stds, 'High CV', 'Within range')
    
    #bind fixed dataframes back together with non high cv points
    #averaged stds
    stds1_all_fixed <- subset(stds1_all_fixed, (!IDs %in% stds1_Abs1_all_fixed$IDs))
    stds1_all_fixed <- rbind(stds1_Abs1_all_fixed, stds1_all_fixed) 
    #all points  
    stds_all_fixed <- subset(stds_all_fixed, (!IDs %in% stds_Abs1_HCV_fixed$IDs))
    stds_all_fixed <- rbind(stds_Abs1_HCV_fixed, stds_all_fixed) 
  }


#make dataframe for fixed data
stds_Abs2_HCV_fixed<-data.frame(matrix(ncol = 12, nrow =0 ))
colnames(stds_Abs2_HCV_fixed)<-colnames(stds1_all)

#filter High CV Samples
stds_Abs2_all_HCV <- stds1_all_fixed %>%  
    filter(str_detect(Abs2_cv_flag, "High CV"))
#this if loop will keep the code from breaking if there are no high cv points
if(!nrow(stds_Abs2_all_HCV)==0){  
    
    #make a loop to filter each high cv sample
    for(x in 1:nrow(stds_Abs2_all_HCV)){
      sing_ID<- stds_all_fixed %>% subset(IDs == stds_Abs2_all_HCV$IDs[x])
      #make a dataframe for cvs
      cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
      
      #make a loop to find the cv for each deleted point
      for(i in 1:nrow(sing_ID)){
        sing_ID1<-sing_ID[-i,] 
        cv_trial[i]<-cv(sing_ID1$Abs2)
      }
      
      # delete the point that gives the lowest cv
      sing_ID[which.min(cv_trial),c(4)]=NA
      stds_Abs2_HCV_fixed<-rbind(stds_Abs2_HCV_fixed,sing_ID)
    }
    #recalculate the average and cv  
    stds1_Abs2_all_fixed <- stds_Abs2_HCV_fixed %>%
      group_by(IDs) %>%
      summarise(Abs1_mean = mean(na.omit(Abs1)), Abs1_sd = sd(na.omit(Abs1)), Abs1_cv = cv(na.omit(Abs1)),
                Abs2_mean = mean(na.omit(Abs2)), Abs2_sd = sd(na.omit(Abs2)), Abs2_cv = cv(na.omit(Abs2)),
                Dilution = first(Dilution),
                Conc_uM = first(Conc_uM),
                Plate = first(Plate))
    #Flag high cvs
    stds1_Abs2_all_fixed$Abs1_cv_flag <- ifelse(stds1_Abs2_all_fixed$Abs1_cv > cv_flag_stds, 'High CV', 'Within range')
    stds1_Abs2_all_fixed$Abs2_cv_flag <- ifelse(stds1_Abs2_all_fixed$Abs2_cv > cv_flag_stds, 'High CV', 'Within range')
    #bind fixed dataframes back together with non high cv points
    #averaged stds
    stds1_all_fixed <- subset(stds1_all_fixed, (!IDs %in% stds1_Abs2_all_fixed$IDs ))
    stds1_all_fixed <- rbind(stds1_Abs2_all_fixed, stds1_all_fixed) 
    #all points  
    stds_all_fixed <- subset(stds_all_fixed, (!IDs %in% stds_Abs2_HCV_fixed$IDs ))
    stds_all_fixed <- rbind(stds_Abs2_HCV_fixed, stds_all_fixed) 
  }

# filter High CV Stds
stds_HCV <- stds1_all_fixed %>%  
subset(Abs1_cv_flag == "High CV"|Abs2_cv_flag == "High CV")
```


```{r subset Stds,include=FALSE}
stds_all_fixed<-filter(stds_all_fixed, !IDs %in% stds_to_remove)
#subset by FE_2
FE_2_stds_Abs1 <- stds_all_fixed %>%  
  filter(str_detect(stds_all_fixed$IDs, "Fe_2")) %>% na.omit(stds_all_fixed$Abs1) 
FE_2_stds_Abs2 <- stds_all_fixed %>%  
  filter(str_detect(stds_all_fixed$IDs, "Fe_2")) %>% filter(!(Conc_uM > Top_STD_Fe2_Abs2_high))%>% na.omit(stds_all_fixed$Abs2)
#subset by FE_3
FE_3_stds_Abs1 <- stds_all_fixed %>%  
  filter(str_detect(stds_all_fixed$IDs, "Fe_3"))%>% na.omit(stds_all_fixed$Abs1)
FE_3_stds_Abs2 <- stds_all_fixed %>%  
  filter(str_detect(stds_all_fixed$IDs, "Fe_3")) %>% filter(!(Conc_uM > Top_STD_Fe3_high)) %>% na.omit(stds_all_fixed$Abs2)

#subset Fe3 STD with in their linear range
FE_3_stds_high <- FE_3_stds_Abs2

#Subset the low standard curve from all 
FE_2_stds_low <- FE_2_stds_Abs1[!(FE_2_stds_Abs1$Conc_uM > Top_STD_Fe2_low),]
FE_3_stds_low <- FE_3_stds_Abs2[!(FE_3_stds_Abs2$Conc_uM > Top_STD_Fe3_low),]

#subset Fe2 abs2 to 100um 
FE_2_stds_high <-FE_2_stds_Abs1
FE_2_stds_diff<-FE_2_stds_Abs2

```


```{r calculate slope and intercept, include=FALSE}
###Fe(2)
#Low curve for Fe(II)
FE_2_low_lm <- lm(FE_2_stds_low$Abs1 ~ FE_2_stds_low$Conc_uM)
summary(FE_2_low_lm)
cf <- coef(FE_2_low_lm)

#create data frame with 1 rows and 0 columns
Slopes1 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes1$Date <- Date_Run
Slopes1$Project <- Project
Slopes1$Curve <- "FE_2_low"
Slopes1$R2 <- summary(FE_2_low_lm)$adj.r.squared
Slopes1$Slope <- cf[2]
Slopes1$Intercept <- cf[1]
Slopes1$Top_STD <- Top_STD_Fe2_low


#High curve for Fe2
 FE_2_high_lm <- lm(FE_2_stds_Abs1$Abs1 ~ FE_2_stds_Abs1$Conc_uM)
 summary(FE_2_high_lm)
 cf <- coef(FE_2_high_lm)
 
#create data frame with 1 rows and 0 columns
Slopes2 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes2$Date <- Date_Run
Slopes2$Project <- Project
Slopes2$Curve <- "FE_2_high"
Slopes2$R2 <- summary(FE_2_high_lm)$adj.r.squared
Slopes2$Slope <- cf[2]
Slopes2$Intercept <- cf[1]
Slopes2$Top_STD <- Top_STD_Fe2_high

 
#curve for Fe(II) Abs2
FE_2_stds_Abs2_lm <- lm(FE_2_stds_Abs2$Abs2 ~ FE_2_stds_Abs2$Conc_uM)
summary(FE_2_stds_Abs2_lm)
cf <- coef(FE_2_stds_Abs2_lm)

#create data frame with 1 rows and 0 columns
Slopes5 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes5$Date <- Date_Run
Slopes5$Project <- Project
Slopes5$Curve <- "FE_2_abs2"
Slopes5$R2 <- summary(FE_2_stds_Abs2_lm)$adj.r.squared
Slopes5$Slope <- cf[2]
Slopes5$Intercept <- cf[1]
Slopes5$Top_STD <- Top_STD_Fe2_Abs2_high


#curve for Fe(II) Abs2-Abs1
FE_2_stds_diff$FE_2_Diff<-FE_2_stds_diff$Abs2-FE_2_stds_diff$Abs1

FE_2_diff_lm <- lm(FE_2_stds_diff$FE_2_Diff ~ FE_2_stds_diff$Conc_uM)
summary(FE_2_diff_lm)
cf <- coef(FE_2_diff_lm)

#create data frame with 1 rows and 0 columns
Slopes7 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes7$Date <- Date_Run
Slopes7$Project <- Project
Slopes7$Curve <- "FE_2_diff"
Slopes7$R2 <- summary(FE_2_diff_lm)$adj.r.squared
Slopes7$Slope <- cf[2]
Slopes7$Intercept <- cf[1]
Slopes7$Top_STD <- Top_STD_Fe2_Abs2_high


###Fe(3)
#Low curve for Fe(III)
FE_3_low_lm <- lm(FE_3_stds_low$Abs2 ~ FE_3_stds_low$Conc_uM)
summary(FE_3_low_lm)
cf <- coef(FE_3_low_lm)

#create data frame with 1 rows and 0 columns
Slopes3 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes3$Date <- Date_Run
Slopes3$Project <- Project
Slopes3$Curve <- "FE_Tot_low"
Slopes3$R2 <- summary(FE_3_low_lm)$adj.r.squared
Slopes3$Slope <- cf[2]
Slopes3$Intercept <- cf[1]
Slopes3$Top_STD <- Top_STD_Fe3_low


#High curve for Fe(III) 
FE_3_high_lm <- lm(FE_3_stds_high$Abs2 ~ FE_3_stds_high$Conc_uM)
summary(FE_3_high_lm)
cf <- coef(FE_3_high_lm)
 
 #create data frame with 1 rows and 0 columns
Slopes4 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes4$Date <- Date_Run
Slopes4$Project <- Project
Slopes4$Curve <- "FE_Tot_high"
Slopes4$R2 <- summary(FE_3_high_lm)$adj.r.squared
Slopes4$Slope <- cf[2]
Slopes4$Intercept <- cf[1]
Slopes4$Top_STD <- Top_STD_Fe3_high


#curve for Fe(III) Abs1
FE_3_abs1_lm <- lm(FE_3_stds_Abs1$Abs1 ~ FE_3_stds_Abs1$Conc_uM )
summary(FE_3_abs1_lm)
cf <- coef(FE_3_abs1_lm)

#create data frame with 1 rows and 0 columns
Slopes6 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes6$Date <- Date_Run
Slopes6$Project <- Project
Slopes6$Curve <- "FE_3_abs1"
Slopes6$R2 <- summary(FE_3_abs1_lm)$adj.r.squared
Slopes6$Slope <- cf[2]
Slopes6$Intercept <- cf[1]
Slopes6$Top_STD <- Top_STD_Fe2_high


#This slope should be small e-6

Slopes<- rbind(Slopes1, Slopes2, Slopes3, Slopes4, Slopes5, Slopes6, Slopes7)

#Slopes6$Slope should be close to 0, e-5
####
```
# Plot standards 
```{r plot all the std curves, echo=FALSE, message=FALSE,fig.keep='last'}
#Plot each of these and calculate the slope, intercept, and R2
#Low curve for Fe(II)
FE_2_low <- ggplot(FE_2_stds_low, aes(Conc_uM, Abs1)) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("Low Fe(II) Absorbance 1 Curve")
FE_2_low

#High curve for Fe2
 FE_2_high <- ggplot(FE_2_stds_Abs1, aes(Conc_uM, Abs1)) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("High Fe(II) Absorbance 1 Curve")
 FE_2_high
 
 #curve for Fe(II) Abs2
FE_2_abs2 <- ggplot(FE_2_stds_Abs2, aes(Conc_uM, Abs2)) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(II) Absorbance 2 Curve")
FE_2_abs2

#curve for Fe(II) Abs2-Abs1
FE_2 <- ggplot(FE_2_stds_diff, aes(Conc_uM,FE_2_Diff)) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(II) Absorbance Difference Curve")
FE_2

#Low curve for Fe(III)
FE_3_low <- ggplot(FE_3_stds_low, aes(Conc_uM, Abs2)) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("Low Fe(III) Absorbance 2 Curve")
FE_3_low

#High curve for Fe(III)
FE_3_high <- ggplot(FE_3_stds_high, aes(Conc_uM, Abs2)) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("High Fe(III) Absorbance 2 Curve")
FE_3_high

#curve for Fe(III) Abs1
FE_3_abs1 <- ggplot(FE_3_stds_Abs1, aes(Conc_uM,Abs1 )) + geom_point() + facet_wrap(~Plate)+ geom_smooth(method = "lm", se = FALSE)+ggtitle("Fe(III) Absorbance 1 Curve")
FE_3_abs1

knitr::kable(Slopes,caption = "Slopes")
ggarrange(FE_2_low,FE_2_high,FE_2_abs2,FE_2,FE_3_low,FE_3_high,FE_3_abs1)
```

# Checking STD Data against QAQC file

```{r read in QAQC file, echo=FALSE}


#read in datafile with all the slopes
qlogO <- read.csv(log_path)
qlogO <- qlogO[,c(1:7)]
colnames(qlogO) <- c("Date", "Project", "Curve", "R2", "Slope", "Intercept", "Top_STD")

#add data to file
qlog <- rbind(Slopes, qlogO)
qlog <- qlog %>% distinct(.keep_all = TRUE)
```

### R2
```{r check R2-1, echo=FALSE,fig.width=3, fig.height=2,warning=FALSE, fig.show='hold',results='hide'}
#plot R2 for all curves
Rsq1_graph<-function(curve){
  qlog1<-qlog %>% filter(Curve == curve)
Rsq1 <- ggplot(data=qlog1, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", linewidth=2)+
  geom_point() + 
  theme_classic() + ylim(0.96, 1.01) + 
  theme(legend.position="none")+ 
  ggtitle(paste0(curve," R2"))+
  guides(x = guide_axis(angle = 70))+
  geom_point(data = qlog %>% filter(Date %in% Slopes$Date & R2 %in% Slopes$R2 & Curve %in% curve),color = "orange")
Rsq1
}
#apply function for each curve
lapply(Slopes[-c(6),]$Curve,Rsq1_graph)
```


```{r check R2-2, echo=FALSE}
#Is the intercept for each plate within 2 standard deviation of the log file? 
for(i in 1:nrow(Slopes)){
  if (Slopes$R2[i] < (r2_cutoff)) {
    ifelse(Slopes$Curve[i]=="FE_3_abs1",print(paste0(Slopes$Curve[i],": R2 will always be low")), print(paste0(Slopes$Curve[i],": Std Curve r2 is below cutoff! - REASSESS")))
  } else {
   print(paste0(Slopes$Curve[i],": Std Curve r2 GOOD"))
  }
  if(!Slopes$Curve[i]=="FE_3_abs1"){
 #writes a flag for Curves that have a low R2   
    dat$QAQC_flag<-if(Slopes$R2[i] <= r2_cutoff){
      ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; ",Slopes$Curve[i] ," Std curve r2 low"),paste0(Slopes$Curve[i]," Std curve r2 low"))                   
    }else{
      ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
    }
  }
}


#Rerun if outside of red lines
#FE_3_abs1 R2 will never be above cut off, This is fine
```
### Slopes
```{r check slopes-1, echo=FALSE,fig.width=3, fig.height=2,warning=FALSE, fig.show='hold', results='hide'}
##plot the slopes to make sure there are no crazy outliers 
#plot slopes for all curves
slope_graph<-function(curve){
  qlog1<-qlog %>% filter(Curve == curve)
slope_Fe <- ggplot(data=qlog1, aes(x=Date, y=Slope)) +
  geom_hline(yintercept= (mean(qlog1$Slope)+ (2*sd(qlog1$Slope))), linetype="dashed", color = "red", linewidth=2)+
  geom_hline(yintercept= (mean(qlog1$Slope)- (2*sd(qlog1$Slope))), linetype="dashed", color = "red", linewidth=2)+
  geom_point() + 
  theme_classic() +  
  theme(legend.position="none") + 
  ggtitle( paste0(curve," Slopes"))+
  guides(x = guide_axis(angle = 70))+
  geom_point(data = qlog %>% filter(Date %in% Slopes$Date & Slope %in% Slopes$Slope & Curve %in% curve),color = "orange")
slope_Fe
}
#apply function for each curve
lapply(Slopes$Curve,slope_graph)
```


```{r check slopes-2, echo=FALSE}
#Is the slope for each plate within 2 standard deviation of the log file? 
for(i in 1:nrow(Slopes)){
  qlog1<-qlog %>% filter(Curve == Slopes$Curve[i])
  if (Slopes$Slope[i] > (mean(qlog1$Slope)+ (2*sd(qlog1$Slope)))| Slopes$Slope[i] < (mean(qlog1$Slope)- (2*sd(qlog1$Slope)))) {
    print(paste0(Slopes$Curve[i],": Std curve slope is 2 sd different from previous slopes! 
      - REASSESS"))
  } else {
    print(paste0(Slopes$Curve[i],":Std curve slope is with 2 sd of previous slopes"))
  }}
#Fe3_abs1$Slope should be close to 0, e-5
#Rerun if outside of red lines
```

### Intercepts
```{r check intercepts-1, echo=FALSE,fig.width=3, fig.height=2,warning=FALSE, fig.show='hold', results='hide'}
##plot the intercepts to make sure there are no crazy outliers 
# Plot intercept for all curves
intercept_graph<-function(curve){
  qlog1<-qlog %>% filter(Curve == curve)
intercept_Fe <- ggplot(data=qlog1, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog1$Intercept)+ (2*sd(qlog1$Intercept))), linetype="dashed", color = "red", linewidth=2)+
  geom_hline(yintercept= (mean(qlog1$Intercept)- (2*sd(qlog1$Intercept))), linetype="dashed", color = "red", linewidth=2)+
  geom_point() + 
  theme_classic() +  
  theme(legend.position="none") + 
  ggtitle( paste0(curve," Intercepts"))+
  guides(x = guide_axis(angle = 70))+
  geom_point(data = qlog %>% filter(Date %in% Slopes$Date & Intercept %in% Slopes$Intercept & Curve %in% curve),color = "orange")

intercept_Fe
}
#apply function for each curve
lapply(Slopes$Curve,intercept_graph)
```


```{r check intercepts-2, echo=FALSE}
#Is the intercept for each plate within 2 standard deviation of the log file? 
for(i in 1:nrow(Slopes)){
  qlog1<-qlog %>% filter(Curve == Slopes$Curve[i])
  if (Slopes$Intercept[i] > (mean(qlog1$Intercept)+ (2*sd(qlog1$Intercept)))| Slopes$Intercept[i] < (mean(qlog1$Intercept)- (2*sd(qlog1$Intercept)))) {
    print(paste0(Slopes$Curve[i],":Std curve intercept is 2 sd different from previous intercepts! 
      - REASSESS"))
  } else {
    print(paste0(Slopes$Curve[i],":Std curve intercept is with 2 sd of previous intercepts"))
  }}

#Rerun if outside of red lines
```


# Calculate Reduction Efficiency 
```{r calculate reduction Efficiency, echo=FALSE}
#pull out Fe_2 and Fe_3 standards at 60uM to check reduction efficiency

#subset by FE_2 and FE_Total
FE_2_60uM <- stds %>%  
  filter(str_detect(stds$IDs, "60 uM Fe_2 A"))  

#subset by FE_2 and FE_Total
FE_3_60uM <- stds %>%  
  filter(str_detect(IDs, "60 uM Fe_3 A"))  

#create data frame with 0 rows and 3 columns
Fe_Red_eff <- data.frame(matrix(ncol = 0, nrow = 3))
Fe_Red_eff$ID <- FE_2_60uM$Wells
Fe_Red_eff$FE2 <- (FE_2_60uM$Abs1)
Fe_Red_eff$FE3 <- FE_3_60uM$Abs2
Fe_Red_eff$Eff <- ((Fe_Red_eff$FE3)/Fe_Red_eff$FE2)*100
Fe_Red_eff$Eff1 <- ((Fe_Red_eff$FE3/0.8)/Fe_Red_eff$FE2)*100 #this one accounts for dilution of the reagents 

Fe_Red_eff$Eff_flag <-  ifelse(Fe_Red_eff$Eff1 > low_recovery_cutoff & Fe_Red_eff$Eff1 < high_recovery_cutoff, 'OK', 'NO, rerun')  
knitr::kable(Fe_Red_eff, caption = "Reduction Efficiency")

#plot recoveries as a bar graph to easily check 
redeffbar <- ggplot(data = Fe_Red_eff, aes(x = ID, y = Eff1,fill=Eff_flag)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK"="darkgreen","NO, rerun"="darkred")) + 
        theme_classic() + labs(x= "Sample ID", y="Reduction Efficiency (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=low_recovery_cutoff, linetype="dashed", color = "black", linewidth=1) + 
        geom_hline(yintercept=high_recovery_cutoff, linetype="dashed", color = "black", linewidth=1)
redeffbar

BadEff <- subset(Fe_Red_eff, Eff_flag == "NO, rerun")

#write out a flag to the sample dataframe if the efficiency have recovery out of range 

dat$QAQC_flag<-if(nrow(BadEff) > 1){
  ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Reduction Efficiency out of range"),"Reduction Efficiency out of range")                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
}

Eff_percent <- (sum(Fe_Red_eff$Eff_flag == "OK")/nrow(Fe_Red_eff))*100
#report out if flags indicate need for rerun
ifelse(Eff_percent >= 60  , ">60% of Reduction Efficiencies are within range",
       "<60% of Reduction Efficiencies are out of range - REASSESS")


```

## Method Minimum Detection Limit (MDL) Calculation
I use the lowest standard to calculate MDL
How to calculate MDL https://www.epa.gov/sites/default/files/2016-12/documents/mdl-procedure_rev2_12-13-2016.pdf
Look at Table 1 on page 5
The Student's t-value used to calculate the method detection limit (MDL) is the one appropriate for a 99% confidence level and a standard deviation estimate with n-1 degrees of freedom
```{r MDL, echo=FALSE}
#make a dataframe of just 0.6 uM Fe2 samples from standard dataframe
MDL_dat <- subset(FE_2_stds_Abs1, Conc_uM == 0.6)

#calculate the standard deviation of abs 1
stddev <- sd(MDL_dat$Abs1)

#calculate n (number of 0.6 uM samples)
n <- nrow(MDL_dat)

#Identify the proper Student's T for n-1 from the table link above
StudentsT <-qt(.99,(n-1))

#Multiply Standard deviation by Student's T
MDL <- stddev*StudentsT
MDL
```


# Check standards QAQC

```{r Assess check stds, echo=FALSE}
#filter std curves
Fe2_low <- subset(Slopes, Curve == "FE_2_low")
Fe2_high <- subset(Slopes, Curve == "FE_2_high")
FeTot_low <- subset(Slopes, Curve == "FE_Tot_low")
FeTot_high <- subset(Slopes, Curve == "FE_Tot_high")
Fe2_abs2 <- subset(Slopes, Curve == "FE_2_abs2")
Fe3_abs1 <- subset(Slopes, Curve == "FE_3_abs1")
Fe2_diff <- subset(Slopes, Curve == "FE_2_diff")

#creat data frame for p values
QAstdchk<-data.frame(matrix(nrow=0,ncol=3))
        colnames(QAstdchk) <- c("Plate","IDs","p_value")

#Make sure check standards are not different from standard concentration
#subset Check Standards from the rest of the dataset
Fe2_std_Chk <- dat %>%  filter(str_detect(IDs, "Fe_2"))
Fe3_std_Chk <- dat %>%  filter(str_detect(IDs, "Fe_3"))

#Calculate Check standard Concentration
Fe2_std_Chk$Conc_uM <- (Fe2_std_Chk$Abs1-Fe2_low$Intercept)/Fe2_low$Slope
Fe3_std_Chk$Conc_uM <-(Fe3_std_Chk$Abs2-FeTot_low$Intercept)/FeTot_low$Slope


#Are the check standards significantly different from the standards?
#subset datasets for comparison
Fe2_std0.6 <- subset(FE_2_stds_low, Conc_uM == 0.6)
Fe2_std9 <- subset(FE_2_stds_low, Conc_uM == 9.0)
Fe2_std60 <- subset(FE_2_stds_low, Conc_uM == 60.0)

Fe3_std0.6 <- subset(FE_3_stds_low, Conc_uM == 0.6)
Fe3_std9 <- subset(FE_3_stds_low, Conc_uM == 9.0)
Fe3_std60 <- subset(FE_3_stds_low, Conc_uM == 60.0)

#this loop compares the check stds for each plate to the std plate
for(i in 1:length(plates)){
#filter each plate  
  Fe2_std_Chk1<-filter(Fe2_std_Chk, Plate %in% plates[i])
  Fe3_std_Chk1<-filter(Fe3_std_Chk, Plate %in% plates[i])
print(plates[i])
#filter out the individual check stds
Fe2_Chkstd0.6 <- subset(Fe2_std_Chk1, IDs %like% "0.6")
Fe2_Chkstd9 <- subset(Fe2_std_Chk1, IDs %like% "9")
Fe2_Chkstd60 <- subset(Fe2_std_Chk1, IDs %like% "60")
Fe3_Chkstd0.6 <- subset(Fe3_std_Chk1, IDs %like% "0.6")
Fe3_Chkstd9 <- subset(Fe3_std_Chk1, IDs %like% "9")
Fe3_Chkstd60 <- subset(Fe3_std_Chk1, IDs %like% "60")

#this if else statement keeps the code from breaking if there are not enough values for the stds or chkstds
if(!nrow(Fe2_std0.6) == 0 & !nrow(Fe2_Chkstd0.6) == 0){
Fe2.t.test.std0.6 <- t.test(Fe2_std0.6$Abs1,Fe2_Chkstd0.6$Abs1,var.equal = T)
Fe2.t.test.std0.6

if(Fe2.t.test.std0.6$p.value > p_value_chkstds){
print("Fe2 Chk Std 0.6 GOOD")
} else {
print("Fe2 Chk Std 0.6 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
  dat$QAQC_flag<-if(Fe2.t.test.std0.6$p.value < p_value_chkstds){
    ifelse(dat$Plate %in% plates[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Fe2 Chk Std 0.6 out of range"),"Fe2 Chk Std 0.6 out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
  } 
#Add p-values to dataframes
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 0.6 Fe2",p_value=Fe2.t.test.std0.6$p.value)) 
}else{
  #if std didn't have enough points for t-test
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 0.6 Fe2",p_value=NA)) 
  print("Std excluded")
}

#this if else statement keeps the code from breaking if there are not enough values for the stds or chkstds
if(!nrow(Fe2_std9) == 0 & !nrow(Fe2_Chkstd9) == 0){
Fe2.t.test.std9 <- t.test(Fe2_std9$Abs1,Fe2_Chkstd9$Abs1,var.equal = T)
Fe2.t.test.std9

if(Fe2.t.test.std9$p.value > p_value_chkstds){
print("Fe2 Chk Std 9 GOOD")
} else {
print("Fe2 Chk Std 9 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
  dat$QAQC_flag<-if(Fe2.t.test.std9$p.value < p_value_chkstds){
    ifelse(dat$Plate %in% plates[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Fe2 Chk Std 9 out of range"),"Fe2 Chk Std 9 out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
  } 
#Add p-values to dataframes
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 9 Fe2",p_value=Fe2.t.test.std9$p.value)) 
}else{
  #if std didn't have enough points for t-test
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 9 Fe2",p_value=NA)) 
  print("Std excluded")
}

#this if else statement keeps the code from breaking if there are not enough values for the stds or chkstds
if(!nrow(Fe2_std60) == 0 & !nrow(Fe2_Chkstd60) == 0){
Fe2.t.test.std60 <- t.test(Fe2_std60$Abs1,Fe2_Chkstd60$Abs1,var.equal = T)
Fe2.t.test.std60

if(Fe2.t.test.std60$p.value > p_value_chkstds){
print("Fe2 Chk Std 60 GOOD")
} else {
print("Fe2 Chk Std 60 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
  dat$QAQC_flag<-if(Fe2.t.test.std60$p.value < p_value_chkstds){
    ifelse(dat$Plate %in% plates[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Fe2 Chk Std 60 out of range"),"Fe2 Chk Std 60 out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
  } 
#Add p-values to dataframes
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 60 Fe2",p_value=Fe2.t.test.std60$p.value)) 
}else{
  #if std didn't have enough points for t-test
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 60 Fe2",p_value=NA)) 
  print("Std excluded")
}

#this if else statement keeps the code from breaking if there are not enough values for the stds or chkstds
if(!nrow(Fe3_std0.6) == 0 & !nrow(Fe3_Chkstd0.6) == 0){
Fe3.t.test.std0.6 <- t.test(Fe3_std0.6$Abs2,Fe3_Chkstd0.6$Abs2,var.equal = T)
Fe3.t.test.std0.6

if(Fe3.t.test.std0.6$p.value > p_value_chkstds){
print("Fe 3 Chk Std 0.6 Fe2 GOOD")
} else {
print("Fe 3 Chk Std 0.6 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
  dat$QAQC_flag<-if(Fe3.t.test.std0.6$p.value < p_value_chkstds){
    ifelse(dat$Plate %in% plates[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Fe3 Chk Std 0.6 out of range"),"Fe3 Chk Std 0.6 out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
  } 
#Add p-values to dataframes
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 0.6 Fe3",p_value=Fe3.t.test.std0.6$p.value)) 
}else{
  #if std didn't have enough points for t-test
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 0.6 Fe3",p_value=NA)) 
  print("Std excluded")
}

#this if else statement keeps the code from breaking if there are not enough values for the stds or chkstds
if(!nrow(Fe3_std9) == 0 & !nrow(Fe3_Chkstd9) == 0){
Fe3.t.test.std9 <- t.test(Fe3_std9$Abs2,Fe3_Chkstd9$Abs2,var.equal = T)
Fe3.t.test.std9

if(Fe3.t.test.std9$p.value > p_value_chkstds){
print("Fe 3 Chk Std 9 GOOD")
} else {
print("Fe 3 Chk Std 9 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
  dat$QAQC_flag<-if(Fe3.t.test.std9$p.value < p_value_chkstds){
    ifelse(dat$Plate %in% plates[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Fe3 Chk Std 9 out of range"),"Fe3 Chk Std 9 out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
  } 
#Add p-values to dataframes
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 9 Fe3",p_value=Fe3.t.test.std9$p.value)) 
}else{
  #if std didn't have enough points for t-test
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 9 Fe3",p_value=NA)) 
  print("Std excluded")
}

#this if else statement keeps the code from breaking if there are not enough values for the stds or chkstds
if(!nrow(Fe3_std60) == 0 & !nrow(Fe3_Chkstd60) == 0){
Fe3.t.test.std60 <- t.test(Fe3_std60$Abs2,Fe3_Chkstd60$Abs2,var.equal = T)
Fe3.t.test.std60

if(Fe3.t.test.std60$p.value > p_value_chkstds){
print("Fe 3 Chk Std 60 GOOD")
} else {
print("Fe 3 Chk Std 60 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
  dat$QAQC_flag<-if(Fe3.t.test.std60$p.value < p_value_chkstds){
    ifelse(dat$Plate %in% plates[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Fe3 Chk Std 60 out of range"),"Fe3 Chk Std 60 out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
  } 
#Add p-values to dataframes
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 60 Fe3",p_value=Fe3.t.test.std60$p.value)) 
}else{
  #if std didn't have enough points for t-test
  QAstdchk<-rbind(QAstdchk,list(Plate=plates[i],IDs="Std 60 Fe3",p_value=NA)) 
  print("Std excluded")
}
}

#flag p-values
QAstdchk$p_flag <-  ifelse(QAstdchk$p_value > p_value_chkstds, 'OK', 'Rerun')

#plot output as a bar graph to easily check
ChkStdbar <- ggplot(data = QAstdchk, aes(x = IDs, y = p_value, fill=p_flag))+
        facet_wrap(~Plate)+
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values = c("OK" = "darkgreen", "Rerun" = "darkred"))+
        theme_classic() + labs(x= "Sample ID", y="T-test p-value") + 
        theme(legend.position="none") +  geom_hline(yintercept=0.05, linetype="dashed", 
                color = "black", linewidth=1)+ 
        ggtitle("Check Standards")

ChkStdbar
```

#No Matrix Checks
We do not use matrix checks for the Fe samples because our matrix checks are not made with Trace metal grade reagents. If we want to do this in the future, we need to use trace metal grade sodium chloride and sodium bicarbonate for making the matrix checks


```{r flag out of range data, echo=FALSE,include=FALSE}

## remove any lines that say stds in IDs
dat1 <- dat %>%  filter(!str_detect(IDs, "Fe_2")) %>%
  filter(!str_detect(IDs, "Fe_3")) %>%
  filter(!str_detect(IDs, "HCl"))

#Now flag any Abs1 that are above the 500uM standard absorbance
FE_2_500uM<- FE_2_stds_high %>%  filter(Conc_uM == Top_STD_Fe2_high)
FE_2_0uM<-subset(na.omit(stds_all_fixed), Conc_uM == 0)

dat1$FE_2_Curve <- ifelse(dat1$Abs1 > mean(FE_2_500uM$Abs1), "High", "Within range")

#and flag any that are below the 0um std + MDL
dat1$FE_2_Curve <- ifelse(dat1$Abs1 <(mean(FE_2_0uM$Abs1)+ MDL), "Below MDL", dat1$FE_2_Curve)

#Now flag any Abs2 that are above the 100uM standard absorbance for Fe2 or Fe3
FE_2_100uM<- FE_2_stds_Abs2 %>%  filter(Conc_uM == Top_STD_Fe2_Abs2_high)
FE_3_100uM<- FE_3_stds_Abs2 %>%  filter(Conc_uM == Top_STD_Fe3_high)

dat1$FE_Tot_Curve <- ifelse(dat1$Abs2 > mean(FE_2_100uM$Abs2), "High", ifelse((dat1$Abs2-dat1$Abs1) > mean(FE_3_100uM$Abs2), "High", "Within range")) 

#and flag any that are below the 0um std + MDL; this only flags samples that have no Fe2 in them
dat1$FE_Tot_Curve <- ifelse(dat1$Abs2 < (mean(FE_2_0uM$Abs2)+ MDL), "Below MDL", dat1$FE_Tot_Curve)

```

# Subset Data and Calculate Concentrations 
```{r calculate concentrations, echo=FALSE,include=FALSE}
###Calculate concentrations of Fe(II)
dat1$FE_2_Conc <- ifelse(dat1$Abs1 > mean(Fe2_std60$Abs1), (dat1$Abs1-Fe2_high$Intercept)/Fe2_high$Slope, (dat1$Abs1-Fe2_low$Intercept)/Fe2_low$Slope)

##since the absorbance for Fe2 changes between Abs1 and Abs2 we need to calculate the expected Abs2 for Fe2
##Expected_diff=m*conc+b
#m=Fe2_diff$Slope #b=Fe2_diff$Intercept
# abs2=abs1+diff
dat1$CalcAbs2<-dat1$Abs1+(dat1$FE_2_Conc*Fe2_diff$Slope+Fe2_diff$Intercept)

#subtract the expected Abs2 from the actual Abs2, this should leave you with the absorbance caused by the Fe3 reducing to Fe2
dat1$Abs2_diff<-dat1$Abs2-dat1$CalcAbs2

#using the slope from FeTot slope calculate Fe3 
#the intercept is not needed as the Abs2_diff starts at 0 
#Fe3_abs1$Slope should be close to 0, I'm unsure if this is the correct place to put it  
dat1$FE_3_Conc <-ifelse(dat1$Abs2 > mean(Fe3_std60$Abs2), (dat1$Abs2_diff)/(FeTot_high$Slope-Fe3_abs1$Slope), (dat1$Abs2_diff)/(FeTot_low$Slope-Fe3_abs1$Slope))


#Account for the dilution factor
dat1$FE_2_Conc_DilCorr <- dat1$FE_2_Conc*dat1$Dilution
dat1$FE_3_Conc_DilCorr <- dat1$FE_3_Conc*dat1$Dilution

#Use ifelse to make any negative values equal to zero 
dat1$FE_2_Conc_Final <- ifelse(dat1$FE_2_Conc_DilCorr <0, 0, dat1$FE_2_Conc_DilCorr)

dat1$FE_3_Conc_Final <- ifelse(dat1$FE_3_Conc_DilCorr <0, 0, dat1$FE_3_Conc_DilCorr)

#Create a column for total iron
dat1$FE_tot_Conc_Final<- dat1$FE_2_Conc_Final+ dat1$FE_3_Conc_Final

head(dat1)

```
## Calculate Averages across wells, std. dev, and cv.  
```{r Calculate Averages across wells and sd, echo=FALSE,fig.width=8,fig.height=5}
##Need to make it so it groups by Plate and Dilution so that reruns of the same ID don't count towards the CV
dat1$ID_full <- paste(dat1$Plate, dat1$Dilution, dat1$IDs, sep = "_")

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat_aver <- dat1 %>%
  group_by(ID_full) %>%
  summarise(IDs=first(IDs),FE_2_mean = mean(na.omit(FE_2_Conc_Final)), FE_2_sd = sd(na.omit(FE_2_Conc_Final)), FE_2_cv = cv(na.omit(FE_2_Conc_Final)),
            FE_3_mean = mean(na.omit(FE_3_Conc_Final)), FE_3_sd = sd(na.omit(FE_3_Conc_Final)), FE_3_cv = cv(na.omit(FE_3_Conc_Final)),
            FE_tot_mean =(mean(FE_2_Conc_Final,na.rm=TRUE)+ mean(FE_3_Conc_Final,na.rm=TRUE)), FE_tot_sd = sqrt(sd(na.omit(FE_3_Conc_Final))^2+sd(na.omit(FE_2_Conc_Final))),
            FE_2_flag = ifelse(mean(na.omit(Abs1)) > mean(FE_2_500uM$Abs1), "High",ifelse(mean(na.omit(Abs1)) < (mean(FE_2_0uM$Abs1)+ MDL), "Below MDL" ,"Within range")),
            FE_3_flag = ifelse(mean(na.omit(Abs2)) > mean(FE_2_100uM$Abs2)|(mean(na.omit(Abs2))-mean(na.omit(Abs1))) > mean(FE_3_100uM$Abs2), "High", ifelse(mean(na.omit(Abs2_diff)) < MDL|mean(na.omit(Abs2)) < (mean(FE_2_0uM$Abs2)+ MDL), "Below MDL" ,"Within range")),
            Dilution = first(Dilution),
            Plate = first(Plate),
            QAQC_flag = first(QAQC_flag))

#flag high cv data
dat_aver$FE_2_cv_flag <- ifelse(dat_aver$FE_2_cv > cv_flag_sample, "High cv rerun", "ok")
dat_aver$FE_3_cv_flag <- ifelse(dat_aver$FE_3_cv > cv_flag_sample, "High cv rerun", "ok")

#Make a list of samples to rerun
HighCVSamples <- subset(dat_aver, FE_2_cv_flag == "High cv rerun"|FE_3_cv_flag == "High cv rerun")

Fe2 <- ggplot(dat_aver, aes(x=IDs, y=FE_2_mean, color=FE_2_cv_flag))+ 
  geom_point(size=4) +  theme_classic() + 
  labs(y="Fe(II) (uM)", x="Sample ID",title=paste(Project, Experiment," Iron(II) Data"))+
  scale_color_manual(values = c("High cv rerun" = "red", "ok" = "darkgreen")) +
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_2_mean-FE_2_sd,
                    ymax=FE_2_mean+FE_2_sd),width=0.3,position=position_dodge(.1))
Fe2

Fe3 <- ggplot(dat_aver, aes(x=IDs, y=FE_3_mean,color=FE_3_cv_flag))+ 
  geom_point(size=4) +  theme_classic() +
  labs(y="Fe(III) (uM)", x="Sample ID",title=paste(Project, Experiment," Iron(III) Data"))+
  scale_color_manual(values = c("High cv rerun" = "red", "ok" = "darkgreen")) +
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_3_mean-FE_3_sd,
                    ymax=FE_3_mean+FE_3_sd),width=0.3,position=position_dodge(.1))
Fe3
```

# Remove Bad Reps 
```{r, echo=FALSE,include=FALSE}
#auto remove bad reps
# filter High CV Samples
dat1_HCV <- subset(dat_aver, FE_2_cv_flag == "High cv rerun"|FE_3_cv_flag == "High cv rerun")
dat_aver1 <- dat_aver

#make dataframe for fixed data
dat1_fixed<-dat1
dat_HCV_fixed_FE2<-data.frame(matrix(ncol = 19, nrow =0 ))
colnames(dat_HCV_fixed_FE2)<-colnames(dat1)
dat_HCV_fixed_FE3<-data.frame(matrix(ncol = 19, nrow =0 ))
colnames(dat_HCV_fixed_FE3)<-colnames(dat1)

#make a loop to filter each high cv sample for Fe 2
dat1_HCV_1<-subset(dat1_HCV,  FE_2_cv_flag == "High cv rerun")
for(x in 1:nrow(dat1_HCV_1)){
  dat_HCV <- subset(dat1_fixed, ID_full %in% dat1_HCV_1$ID_full )
  sing_ID<- dat_HCV %>% subset(ID_full == dat1_HCV_1$ID_full[x])
  #make a dataframe for cvs
  cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
  
  #make a loop to find the cv for each deleted point
  if(nrow(sing_ID)>2){
  for(i in 1:nrow(sing_ID)){
    sing_ID1<-sing_ID[-i,]  
    cv_trial[i]<-cv(sing_ID1$FE_2_Conc_Final)
  }
  
  # delete the point that gives the lowest cv
  if(which.min(cv_trial)<cv(sing_ID$FE_2_Conc_Final)){sing_ID[which.min(cv_trial),c(3,16)]=NA}
  dat_HCV_fixed_FE2<-rbind(dat_HCV_fixed_FE2,sing_ID)
  }else{
    dat_HCV_fixed_FE2<-rbind(dat_HCV_fixed_FE2,sing_ID) 
  }
}

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1_HCV_fixed_FE2 <- dat_HCV_fixed_FE2 %>%
  group_by(ID_full) %>%
  summarise(IDs=first(IDs),FE_2_mean = mean(na.omit(FE_2_Conc_Final)), FE_2_sd = sd(na.omit(FE_2_Conc_Final)), FE_2_cv = cv(na.omit(FE_2_Conc_Final)),
            FE_3_mean = mean(na.omit(FE_3_Conc_Final)), FE_3_sd = sd(na.omit(FE_3_Conc_Final)), FE_3_cv = cv(na.omit(FE_3_Conc_Final)),
            FE_tot_mean =(mean(FE_2_Conc_Final,na.rm=TRUE)+ mean(FE_3_Conc_Final,na.rm=TRUE)), FE_tot_sd = sqrt(sd(na.omit(FE_3_Conc_Final))^2+sd(na.omit(FE_2_Conc_Final))),
            FE_2_flag = ifelse(mean(na.omit(Abs1)) > mean(FE_2_500uM$Abs1), "High",ifelse(mean(na.omit(Abs1)) < (mean(FE_2_0uM$Abs1)+ MDL), "Below MDL" ,"Within range")),
            FE_3_flag = ifelse(mean(na.omit(Abs2)) > mean(FE_2_100uM$Abs2)|(mean(na.omit(Abs2))-mean(na.omit(Abs1))) > mean(FE_3_100uM$Abs2), "High", ifelse(mean(na.omit(Abs2_diff)) < MDL|mean(na.omit(Abs2)) < (mean(FE_2_0uM$Abs2)+ MDL), "Below MDL" ,"Within range")),
            Dilution = first(Dilution),
            Plate = first(Plate),
            QAQC_flag = first(QAQC_flag))


#Flag high cvs 
dat1_HCV_fixed_FE2$FE_2_cv_flag <- ifelse(dat1_HCV_fixed_FE2$FE_2_cv > cv_flag_sample, "High cv rerun", "ok")
dat1_HCV_fixed_FE2$FE_3_cv_flag <- ifelse(dat1_HCV_fixed_FE2$FE_3_cv > cv_flag_sample, "High cv rerun", "ok")

#bind fixed dataframes back together with non high cv points
#averaged data
dat_aver1 <- subset(dat_aver1, !(ID_full %in% dat1_HCV_fixed_FE2$ID_full ))
dat_aver1 <- rbind(dat1_HCV_fixed_FE2, dat_aver1)
head(dat_aver1)
#all points
dat1_fixed <- subset(dat1_fixed, (!ID_full %in% dat_HCV_fixed_FE2$ID_full ))
dat1_fixed <- rbind(dat_HCV_fixed_FE2, dat1_fixed)

#make a loop to filter each high cv sample for Fe 3
dat1_HCV_1<-subset(dat1_HCV,  FE_3_cv_flag == "High cv rerun")
for(x in 1:nrow(dat1_HCV_1)){
  dat_HCV <- subset(dat1_fixed, ID_full %in% dat1_HCV_1$ID_full )
  sing_ID<- dat_HCV %>% subset(ID_full == dat1_HCV_1$ID_full[x])
  #make a dataframe for cvs
  cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
  
  #make a loop to find the cv for each deleted point
  if(nrow(sing_ID)>2){
  for(i in 1:nrow(sing_ID)){
    sing_ID1<-sing_ID[-i,] 
    cv_trial[i]<-cv(sing_ID1$FE_3_Conc_Final)
  }
  
  # delete the point that gives the lowest cv
  if(which.min(cv_trial)<cv(sing_ID$FE_3_Conc_Final)){sing_ID[which.min(cv_trial),c(4,12,17)]=NA}
  dat_HCV_fixed_FE3<-rbind(dat_HCV_fixed_FE3,sing_ID)
  }else{
    dat_HCV_fixed_FE2<-rbind(dat_HCV_fixed_FE3,sing_ID) 
  }}
#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1_HCV_fixed_FE3 <- dat_HCV_fixed_FE3 %>%
  group_by(ID_full) %>%
  summarise(IDs=first(IDs),FE_2_mean = mean(na.omit(FE_2_Conc_Final)), FE_2_sd = sd(na.omit(FE_2_Conc_Final)), FE_2_cv = cv(na.omit(FE_2_Conc_Final)),
            FE_3_mean = mean(na.omit(FE_3_Conc_Final)), FE_3_sd = sd(na.omit(FE_3_Conc_Final)), FE_3_cv = cv(na.omit(FE_3_Conc_Final)),
            FE_tot_mean =(mean(FE_2_Conc_Final,na.rm=TRUE)+ mean(FE_3_Conc_Final,na.rm=TRUE)), FE_tot_sd = sqrt(sd(na.omit(FE_3_Conc_Final))^2+sd(na.omit(FE_2_Conc_Final))),
            FE_2_flag = ifelse(mean(na.omit(Abs1)) > mean(FE_2_500uM$Abs1), "High",ifelse(mean(na.omit(Abs1)) < (mean(FE_2_0uM$Abs1)+ MDL), "Below MDL" ,"Within range")),
            FE_3_flag = ifelse(mean(na.omit(Abs2)) > mean(FE_2_100uM$Abs2)|(mean(na.omit(Abs2))-mean(na.omit(Abs1))) > mean(FE_3_100uM$Abs2), "High", ifelse(mean(na.omit(Abs2_diff)) < MDL|mean(na.omit(Abs2)) < (mean(FE_2_0uM$Abs2)+ MDL), "Below MDL" ,"Within range")),
            Dilution = first(Dilution),
            Plate = first(Plate),
            QAQC_flag = first(QAQC_flag))


#Flag high cvs
dat1_HCV_fixed_FE3$FE_2_cv_flag <- ifelse(dat1_HCV_fixed_FE3$FE_2_cv > cv_flag_sample, "High cv rerun", "ok")
dat1_HCV_fixed_FE3$FE_3_cv_flag <- ifelse(dat1_HCV_fixed_FE3$FE_3_cv > cv_flag_sample, "High cv rerun", "ok")

#bind fixed dataframes back together with non high cv points
#averaged data
dat_aver1 <- subset(dat_aver1, !(ID_full %in% dat1_HCV_fixed_FE3$ID_full ))
dat_aver1 <- rbind(dat1_HCV_fixed_FE3, dat_aver1)
head(dat_aver1)
#all points
dat1_fixed <- subset(dat1_fixed, (!ID_full %in% dat_HCV_fixed_FE3$ID_full ))
dat1_fixed <- rbind(dat_HCV_fixed_FE3, dat1_fixed)
```

## Flagged data
```{r, echo=FALSE}
#filter the high cv samples  
HighCVSamples <- subset(dat_aver1, FE_2_cv_flag == "High cv rerun"|FE_3_cv_flag == "High cv rerun")  
#filter the samples that need to be run at a higher dilution 
FE_DiluteRerun <- subset(dat_aver1, FE_2_flag == "High"| FE_3_flag == "High")
ADL<-subset(dat_aver1 ,(IDs %in% FE_DiluteRerun$IDs))
ADL<-filter(ADL,!FE_2_flag=="High" & !FE_3_flag=="High") 
FE_DiluteRerun<-FE_DiluteRerun %>% 
  subset((!IDs %in% ADL$IDs)) %>% 
  subset((!IDs %like% " dup")) %>% 
  subset((!IDs %like% " spk"))
#filter the samples that need to be run at a lower dilution 
FE_bdl <- subset(dat_aver1, FE_2_flag == "Below MDL"& Dilution!=1| FE_3_flag == "Below MDL"& Dilution!=1)
BDL<-subset(dat_aver1 ,(IDs %in% FE_bdl$IDs))
BDL<-filter(BDL, !FE_2_flag == "Below MDL" & !FE_3_flag == "Below MDL") 
FE_bdl<-FE_bdl %>% 
  subset((!IDs %in% BDL$IDs)) %>% 
  subset((!IDs %like% " dup")) %>% 
  subset((!IDs %like% " spk"))

knitr::kable(HighCVSamples[, c(1,3,5,6,8,16,17)],col.names = gsub("_", " ", names(HighCVSamples[, c(1,3,5,6,8,16,17)])), caption = "High CV Samples")
knitr::kable(FE_DiluteRerun[, c(1,3,6,9,11,12)],col.names = gsub("_", " ", names(FE_DiluteRerun[, c(1,3,6,9,11,12)])), caption = "Samples Above the Detection Limit that should be rerun at a higher dilution")
knitr::kable(FE_bdl[, c(1,3,6,9,11,12)], col.names = gsub("_", " ", names(FE_bdl[, c(1,3,6,9,11,12)])),caption = "Samples Below the Detection Limit that should be rerun at a lower dilution")

```

## Plot data after bad reps were removed 

```{r, echo=FALSE,fig.width=8,fig.height=5,fig.show='hold'}
#plot data and sd's just to check and see what they look like - just a quick first look 
FE2 <- ggplot(dat_aver1, aes(x=ID_full, y=FE_2_mean, color=FE_2_flag))+ 
  geom_point(size=4)+ 
  scale_color_manual(values = c("High" = "red", "Below MDL" = "orange", "Within range" = "black")) +  theme_classic() + 
  labs(y="Fe(II) (uM)", x="Sample ID",title=paste(Project, Experiment," Iron(II) Data")) + 
   geom_errorbar(aes(ymin=FE_2_mean-FE_2_sd,
                    ymax=FE_2_mean+FE_2_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
FE2

FE3 <- ggplot(dat_aver1, aes(x=ID_full, y=FE_3_mean, color=FE_3_flag))+ 
  geom_point(size=4)+ 
  scale_color_manual(values = c("High" = "red", "Below MDL" = "orange", "Within range" = "black")) +  theme_classic() + 
  labs(y="Fe(III) (uM)", x="Sample ID",title=paste(Project, Experiment," Iron(III) Data")) + 
   geom_errorbar(aes(ymin=FE_3_mean-FE_3_sd,
                    ymax=FE_3_mean+FE_3_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
FE3

FE_tot <- ggplot(dat_aver1, aes(x=IDs, y=FE_tot_mean, color=Dilution))+ 
  geom_point(size=4) +  theme_classic() +
  labs(y="Fe(II)+Fe(III) (uM)", x="Sample ID",title=paste(Project, Experiment," Total Iron Data")) + 
   geom_errorbar(aes(ymin=FE_tot_mean-FE_tot_sd,
                    ymax=FE_tot_mean+FE_tot_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
FE_tot

Fe2 <- ggplot(dat_aver1, aes(x=ID_full, y=FE_2_mean, color=FE_2_cv_flag))+ 
  geom_point(size=4) +  theme_classic() +  
  scale_color_manual(values = c("High cv rerun" = "red", "ok" = "darkgreen"))+
  labs(y="Fe(II) (uM)", x="Sample ID",title=paste(Project, Experiment," Iron(II) Data"))+
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_2_mean-FE_2_sd,
                    ymax=FE_2_mean+FE_2_sd),width=0.3,position=position_dodge(.1))
Fe2

Fe3 <- ggplot(dat_aver1, aes(x=ID_full, y=FE_3_mean, color=FE_3_cv_flag))+ 
  geom_point(size=4) +  theme_classic() + 
  scale_color_manual(values = c("High cv rerun" = "red", "ok" = "darkgreen"))+
  labs(y="Fe(III) (uM)", x="Sample ID",title=paste(Project, Experiment," Iron(III) Data"))+
  guides(x = guide_axis(angle = 70)) + 
  geom_errorbar(aes(ymin=FE_3_mean-FE_3_sd,
                    ymax=FE_3_mean+FE_3_sd),width=0.3,position=position_dodge(.1))
Fe3



```

## Check Dups for QAQC 
```{r Dups, echo=FALSE,fig.width=8,fig.height=5,warning=FALSE}

#pull out any rows that have "d" in the SampleID column
dups <- dat_aver1 %>%  
  filter(str_detect(IDs, "dup")) 

#remove these from dat2
dat3 <-  dat_aver1  %>%  
  filter(!str_detect(IDs, "dup")) %>%  
  filter(!str_detect(IDs, "spk")) 

#remove the d from these IDs so we will have duplicates 
dups$ID_full<-gsub(" dup",'',dups$ID_full)
dups <- dups[ ,-c(4,5,7:8,10:17)]

colnames(dups) <- c('ID_full','IDs', 'FE_2_mean_dup', "FE_3_mean_dup", "FE_tot_mean_dup")


#put it back together with the old data set and look for duplicates 
QAdups <- merge(dat3, dups, by= "ID_full")
QAdups <- QAdups[ ,-c(4:5,7:8,10:12,15:18)]

QAdups$FE_2_dups_chk <- ((abs(QAdups$FE_2_mean-QAdups$FE_2_mean_dup))/((QAdups$FE_2_mean+QAdups$FE_2_mean_dup)/2))*100
QAdups$FE_2_dups_chk <- ifelse(!is.na(QAdups$FE_2_dups_chk),QAdups$FE_2_dups_chk,0 )
QAdups$FE_2_dups_flag <-  ifelse(QAdups$FE_2_dups_chk <dups_perc_diff, 'OK', 'Rerun')

QAdups$FE_3_dups_chk <- ((abs(QAdups$FE_3_mean-QAdups$FE_3_mean_dup))/((QAdups$FE_3_mean+QAdups$FE_3_mean_dup)/2))*100
QAdups$FE_3_dups_chk <- ifelse(!is.na(QAdups$FE_3_dups_chk),QAdups$FE_3_dups_chk,0 )
QAdups$FE_3_dups_flag <-  ifelse(QAdups$FE_3_dups_chk <dups_perc_diff, 'OK', 'Rerun')

QAdups$FE_tot_dups_chk <- ((abs(QAdups$FE_tot_mean-QAdups$FE_tot_mean_dup))/((QAdups$FE_tot_mean+QAdups$FE_tot_mean_dup)/2))*100
QAdups$FE_tot_dups_chk <- ifelse(!is.na(QAdups$FE_tot_dups_chk),QAdups$FE_tot_dups_chk,0 )
QAdups$FE_tot_dups_flag <-  ifelse(QAdups$FE_tot_dups_chk <dups_perc_diff, 'OK', 'Rerun')

#plot dups output as a bar graph to easily check - want any over 10% to be red need to work on this 
dupsbar_Fe2 <- ggplot(data = QAdups, aes(x = ID_full, y = FE_2_dups_chk, fill=FE_2_dups_flag)) +
       geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK" ='darkgreen' , "Rerun" = 'darkred')) + 
        theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
        theme(legend.position="none") +  geom_hline(yintercept=dups_perc_diff, linetype="dashed", 
                color = "black", linewidth=1)

dupsbar_Fe2


dupsbar_Fe3 <- ggplot(data = QAdups, aes(x = ID_full, y = FE_3_dups_chk, fill=FE_3_dups_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values=c("OK" ='darkgreen' , "Rerun" = 'darkred')) + 
        theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
        theme(legend.position="none") +  geom_hline(yintercept=dups_perc_diff, linetype="dashed", 
                color = "black", linewidth=1)

dupsbar_Fe3

dupsbar_Fetot <- ggplot(data = QAdups, aes(x = ID_full, y = FE_tot_dups_chk, fill=FE_tot_dups_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values=c("OK" ='darkgreen' , "Rerun" = 'darkred')) + 
        theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
        theme(legend.position="none") +  geom_hline(yintercept=dups_perc_diff, linetype="dashed", 
                color = "black", linewidth=1)

dupsbar_Fetot
#check for any no's that would warrant reruns
Baddups_Fe2 <- subset(QAdups, FE_2_dups_flag == "Rerun")
Baddups_Fe3 <- subset(QAdups, FE_3_dups_flag == "Rerun")



dat_aver1<-dat_aver1 %>% group_by(Plate) %>% mutate(QAQC_flag=ifelse(Plate %in% Baddups_Fe2$Plate, ifelse(dat_aver1$QAQC_flag != "", paste0(dat_aver1$QAQC_flag, "; Fe2 Dup perc diff out of range"),"Fe2 Dup perc diff out of range"),paste0( dat_aver1$QAQC_flag,"") ) )%>% mutate(QAQC_flag=ifelse(Plate %in% Baddups_Fe3$Plate, ifelse(dat_aver1$QAQC_flag != "", paste0(dat_aver1$QAQC_flag, "; Fe3 Dup perc diff out of range"),"Fe3 Dup perc diff out of range"),paste0( dat_aver1$QAQC_flag,"") ) ) %>% ungroup()

dat1<-dat1 %>% group_by(Plate) %>% mutate(QAQC_flag=ifelse(Plate %in% Baddups_Fe2$Plate, ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, "; Fe2 Dup perc diff out of range"),"Fe2 Dup perc diff out of range"),paste0( dat1$QAQC_flag,"") ) )%>% mutate(QAQC_flag=ifelse(Plate %in% Baddups_Fe3$Plate, ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, "; Fe3 Dup perc diff out of range"),"Fe3 Dup perc diff out of range"),paste0( dat1$QAQC_flag,"") ) ) %>% ungroup()


dups_percent_Fe2 <- (sum(QAdups$FE_2_dups_flag == "OK")/nrow(QAdups))*100
#report out if flags indicate need for rerun
ifelse(dups_percent_Fe2 >= 60  , ">60% of Fe 2 Duplicates are within <10%",
       "<60% of Duplicates are within <10% - REASSESS")

dups_percent_Fe3 <- (sum( QAdups$FE_3_dups_flag == "OK")/nrow(QAdups))*100
#report out if flags indicate need for rerun
ifelse(dups_percent_Fe3 >= 60  , ">60% of Fe 3 Duplicates are within <10%",
       "<60% of Duplicates are within <10% - REASSESS")


```


## Check the spks for QAQC 
```{r Spikes, echo=FALSE,fig.width=8,fig.height=5,warning=FALSE}

#pull out any rows that have "spk" in the SampleID column
spks <- dat_aver1 %>%  
  filter(str_detect(IDs, "spk")) 

#remove the spk from these IDs so we will have duplicates 
spks$ID_full<-gsub(" spk","",spks$ID_full)
spks <- spks[ ,-c(4:14)]
colnames(spks) <- c('ID_full','IDs', 'FE_2_mean_spk')

#put it back together with the old data set and look for duplicates 
QAspks <- merge(dat3, spks, by= "ID_full")
QAspks <- QAspks[ ,-c(4:5,7:10,15)]

#now we need to calculate the spike concentration and calculate the spike recovery 
#spike for these samples was 55uL of the 60uM standard
                      
QAspks$FE_Total_unspkd <- (QAspks$FE_2_mean/QAspks$Dilution)*(sample_vol/Con1) #gives us the total fe in the sample in umoles
QAspks$FE_Total_spkd <-    (QAspks$FE_2_mean_spk/QAspks$Dilution)*((sample_vol+spkvol)/Con1)        ##total fe in spiked sample in umoles 
QAspks$FE_expctd_spkd <-  (QAspks$FE_Total_unspkd + spk_Conc)
QAspks$spk_recovery <-    (QAspks$FE_Total_spkd/QAspks$FE_expctd_spkd)*100
QAspks$FE_spks_flag <-  ifelse(QAspks$spk_recovery >low_recovery_cutoff & QAspks$spk_recovery < high_recovery_cutoff, 'OK', 'NO, rerun') 


#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 

spksbar <- ggplot(data = QAspks, aes(x = ID_full, y = spk_recovery, fill=FE_spks_flag)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK"="darkgreen","NO, rerun"="darkred")) + 
        theme_classic() + labs(x= "Sample ID", y="Spike Recovery (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=low_recovery_cutoff, linetype="dashed", color = "black", linewidth=1) + 
        geom_hline(yintercept=high_recovery_cutoff, linetype="dashed", color = "black", linewidth=1)

spksbar

#check for any no's that would warrant reruns!
Badspks <- subset(QAspks, FE_spks_flag == "NO, rerun")

dat_aver1<-dat_aver1 %>% group_by(Plate) %>% mutate(QAQC_flag=ifelse(Plate %in% Badspks$Plate, ifelse(dat_aver1$QAQC_flag != "", paste0(dat_aver1$QAQC_flag, "; Spk recovery out of range"),"Spk recovery out of range"),paste0( dat_aver1$QAQC_flag,"") ) ) %>% ungroup()

dat_aver1 <- dat_aver1 %>%
  mutate(Run_notes = Run_notes)

dat1<-dat1 %>% group_by(Plate) %>% mutate(QAQC_flag=ifelse(Plate %in% Baddups_Fe2$Plate, ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, "; Spk recovery out of range"),"Spk recovery out of range"),paste0( dat1$QAQC_flag,"") ) ) %>% ungroup()

dat1 <- dat1 %>%
  mutate(Run_notes = Run_notes)

#write out a flag to the sample dataframe if any spks have recovery out of range 


spks_percent <- (sum(QAspks$FE_spks_flag == "OK")/nrow(QAspks))*100
#report out if flags indicate need for rerun
ifelse(spks_percent >= 60  , ">60% of Spikes are within range",
       "<60% of Spikes are out of range - REASSESS")
```
## Export full data then just final data 
```{r, echo=FALSE}

#Read out the summarized data
write.csv(dat_aver1, file= summarizeddata_path)

#Read out all the data
write.csv(dat1, file= fulldata_path)

#Now take out the absorbance and stuff 
dat4 <- dat1[ ,-c(1:6,8:15)]
write.csv(dat4, file= shortdata_path)

Samples_outof_range<-rbind(FE_DiluteRerun,FE_bdl)
write.csv(Samples_outof_range, out_of_Range)
write.csv(HighCVSamples, high_CV)

#save to QAQC Data to QAQC File
write.csv(qlog, file= log_path, row.names=FALSE)

```

