---
title: "COMPASS_Synoptic_SmartChamberFluxes_2023_CO2"
author: "Stephanie J. Wilson"
date: "2025-05-08"
output: html_document
---
## Goals of this code: 
1.	Reads in raw smart chamber
2.	If no metadatafile, write one out & fields will be: 
Label 
the IDs pulled from the label, 
Plot *fixed so that if there are any issues in the label they are good here 
Drop column – false as long as label looks right 
adjust time – would be initially populated by the start time in the OR blank 
adjust observation length – initially populated by the obs length  OR blank 
notes / comment 
3.	Code reads in the meta data and applies it to the data 
a.	i.e. drops fluxes that have the drop column TRUE or FLAG this 
b.	adjusting times as needed 
4.	Now you compute fluxes 
5.	Visualize and look for issues 
6.	If found, report to meta data file then go back to 3 and reanalyze 
7.	Apply algorithmic qaqc and generate flags 


## Set Up & Load Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(stringr)

```

## Pull in Smart Chamber Files from the Raw Data folder
```{r}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
zip_contents <- unzip("Raw Data.zip", list = TRUE)

# Keep only JSON files
json_files <- zip_contents$Name[grepl("json$", zip_contents$Name, ignore.case = TRUE)]

# Extract only the JSON files
unzip(
  "Raw Data.zip",
  files = json_files,
  exdir = tmp_dir
)

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "json$",
  full.names = TRUE,
  recursive = TRUE
)

head(file.list)

# Read in the data
alldat1 <- lapply(file.list, ffi_read_LIsmartchamber)

# Combine into one data.table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)


```

## Create a unique Plot_ID, check if there is a meta-data file, if not, create a meta-data file
```{r}

############### only run this first part if you don't already have a meta-data file 
        #because if you do then all the data for each one will have a plot and you won't be able to change start times


#create unique Plot ID for each measurement: 
#detach("package:data.table", unload=TRUE)
alldat$month <- as.character(month(alldat$TIMESTAMP, label=TRUE, abbr=TRUE)) #will not run if you have data.table loaded 

alldat$year <- year(alldat$TIMESTAMP)

alldat <- alldat %>% 
  mutate(Plot_ID = str_c(pmin(month, RepNum), 
                      pmax(month, RepNum),  sep = '-'))
alldat <- alldat %>% 
  mutate(Plot_ID = str_c(pmin(Plot_ID, label), 
                     pmax(Plot_ID, label),  sep = '-'))  
#this needs to be PLot ID so that later we ID plot from metadata and not the metadata row 

# Path inside the zip
zip_file <- "Raw Data.zip"
csv_inside_zip <- "Raw Data/COMPASS_SmartChamber_Metadata_2023.csv"

# Make sure the destination folder exists
dir.create("Raw Data", showWarnings = FALSE)

# Extract only the metadata CSV if it doesn't already exist
if (!file.exists(csv_inside_zip)) {
  unzip(
    zip_file,
    files = csv_inside_zip,
    exdir = "."
  )
}

# Define the file path where you want to save the CSV file
file_path <- "Raw Data/COMPASS_SmartChamber_Metadata_2023.csv"

# Check if the CSV file exists
if (file.exists(file_path)) {
  # If it exists, read it back into R
  metadata <- read.csv(file_path)
  print("CSV file exists and has been read into the code.")
} else {
  # If it does not exist, create the CSV file
    #create a metadata file from the Plot IDs 
      # pull timestart timestamps to put in the metadata and obs lengths 
  metadata <- alldat %>%
  group_by(Plot_ID) %>%
  summarise(
    Date = format(min(TIMESTAMP), "%Y-%m-%d"),
    Time_start = format(min(TIMESTAMP), "%H:%M:%S"),
    First_Timestamp = min(TIMESTAMP),  # Minimum (earliest) timestamp for each Plot_ID
    Last_Timestamp = max(TIMESTAMP),   # Maximum (latest) timestamp for each Plot_ID
    Total_Time = as.numeric(difftime(max(TIMESTAMP), min(TIMESTAMP), units = "secs")),  # Time difference in seconds
    ) %>%
  ungroup() %>%
  mutate(drop = FALSE) %>% # Add a 'drop' column with 'FALSE' for every row 
  #mutate(notes = " ")  # Add a 'drop' column with 'FALSE' for every row
  #mutate(Deadband = 5)
    mutate(Notes = "") #%>%
    #mutate(Plot = Plot_ID) #think need this because we need to make a plot column
  
  # Write df_unique to CSV
  write.csv(metadata, file = file_path, row.names = FALSE)
  
  print("CSV file does not exist. It has been created.")
}


```

## Mark fluxes that were retaken - NOT RUNNING FOR NOW   - this is the first issue! Need to get all but last flux to be TRUE for drop  
```{#r}

#so first we need to make it so that the fluxes that were retaken have similar names or some way to identify them 
  # if they were retaken it will either have a letter after it or a retake written 

#change drop column to TRUE if the plot doesn't include these strings 
metadata <- metadata %>%
  mutate(
    drop = ifelse(
      !grepl("SWH|GCW|MSM|GWI|GCReW", Plot), 
      # Check if Plot_ID does not contain the specified strings
      TRUE,  # Set drop to TRUE if it doesn't contain any of the strings
      drop   # Keep the original value of drop if it contains one of the strings
    )
  )

#if they say retake then we want to replace that with an "a" and if it has -A then we also want to remove that and make it "a" 
metadata <- metadata %>%
  mutate(Plot = if_else(str_detect(Plot, "-retake"), 
                        str_replace(Plot, "-retake", "a"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-Retake"), 
                        str_replace(Plot, "-Retake", "a"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-b"), 
                        str_replace(Plot, "-b", "b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "1-May-GCW-TR-3-A"), 
                        str_replace(Plot, "1-May-GCW-TR-3-A", "1-May-GCW-TR-3a"), 
                        Plot)) 

#now need to figure out if the flux was taken again and use the latest measurement 

# Extract the base plot (e.g., "GCW-TR-1" from "GCW-TR-1", "GCW-TR-1a", etc.)
metadata$base_plot <- str_extract(metadata$Plot,  "^[^-]+-[A-Za-z]+-[A-Za-z]+-[A-Za-z]+")

# Extract the replicate number (the number after the last hyphen)
metadata$replicate <- as.numeric(str_extract(metadata$Plot, "(?<=-)[0-9]+(?=[a-zA-Z]?$)"))

# Extract the suffix (if any, the letter after the replicate number)
metadata$suffix <- str_extract(metadata$Plot, "(?<=\\d)[a-zA-Z]$")

# If there's no suffix, set it to an empty string
metadata$suffix[is.na(metadata$suffix)] <- ""

# Sort by base plot, replicate number, and suffix (to keep highest suffix) - #####INSTEAD WE WANT TO CHANGE DROP TO TRUE HERE#######
#allflux <- metadata %>%
 # group_by(base_plot, replicate) %>%
#  arrange(base_plot, replicate, desc(suffix), .by_group = TRUE) %>%
 # slice_head(n = 1) %>%
#  ungroup()



```

##Dont need to pull L1 temp data or any other temps like we do for macrophyte because the smartchamber records the temp

## Match up the metadata to the rest of the data   
```{r}

#remove any metadata rows that have NAs for the PlotID
metadata <- metadata %>%
  filter(!is.na(Plot_ID)) %>%
  filter(str_detect(Plot_ID, paste(c("GCW", "MSM", "GWI", "SWH"), collapse = "|"))) %>%
  filter(!str_detect(Plot_ID, paste(c("1-Nov-MSM-UP-2"), collapse = "|"))) # This one has a weird Obs Length need to flag and remove

#match the altered metadata to the concentration data with fluxfinder ffi_metadata_match
alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = alldat$TIMESTAMP,
  start_dates = metadata$Date,
  start_times = metadata$Time_start,
  obs_lengths = metadata$Total_Time) 
head(alldat)

#alldat$Deadband <- metadata$Deadband[alldat$metadat_row] #flux calcs function cannot handle a column as a deadband and that is a problem to fix
alldat$Plot <- metadata$Plot_ID[alldat$metadat_row] 

#alldat$Drop <- metadata$Drop[alldat$metadat_row] #don't need this for now because we remove the fluxes that were retaken below 

```

## Change the units from ppm to umol  
```{r}
#change the units
alldat$CO2_umol <- ffi_ppm_to_umol(alldat$co2, 
                                volume = (alldat$TotalVolume / 1000000), # have to divide to get to m3
                                temp = alldat$chamber_t)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CO2_umol_m2 <- alldat$CO2_umol / 0.0318 #adjust for the area 


```

## Calculate fluxes
```{r}

#need to create a new line that would say only calculate the flux if the drop column is false
    #before we add this we need to figure out how to drop except for the last flux taken 


#calculate fluxes for all other sites 
fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CO2_umol_m2", 
                             dead_band = 5)
head(fluxes)

allflux <- fluxes


```

## Remove fluxes that were retaken - this works, but I think we want to do this earlier with the drop column???   
```{r}

#so first we need to make it so that the fluxes that were retaken have similar names or some way to identify them 
  # if they were retaken it will either have a letter after it or a retake written 


#remove any lines that don't have a site code in the Plot ID 
fluxes <-  fluxes %>%  
  filter(str_detect(Plot, "SWH|MSM|GWI|GCW|GCReW"))

#if they say retake then we want to replace that with an "a" and if it has -A then we also want to remove that and make it "a" 
fluxes <- fluxes %>%
  mutate(Plot = if_else(str_detect(Plot, "-retake"), 
                        str_replace(Plot, "-retake", "a"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "real"), 
                        str_replace(Plot, "real", "b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-Retake"), 
                        str_replace(Plot, "-Retake", "a"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-b"), 
                        str_replace(Plot, "-b", "b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "1A"), 
                       str_replace(Plot, "1A", "1a"), 
                        Plot))%>%
  mutate(Plot = if_else(str_detect(Plot, "1B"), 
                       str_replace(Plot, "1B", "1b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "2A"), 
                       str_replace(Plot, "2A", "2a"), 
                        Plot))%>%
  mutate(Plot = if_else(str_detect(Plot, "2B"), 
                       str_replace(Plot, "2B", "2b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "2C"), 
                       str_replace(Plot, "2C", "2c"), 
                        Plot))%>%
  mutate(Plot = if_else(str_detect(Plot, "1real"), 
                        str_replace(Plot, "1real", "1b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-1B"), 
                       str_replace(Plot, "-1B", "-1b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-4B"), 
                       str_replace(Plot, "-4B", "-4b"), 
                        Plot))

#Do the same for alldat (individual flux point data)
#if they say retake then we want to replace that with an "a" and if it has -A then we also want to remove that and make it "a" 
#remove any lines that don't have a site code in the Plot ID 
alldat <-  alldat %>%  
  filter(str_detect(Plot, "SWH|MSM|GWI|GCW|GCReW"))

alldat <- alldat %>%
  mutate(Plot = if_else(str_detect(Plot, "1A"), 
                       str_replace(Plot, "1A", "1a"), 
                        Plot))%>%
  mutate(Plot = if_else(str_detect(Plot, "1B"), 
                       str_replace(Plot, "1B", "1b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "2A"), 
                       str_replace(Plot, "2A", "2a"), 
                        Plot))%>%
  mutate(Plot = if_else(str_detect(Plot, "2B"), 
                       str_replace(Plot, "2B", "2b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "2C"), 
                       str_replace(Plot, "2C", "2c"), 
                        Plot))%>%
  mutate(Plot = if_else(str_detect(Plot, "1real"), 
                        str_replace(Plot, "1real", "1b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-1B"), 
                       str_replace(Plot, "-1B", "-1b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-4B"), 
                       str_replace(Plot, "-4B", "-4b"), 
                        Plot)) %>%
  mutate(Plot = if_else(str_detect(Plot, "-b"), 
                        str_replace(Plot, "-b", "b"), 
                        Plot))

# Extract the base plot (e.g., "GCW-TR-1" from "GCW-TR-1", "GCW-TR-1a", etc.)
fluxes$base_plot <- str_extract(fluxes$Plot,  "^[^-]+-[A-Za-z]+-[A-Za-z]+-[A-Za-z]+")

# Extract the replicate number (the number after the last hyphen)
fluxes$replicate <- as.numeric(str_extract(fluxes$Plot, "(?<=-)[0-9]+(?=[a-zA-Z]?$)"))

# Extract the suffix (if any, the letter after the replicate number)
fluxes$suffix <- str_extract(fluxes$Plot, "(?<=\\d)[a-zA-Z]$")

# If there's no suffix, set it to an empty string
fluxes$suffix[is.na(fluxes$suffix)] <- ""

#create ID columns from the plots 
IDs <- data.frame(do.call('rbind', strsplit(as.character(fluxes$Plot),'-',fixed=TRUE)))
colnames(IDs) <- c("RepNumber", "Month", "Site", "Zone", "Replicate") 
head(IDs)

#combine these dataframes together 
allflux <- cbind(IDs, fluxes)

# Sort by base plot, replicate number, and suffix (to keep highest suffix)
allflux <- allflux %>%
  group_by(base_plot, replicate) %>%
  arrange(Month, Site, Zone, replicate, desc(suffix), .by_group = TRUE) %>%
  slice_head(n = 1) %>%
  ungroup()



```

## Visualize fluxes and sort by r2 
```{r}

#take a look at the fluxes we calculated with R2 
ggplot(allflux, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + ylim(-5,100) + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CO2 flux (mmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(allflux, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate > 1, select = Plot:TIMESTAMP_max))
fluxes_negative <- as.data.frame(subset(allflux,  lin_flux.estimate < 0, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(allflux$lin_flux.estimate, 0.05) 
high <- quantile(allflux$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(allflux, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(allflux, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r}

tobechecked <- c("2-Aug-SWH-TR-9" , "2-Jul-MSM-TR-1"      )

colnames <- c("Plot")

#Loop to run five random plots subset above 
for(P in tobechecked){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), co2)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

#r1 <- subset(alldat, Plot ==  "2-May-GWI-TR-4b")
#p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), ch4)) + geom_point() + 
 # scale_x_datetime(breaks = "15 sec") + 
 # theme_classic() +
 # ylab(expression(paste( CH [4], " (ppb)"))) +
 # xlab("Time") + labs(title = "plot name above") + 
  #geom_smooth()
#print(p1)


#list of the ones that I changed and checked and looked good? 
fixedplots <- c("2-Jul-GWI-UP-3", "2-Aug-SWH-TR-6a", "1-Oct-SWH-TR-9", "1-Sep-SWH-TR-9",
                "2-May-GWI-TR-4b", "1-Sep-SWH-UP-2", "2-Oct-SWH-TR-1a", "2-Oct-SWH-UP-2", "2-Nov-SWH-SWAMP-2a",
                "2-Oct-GWI-UP-3", "1-Aug-SWH-TR-1", "2-May-SWH-SWAMP-2", "2-Nov-SWH-SWAMP-1", "2-Aug-SWH-TR-1",
                 "2-Aug-SWH-SWAMP-2", "1-Sep-SWH-TR-8b", "1-Oct-SWH-TR-10", "1-Oct-SWH-SWAMP-1a", "1-Nov-SWH-TR-4a", 
                "1-Nov-SWH-SWAMP-2a", "1-Nov-SWH-SWAMP-1",  "1-May-SWH-TR-4a", "1-May-SWH-SWAMP-2", "1-May-GCW-TR-4", 
                "1-Jun-SWH-TR-8b", "2-Jun-SWH-TR-3a", "1-Aug-SWH-TR-8a", "1-Aug-SWH-SWAMP-2", "2-May-GWI-UP-5", "1-Nov-SWH-UP-2",
                "2-Aug-SWH-TR-2c", "1-Oct-SWH-SWAMP-3a", "1-Jun-SWH-UP-4", "1-May-SWH-TR-5a", 
                 "2-Nov-SWH-SWAMP-5a", "2-Aug-SWH-TR-9a", "2-Sep-GCW-TR-5", "2-Aug-SWH-SWAMP-1a", "2-Oct-SWH-SWAMP-2", 
                "1-Aug-SWH-TR-2c", "2-Oct-GCW-TR-4", "2-Aug-SWH-SWAMP-1a", "1-Nov-SWH-UP-4", 
                "1-Jun-SWH-TR-5" , "1-Aug-SWH-TR-5", "2-Oct-GCW-UP-4", "1-Jul-SWH-UP-3", "1-Oct-SWH-SWAMP-4", "2-Sep-SWH-TR-2a", 
                 "2-Jun-SWH-TR-1a", "2-Jun-SWH-SWAMP-1", "2-Jul-SWH-SWAMP-2a", "2-Jul-SWH-SWAMP-1a", "1-Oct-SWH-TR-2a", "2-Sep-GCW-UP-5", 
                 "2-Jun-GCW-UP-5", "2-Aug-GCW-UP-5", "2-Jul-GCW-UP-1", "2-Oct-SWH-TR-3a", "1-Aug-SWH-TR-7a", "2-Nov-SWH-UP-1", "2-Sep-SWH-UP-1",
                "2-Jun-GCW-TR-3", "1-Sep-SWH-TR-3a", "2-Sep-SWH-TR-5", "2-Sep-GWI-UP-4", "2-Oct-SWH-SWAMP-1a", "2-May-GWI-TR-1b", "2-Jul-GCW-TR-2", 
                 "2-Jun-GCW-UP-3", "1-Jul-SWH-UP-2", "2-Jun-GCW-TR-1", "1-Nov-SWH-TR-7a", "2-Nov-SWH-TR-10", "2-Nov-SWH-SWAMP-4",
                 "1-Nov-SWH-TR-8", "1-Apr-MSM-UP-3", "1-Sep-SWH-SWAMP-3", "1-Apr-MSM-TR-1", "2-Jun-SWH-UP-5", "1-Oct-SWH-SWAMP-5a", 
                "2-Sep-SWH-TR-8" , "1-Sep-SWH-TR-7a", "1-Aug-SWH-TR-6" , "2-Jul-MSM-TR-1", "2-Aug-SWH-TR-9")

Missing_data <- c("1-Nov-SWH-UP-1b", "2-Nov-SWH-UP-1", "1-Apr-MSM-UP-4B", "2-Apr-MSM-UP-4b", "1-Jul-SWH-TR-1b", "2-Jul-SWH-TR-1b",  "1-Nov-SWH-UPCON-1b",   "2-Nov-SWH-UP-1b", "1-Nov-SWH-UP-2c", "2-Nov-SWH-UP-2c")

```


## Spot check a few fluxes with R2 > 0.9  
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), co2)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

####################
```

## Take a look at the  fluxes <0.9 and a flux estimate below 1
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot)

#Loop to run five random plots subset above 
for(P in low_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO2 [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

####################

```

## Take a look at the bad fluxes <0.9 and have a high flux estimate 
# See if we want to remove these fluxes or if we want to try and recalculate these ones? 
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

#Loop to run five random plots subset above 
for(P in bad_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################


```

# Look at any CO2 fluxes that are negative and remove those plots 
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
neg_plots <- unique(fluxes_negative$Plot)

#Loop to run five random plots subset above 
for(P in neg_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################


```


## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

#Loop to run five random plots subset above 
for(P in tail_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Check all fluxes
```{r, echo=FALSE}

##Put all fluxes together
fluxes_all <- as.data.frame(subset(allflux, select = Plot:TIMESTAMP_max))
all_plots <- unique(fluxes_all$Plot)

#Loop to run five random plots subset above
for(P in all_plots){
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################
```


## Remove any fluxes that do not fit criteria 
```{r, echo=FALSE}

# Create a new dataframe by removing rows that were present in fluxes_weird which have a low R2 and a flux higher than 1 (23 obs) and negative CO2 fluxes
allflux_filtered <- allflux[!(allflux$Plot %in% fluxes_weird$Plot), ]

allflux_filtered <- allflux_filtered[!(allflux_filtered$Plot %in% fluxes_negative$Plot), ]

#Any others that are not in the weird category that need to be removed: 
  #also removing ones that are crazy off but were not in the fluxes_weird dataframe (did this by listing them below)

allflux_filtered <- allflux_filtered %>%
  filter(!Plot %in% c("2-Jul-SWH-TR-9", "1-Jul-SWH-TR-9", "1-May-GCW-TR-3", "2-May-GCW-TR-3", "1-Aug-SWH-TR-10a", "1-Sep-GCW-UP-3", "2-Jul-SWH-TR-9")) 

# 1-Aug-SWH-TR-10a, V-shaped
# 1-Sep-GCW-UP-3, V-shaped
# 2-Jul-SWH-TR-9, has a starting CO2 value of over 30,000

##Ask Steph
# 1-May-SWH-UP-5, asked Steph
# 1-Sep-MSM-TR-1, asked Steph
# 2-Oct-GCW-TR-3, asked Steph
# 2-Sep-SWH-TR-1, asked Steph
# 2-Sep-SWH-TR-3, asked Steph
# 2-Sep-SWH-UP-1, asked Steph


# View the result
head(allflux_filtered)


```


## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(allflux_filtered, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(allflux_filtered, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(allflux_filtered, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere 
```{r, echo=FALSE}

#can remove base_plot column, the replicate column, and the suffix column
allflux_filtered <- allflux_filtered[,-c(31:33)]

#need to deal with SWH UPCON, UP, and SWAMP 
  #in 2023 when sampling SWH the UP site was known as UPCON and the SWAMP site was UP 
  #this was changed in 2024, so for consistency we need to change these here to prevent confusion 
allflux_filtered <- allflux_filtered %>%
  mutate(
    Zone = case_when(
      Site == "SWH" & Zone == "UP" ~ "SWAMP",
      Site == "SWH" & Zone == "UPCON" ~ "UP",
      TRUE ~ Zone )) %>%  # leave other values unchanged 
  mutate(
    Plot = case_when(
      Site == "SWH" ~ str_replace_all(Plot, c( "-UP-" = "-SWAMP-",  "-UPCON-" = "-UP-")),
      TRUE ~ Plot) ) 

#export as a csv to the folder that you want 
write.csv(allflux_filtered, file="Processed Data/COMPASS_Synoptic_CB_SmartChamber_Fluxes_Final_2023_CO2.csv")           #Change file name


```



