---
title: "COMPASS_FE_20221220"
author: "Stephanie J. Wilson"
date: '2023-01-10'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

#notes I don't have a system yet to compare old slopes to current slopes for that day 
#I don't have a system for eliminating wells if the sd is high - if you think one of your sample wells was off you would need to remove it manually... 

#Files Required: 
#csv of standards as list not plate with columns: Fe(II), Abs,	Curve,	FE ; these are: Known std concentration, Absorbance, Curve type (High or Low), and FE Curve (FE_2 or FE_Total)
#csv of samples as list not plate with columns: Date,	Plate,	Well_ID,	SampleID,	Dil_Factor,	Abs1,	Abs2 ; these are Date of run, Plate ID, Well ID, Sample ID, Dilution factor, Absorbance 1 (Fe2 or first read), and Absorbance 2 (feTot or second read)
# I save these files in a folder for each day of FE runs 


## Read in Standards and subset 
```{r}

stds <- read.csv("Raw Data/FE_20230110_Stds.csv")
head(stds)

stdid <- read.csv("Raw Data/Stds_IDs_20230110.csv")
head(stdid)

stds1 <- merge(stdid, stds)
head(stds1)

#subset by FE_2 and FE_Total
FE_2_stds <- stds1[stds1$Fe_2_High == "Yes", ]                               # One factor level
head(FE_2_stds) 

FE_Tot_stds <- stds1[stds1$Fe_3 == "Yes", ]                               # One factor level
head(FE_Tot_stds)

#subset curves for high and low 
FE_2_stds_high <-  FE_2_stds[FE_2_stds$Fe_2_High == "Yes", ]                               # One factor level
head(FE_2_stds_high) 

FE_2_stds_low <-  FE_2_stds[FE_2_stds$Fe_2_Low == "Yes", ]                               # One factor level
head(FE_2_stds_low)

#FE_Tot_stds_high <-  FE_Tot_stds[FE_Tot_stds$Curve == "High", ]                               # One factor level
#head(FE_Tot_stds_high) 

#FE_Tot_stds_low <-  FE_Tot_stds[FE_Tot_stds$Curve == "Low", ]                               # One factor level
#head(FE_Tot_stds_low)

```

## Plot standards 
```{r}
#Plot each of these and calculate the slope, intercept, and R2 
library(ggplot2)

#Low curve for Fe(II)
FE_2_Low <- ggplot(FE_2_stds_low, aes(Concentration, Abs1)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
FE_2_Low

FE_2_Low_lm <- lm(FE_2_stds_low$Abs1 ~ FE_2_stds_low$Concentration)
summary(FE_2_Low_lm)
cf <- coef(FE_2_Low_lm)

#create data frame with 1 rows and 0 columns
Slopes1 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes1$Curve <- "FE_2_Low"
Slopes1$R2 <- summary(FE_2_Low_lm)$adj.r.squared
Slopes1$Slope <- cf[2]
Slopes1$Intercept <- cf[1]
head(Slopes1)


#High curve for Fe(III)
FE_2_High <- ggplot(FE_2_stds_high, aes(Concentration, Abs1)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
FE_2_High

FE_2_High_lm <- lm(FE_2_stds_high$Abs1 ~ FE_2_stds_high$Concentration)
summary(FE_2_High_lm)
cf <- coef(FE_2_High_lm)

#create data frame with 1 rows and 0 columns
Slopes2 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes2$Curve <- "FE_2_High"
Slopes2$R2 <- summary(FE_2_High_lm)$adj.r.squared
Slopes2$Slope <- cf[2]
Slopes2$Intercept <- cf[1]
head(Slopes2)


#Low curve for Fe(Total)
FE_Tot_Low <- ggplot(FE_2_stds_low, aes(Concentration, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
FE_Tot_Low

FE_Tot_Low_lm <- lm(FE_2_stds_low$Abs2 ~ FE_2_stds_low$Concentration)
summary(FE_Tot_Low_lm)
cf <- coef(FE_Tot_Low_lm)

#create data frame with 1 rows and 0 columns
Slopes3 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes3$Curve <- "FE_Tot_Low"
Slopes3$R2 <- summary(FE_Tot_Low_lm)$adj.r.squared
Slopes3$Slope <- cf[2]
Slopes3$Intercept <- cf[1]
head(Slopes3)


#High curve for Fe(III)
FE_Tot_High <- ggplot(FE_2_stds_high, aes(Concentration, Abs2)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
FE_Tot_High

FE_Tot_High_lm <- lm(FE_2_stds_high$Abs2 ~ FE_2_stds_high$Concentration)
summary(FE_Tot_High_lm)
cf <- coef(FE_Tot_High_lm)

#create data frame with 1 rows and 0 columns
Slopes4 <- data.frame(matrix(ncol = 0, nrow = 1))
Slopes4$Curve <- "FE_Tot_High"
Slopes4$R2 <- summary(FE_Tot_High_lm)$adj.r.squared
Slopes4$Slope <- cf[2]
Slopes4$Intercept <- cf[1]
head(Slopes4)


#pull all the slopes together into one dataframe 
Slopes <- rbind(Slopes1, Slopes2, Slopes3, Slopes4)
head(Slopes)

####
```

## Read in Sample Data subset 
```{r}

dat <- read.csv("Raw Data/FE_20230110_Samples.csv")
head(dat)

## remove any lines that say stds in SampleID
library(dplyr)

dat <- dat %>%  filter(!SampleID =='stds')
head(dat)

#Now flag any Abs1 that are above the 60uM standard absorbance (0.2)
dat$FE_2_Curve <- ifelse(dat$Abs1 >0.2, "High", "Low")
head(dat)

#Now flag any Abs2 that are above the 60uM standard absorbance (0.3)
dat$FE_Tot_Curve <- ifelse(dat$Abs2 >0.3, "High", "Low")
head(dat)


```


## Subset Data and Calculate Concentrations 
```{r}
#Calculate concentrations of Fe(II)
dat$FE_2_Conc <- ifelse(dat$Abs1 >0.2, (dat$Abs1-Slopes2$Intercept)/Slopes2$Slope, (dat$Abs1-Slopes1$Intercept)/Slopes1$Slope )
head(dat)

#calculate concentrations of Fe(total)
dat$FE_Tot_Conc <- ifelse(dat$Abs2 >0.3, (dat$Abs2-Slopes4$Intercept)/Slopes4$Slope,                                       (dat$Abs2-Slopes3$Intercept)/Slopes3$Slope )
head(dat)


#Account for the dilution factor
dat$FE_2_Conc_DilCorr <- dat$FE_2_Conc*dat$dilution.factor

dat$FE_Tot_Conc_DilCorr <- dat$FE_Tot_Conc*dat$dilution.factor

head(dat)
```


## Calculate Fe(III) 
```{r}
#Calculate concentrations of Fe(III)
dat$FE_3_Conc <- dat$FE_Tot_Conc_DilCorr-dat$FE_2_Conc_DilCorr

#Use ifelse to make any negative values equal to zero 
dat$FE_2_Conc_Final <- ifelse(dat$FE_2_Conc_DilCorr <0, 0, dat$FE_2_Conc_DilCorr)

dat$FE_3_Conc_Final <- ifelse(dat$FE_3_Conc <0, 0, dat$FE_3_Conc)

#Use ifelse to make any negative values equal to zero 
dat$FE_2_Info <- ifelse(dat$FE_2_Conc_DilCorr <0, 'bdl', 'within range')

#Use ifelse to make any negative values equal to zero 
dat$FE_3_Info <- ifelse(dat$FE_3_Conc <0, 'bdl', 'within range')

head(dat)
```


## Calculate Averages across wells and std. dev.  
```{r}
head(dat)

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1 <- dat %>%
  group_by(SampleID) %>%
  summarise(FE_2_mean = mean(FE_2_Conc_Final), FE_2_sd = sd(FE_2_Conc_Final), 
            FE_3_mean = mean(FE_3_Conc_Final), FE_3_sd = sd(FE_3_Conc_Final), 
            FE_2_flag = first(FE_2_Info), FE_3_flag = first(FE_3_Info))

head(dat1)


```


## Export full data then just final data 
```{r}

#Read in meta-data and sample IDs
IDs <- read.csv("Raw Data/FE_SampleIDs_2022.csv")
head(IDs)

dat_IDs <- merge(IDs, dat1)
head(dat_IDs)

#Read out the summarized data
head(dat_IDs)
write.csv(dat_IDs, file="Processed Data/FE_20230110_Summary_Data.csv")

#Read out all the data in dat 
head(dat)
write.csv(dat, file="Processed Data/FE_20230110_Data.csv")


#Now take out the absorbance and stuff 
head(dat)
dat2 <- dat[ ,-(2:3)]
dat2 <- dat2[ ,-(3:12)]
head(dat2)
write.csv(dat2, file="Processed Data/FE_20230110_short_Data.csv")

```

### END



