---
title: "Synoptic Trees"
author: "A. Stearns"
date: "2026-03-29"
output: html_document
---
# Set up Library

```{r}

library(readr)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(zoo)
theme_set(theme_bw())
library(terra)
library(scales)

```

## Read in dataset

```{r}

trees <- read_csv("CB_TreePlotInventory.csv")
show_col_types = FALSE

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Pivot to long format since we have multiple years with DBH and status as column headers

```{r}

long_data <- trees %>%
  
  # take out columns we don't care about
  select(-Sapflow, -starts_with("Notes_")) %>%
  
    pivot_longer(
      cols = matches("_(\\d{4})$"),       # columns ending with a year
      names_to = c(".value", "Year"),
      names_pattern = "^(.*)_(\\d{4})$"
    ) %>%
    mutate(
      Year = as.integer(Year)
    ) %>%

  # filter out rows that have blanks and NA's in the  dbh column
   filter(if_all(starts_with("DBH"), ~ !is.na(.) 
                 & . != ""))
# convert year to integer
 long_data <- long_data %>%
  mutate(Year = as.integer(Year))

```

# Standardize status codes

```{r}
 long_data <- long_data %>%
    mutate(
      Status = toupper(trimws(Status)),  # LI, DS, DC, DN
      Status_code = case_when(
        Status == "LI" ~ 1,                   # alive
        Status %in% c("DS", "DC", "DN") ~ 0, # dead
        TRUE ~ NA_real_                       # anything else
      )
    )
```

# Calculate growth

```{r}
long_data <- long_data %>%
    arrange(Site, Plot, Tag, Year) %>%
    group_by(Site, Plot, Tag) %>%
    mutate(
      growth_cm = DBH_cm - lag(DBH_cm),
      annual_growth = growth_cm / (Year - lag(Year))
    ) %>%
    ungroup()
```

# Calculate mortality

```{r}
 mortality_data <- long_data %>%
    arrange(Site, Plot, Tag, Year) %>%
    group_by(Site, Plot, Tag) %>%
    mutate(
      prev_status_code = lag(Status_code),
      died_this_year = Status_code == 0 & prev_status_code == 1,
      first_observed_dead = is.na(prev_status_code) & Status_code == 0
    ) %>%
    ungroup()
```

# Site level growth and mortality

```{r}

#growth 

growth_summary <- long_data %>%
    group_by(Site, Year) %>%
    summarize(
      mean_growth = mean(annual_growth, na.rm = TRUE),
      median_growth = median(annual_growth, na.rm = TRUE),
      n_trees = sum(!is.na(annual_growth)),
      .groups = "drop"
    )

# visualize

ggplot(growth_summary, aes(x = Year, y = mean_growth, color = Site)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Mean Annual Growth by Site", y = "DBH Growth (cm/year)", x = "Year") +
  theme_minimal()

# mortality

 mortality_summary <- mortality_data %>%
    group_by(Site, Year) %>%
    summarize(
      n_alive_prev = sum(prev_status_code == 1, na.rm = TRUE),
      n_died = sum(died_this_year, na.rm = TRUE),
      mortality_rate = n_died / n_alive_prev,
      .groups = "drop"
    )


```

# Plot level growth and mortality

```{r}

# growth

plot_growth_summary <- long_data %>%
    group_by(Site, Plot, Year) %>%
    summarize(
      mean_growth = mean(annual_growth, na.rm = TRUE),
      n_trees = sum(!is.na(annual_growth)),
      .groups = "drop"
    )

# mortality

 plot_mortality_summary <- mortality_data %>%
    group_by(Site, Plot, Year) %>%
    summarize(
      n_alive_prev = sum(prev_status_code == 1, na.rm = TRUE),
      n_died = sum(died_this_year, na.rm = TRUE),
      mortality_rate = n_died / n_alive_prev,
      .groups = "drop"
    )
 
 # visualize heatmap
 
 ggplot(plot_mortality_summary, aes(x = Year, y = Plot, fill = mortality_rate)) +
  geom_tile(color = "white") +
  facet_wrap(~Site) +
  scale_fill_viridis_c(option = "plasma", labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Plot-level Mortality Rate by Year", y = "Plot", x = "Year", fill = "Mortality") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
```

# Growth vs. Mortality

```{r}
plot_growth_vs_mortality <- function(growth_summary, mortality_summary) {
  
  # Merge site-level summaries
  site_summary <- growth_summary %>%
    left_join(mortality_summary, by = c("Site", "Year"))
  
  # Scatter plot: Mean growth vs Mortality rate
  ggplot(site_summary, aes(x = mean_growth, y = mortality_rate, color = Site)) +
    geom_point(size = 3) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "gray30") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      title = "Site-level Growth vs Mortality",
      x = "Mean Annual Growth (cm/year)",
      y = "Mortality Rate",
      color = "Site"
    ) +
    theme_minimal()
}
```



## Auto plot dashboard function

```{r}

library(ggplot2)
library(scales)
library(gridExtra)  # for arranging multiple plots

plot_dashboard <- function(growth_summary, mortality_summary, plot_growth_summary = NULL, plot_mortality_summary = NULL) {
  
  # -----------------------------
  # 1. Site-level growth line plot
  # -----------------------------
  p1 <- ggplot(growth_summary, aes(x = Year, y = mean_growth, color = Site)) +
    geom_line(size = 1.2) +
    geom_point(size = 2) +
    labs(title = "Mean Annual Growth by Site", y = "DBH Growth (cm/year)", x = "Year") +
    theme_minimal()
  
  # -----------------------------
  # 2. Site-level mortality line plot
  # -----------------------------
  p2 <- ggplot(mortality_summary, aes(x = Year, y = mortality_rate, color = Site)) +
    geom_line(size = 1.2) +
    geom_point(size = 2) +
    scale_y_continuous(labels = percent_format(accuracy = 1)) +
    labs(title = "Annual Mortality Rate by Site", y = "Mortality Rate", x = "Year") +
    theme_minimal()
  
  plots <- list(p1, p2)
  
  # -----------------------------
  # Optional: Plot-level growth heatmap
  # -----------------------------
  if (!is.null(plot_growth_summary)) {
    p3 <- ggplot(plot_growth_summary, aes(x = Year, y = Plot, fill = mean_growth)) +
      geom_tile(color = "white") +
      facet_wrap(~Site) +
      scale_fill_viridis_c(option = "viridis") +
      labs(title = "Plot-level Growth Heatmap", y = "Plot", x = "Year", fill = "Growth") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    plots <- c(plots, list(p3))
  }
  
  # -----------------------------
  # Optional: Plot-level mortality heatmap
  # -----------------------------
  if (!is.null(plot_mortality_summary)) {
    p4 <- ggplot(plot_mortality_summary, aes(x = Year, y = Plot, fill = mortality_rate)) +
      geom_tile(color = "white") +
      facet_wrap(~Site) +
      scale_fill_viridis_c(option = "plasma", labels = percent_format(accuracy = 1)) +
      labs(title = "Plot-level Mortality Heatmap", y = "Plot", x = "Year", fill = "Mortality") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    plots <- c(plots, list(p4))
  }
  
  # Arrange all plots in a grid
  gridExtra::grid.arrange(grobs = plots, ncol = 1)
}

```














